---
title: "Bradford: GVA vs jobs 2D percent change plots"
author:
  - name: "Dan Olner"
format: 
  html:
    output-file: "Bradford_2D_gvajobs_percentchangeplots_July2025"
    fig-width: 9
    fig-height: 9
execute:
  echo: false
  message: false
  warning: false
fig-responsive: false
---

```{r setup}
library(tidyverse)
library(sf)#for geo stuff
library(zoo)#for making smoothed moving averages
library(patchwork)#combine ggplots easily
# library(plotly)#For interactive plots
library(RColorBrewer)

source('../functions/misc_functions.R')
source('../functions/adhoc_functions.R')
#source('functions/misc_functions.R')
#source('functions/adhoc_functions.R')

options(scipen = 999)

#Set ggplot theme
theme_set(theme_light())




```


## GVA (chained volume) vs jobs percent change plots: SIC 2 digit sectors

These plots show **percent change in chained volume GVA and job count** on different axes, with **Bradford's change overlaid in blue** and job count plus percent of regional economy included in title (for the most recent year). It uses a 3 year moving average to smooth out jumps.

Only places where a sector makes up 1.5% or more of the economy are included - except for Bradford, which is included in all plots.

This is done for the shorter 2 digit SIC categories in the regional GVA data, joined to BRES jobs data.

Any arrows in the **top left grey diagonal section** will have seen productivity drops (measured as GVA per FT job) over that time. In the bottom left diagonal, productivity went up.

Arrows in the bottom right quadrant have seen **a combination of productivity increases with job decreases**.

**NOTE: All sectors are included even if they're very small proportions of Bradford's economy - so percent changes could appear large for very tiny sectors. Look at the job count in the title to check. Including them all for completeness.**


```{r plots}
# OUTPUT ALL GVA * JOBS 2D PLOTS FOR BRAD SECTORS OVER 1% OF GVA----

bres.gva.2digit.2023 <- readRDS('../data/regionalGVA_plus_BRESjobcounts/regionalGVA_chainedvolume_BRES_FT_jobcount_bespoke2digitSIC_nONLY_MATCHING_GEOGs_2015_2023.rds') %>% 
  filter(JOBCOUNT > 0, !qg('membership|activities of households|agri', SIC07_description)) %>% 
  mutate(gvaperjob = GVA / JOBCOUNT)

shortsectornames <- read_csv('../data/shortsectornames_for_regionalGVA_2digitSICs.csv')

#check match, tick
# table(unique(bres.gva.2digit.2023$SIC07_description) %in% shortsectornames$SIC07_description)
bres.gva.2digit.2023 <- bres.gva.2digit.2023 %>% 
  left_join(
    shortsectornames, by = 'SIC07_description'
  )

#Add in sector GVA proportions to be able to filter by sector size
bres.gva.2digit.2023 <- bres.gva.2digit.2023 %>% 
  group_split(DATE) %>%
  map(add_location_quotient_and_proportions,
      regionvar = Region_name,
      lq_var = SIC07_description_shortened,
      # valuevar = GVA) %>% #This is CV GVA, can't use that for proportions, involves summing!
      valuevar = JOBCOUNT) %>% 
  bind_rows() %>% 
  select(DATE:SIC07_description_shortened,JOBS_sector_regional_proportion = sector_regional_proportion)


smoothband = 3

gva.jobs.ITL3.for2dplot <- bres.gva.2digit.2023 %>% 
  arrange(DATE) %>% 
  group_by(Region_name,SIC07_description_shortened) %>%
  mutate(
    JOBS_sector_regional_percent_movingav = rollapply(JOBS_sector_regional_proportion * 100,smoothband,mean,align='center',fill=NA),
    jobcount_movingav = rollapply(JOBCOUNT,smoothband,mean,align='center',fill=NA),
    gva_movingav = rollapply(GVA,smoothband,mean,align='center',fill=NA),
    `gva/job moving av` = rollapply(gvaperjob * 1000,smoothband,mean,align='center',fill=NA)
  ) %>% 
  ungroup()


#Err
# gva.jobs.ITL3.for2dplot %>% filter(qg('accomm|admin',SIC07_description_shortened), Region_name == 'Bradford', DATE == 2022) %>% View

#Filter by sector proportion for sector we're looking at, in latest smoothed years

# selectedsector = getdistinct('chemicals', gva.jobs.ITL3.for2dplot$SIC07_description_shortened)
# selectedsector = getdistinct('transp manuf', gva.jobs.ITL3.for2dplot$SIC07_description_shortened)
# selectedsector = getdistinct('food manuf', gva.jobs.ITL3.for2dplot$SIC07_description_shortened)
# selectedsector = getdistinct('metals', gva.jobs.ITL3.for2dplot$SIC07_description_shortened)
# selectedsector = getdistinct('head office', gva.jobs.ITL3.for2dplot$SIC07_description_shortened)

#Saving for use elsewhere!
#saveRDS(gva.jobs.ITL3.for2dplot,'local/data/gvajobsfor2dplot.rds')


for(sector in unique(bres.gva.2digit.2023$SIC07_description_shortened)[order(unique(bres.gva.2digit.2023$SIC07_description_shortened))]){

  placestokeep <- gva.jobs.ITL3.for2dplot %>% 
    filter(!is.na(jobcount_movingav)) %>% 
    filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
    filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
    select(Region_name) %>% 
    distinct() %>% 
    pull
  
  #Always plot Bradford! Add in if not there
  if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")
  
  #Add into 2D percent plot
  p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
  
  print(p)
  
  # ggsave(paste0("local/outputs/gva_jobs_2dpercentplots_Bradfordhighlight/chainedvol/",gsub('/| ','_',sector),".png"), dpi = 150, width = 9)

   

}#end for


```
