---
title: "Bradford sectors and clusters: past, present"
author:
  - name: "Dan Olner"
format: 
  html:
    output-file: "Bradford_sectorclusters_July2025"
    fig-width: 8
    fig-height: 8
execute:
  echo: false
  message: false
  warning: false
fig-responsive: false
collapse: false
---

```{r setup}
library(tidyverse)
# library(spdep)
library(sf)#for geo stuff
# library(tmap)#for mapping 
library(zoo)#for making smoothed moving averages
library(patchwork)#combine ggplots easily
library(plotly)#For interactive plots
# # library(ggdist)
# library(tidyr)
# library(distributional)
library(RColorBrewer)

source('../functions/misc_functions.R')
source('../functions/adhoc_functions.R')
#source('functions/misc_functions.R')
#source('functions/adhoc_functions.R')

options(scipen = 999)

#Set ggplot theme
theme_set(theme_light())


```

# INTRODUCTION

This document contains the following sections - all accessible via the table of contents on the right hand side of this page. **Click on a top-level heading to open up its subheadings.**

1. A summary/synthesis of the findings 
2. Connections to other sector strategies - including a detailed data driven link to the UK Industrial Strategy
3. Headlines from the data analysis, including a sector-by-sector breakdown of sectors (examining strongest and other important sectors). Hover on figure references for a preview or click to jump to them.
4. A series of data sections used as the basis for the analysis - see the 'data' section at the top of the headlines below for a full explanation of what has been used.

Here are links to other sources used in the analysis below to look at Bradford's sectors/clusters in context (all open in new tabs).

1. A couple of data catalogues with plots for the full range of sectors in Bradford. These are:
    - [Full 5 digit SIC employment location quotients](https://danolner.github.io/RegionalEconomicTools/quarto_docs/Bradford_BRES_5digit_LQs_July2025.html){target="_blank"} broken down by parent 2 digit SIC.
    - [GVA vs job percent change 2D plots](https://danolner.github.io/RegionalEconomicTools/quarto_docs/Bradford_2D_gvajobs_percentchangeplots_July2025.html){target="_blank"}. These show how much (inflation-adjusted) GVA and job counts have changed between their 2015-2017 average and 2021-23 average for ITL3 zones, with Bradford highlighted. Putting GVA and job percent change on different axes also shows exactly how productivity (GVA per full-time job) has changed and whether productivity gains have included job gains or losses.
2. A set of treemaps showing the relative size of sectors in Bradford (as well as comparing Bradford to other ITL3s in West Yorkshire). These are:
    - Treemap for [BRES job count data (2023) for West Yorkshire ITL3 zones](https://danolner.github.io/RegionalEconomicTools/miscdocs/WYLAs_BRES_treemap.html){target="_blank"} including Bradford - sectors with matching colours. Click to drill down into place then 3 SIC levels (section, 2 digit and 5 digit).
    - Treemap for [Companies House job count data (up to end of June 2025) for West Yorkshire ITL3 zones](https://danolner.github.io/RegionalEconomicTools/miscdocs/WYLAs_CompaniesHouse2025_treemap.html){target="_blank"}.
3. [Companies House map](https://danolner.github.io/companieshouseopen/maps/companieshouse_modal_SICsection_byemployeecount_min10plus_hex1000m_wLADs.html){target="_blank"} showing most common broad sector per 100m^2 hex. This is useful for getting a sense of Bradford's economic geography and the surrounding areas.




# SUMMARY

TBW!


# CONNECTION TO OTHER SECTOR STRATEGIES

**Including a detailed link to the UK Industrial Stategy**

## UK Industrial Strategy

### Overview

The [UK Industrial Strategy](https://www.gov.uk/government/collections/the-uks-modern-industrial-strategy-2025) was released in 2025, with accompanying plans for 6 out of the 8 target "IS-8" sectors. It also has a [sector definition list](https://www.gov.uk/government/publications/industrial-strategy/industrial-strategy-sector-definitions-list) that provides SIC codes being targeted - top level sectors and 'frontier' subsectors (a few of which overlap; most do not). These range in specificity between 2 digit and 5 digit SIC codes.

As that webpage says, most of these SIC codes cannot capture the full range of possible firms the strategy aims to cover, and it also contains entire sections that SIC codes can't cover at all - clean energy and certain digital sectors in particular. However, it does allow an analysis of which sectors are *definitely* within those industrial strategy categories. This section gathers job details for those specific sectors - a distinct range of types and breadth - and compares Bradford to the rest of the UK where possible, to see which sectors align with the industrial strategy and other analysis above. This is done for both BRES and Companies House job counts, to pick up on different job types.

This document has [a data section ("UK INDUSTRIAL STRATEGY LINKAGE")](https://danolner.github.io/RegionalEconomicTools/quarto_docs/Bradford_sectorclusters_July2025.html#uk-industrial-strategy-linkage) that pulls out Bradford sectors in this 'IndStrat' list - giving job counts, LQs and change over time from both BRES and Companies House data.

The table below lists the IS-8 broad sector headings. In the IndStrat plots, the abbreviations for these IS-8 headings appear before the sector name. Clean energy - as mentioned above - has no SIC codes and so can't currently be matched. There are no matching defence sectors or firms in Bradford.

```{r brad_indstrattable}

indstrat <- tibble(
  `IS8 sector` = c(
    'Advanced Manufacturing',
    'Clean Energy Industries',
    'Creative Industries',
    'Defence',
    'Digital and Technologies', 
    'Financial Services',
    'Life Sciences', 
    'Professional and Business Services'),
  `Abbreviation` = c(
    'ADV MANUF',
    '--',
    'CREATIVE',
    '--',
    'DIGITAL',
    'FIN',
    'LIFESCI',
    'PBS'
    )
)

#escape = FALSE so that markdown in the cells (the links) is rendered correctly.
knitr::kable(indstrat, escape = FALSE, caption = "Table: Industrial Strategy IS-8 names and abbreviations (dashes mean that sector doesn't appear for Bradford)")

```


Also:

* The LQ plots only shows IndStrat SIC sectors where Bradford has 100+ employees (for 2021-23 average)
* And breaks them down into 'frontier' sectors and the others (a few of which overlap)



### West-yorkshire relevant points in the Industrial Strategy

Here are bullet points from the main Industrial Strategy policy document and the accompanying sector strategy docs (see link at the start of this section) where mentions of West Yorkshire occur (WY authorities are only mentioned once - namely Wakefield's Production Park, see below).

**From main document:**

* "The British Business Bank will introduce a new Cluster Champions programme in 10 places"  including West Yorkshire. (p.98)
* "Champions with deep expertise and local knowledge will coordinate investment-readiness programmes, strengthen financial networks, and connect high-potential firms in the IS-8 to investors" (again, WY included: "Greater Manchester, West Yorkshire, the West Midlands, Liverpool City Region, South Yorkshire, North East, West of England, Glasgow City Region, Cardiff City Region, and Belfast City Region.")
* National wealth fund available WY (p.99)
* Local Innovation Partnerships Fund - WY (p.100)
* Intracity connectivity / creative place growth fund (part for WY, p.109)
* A "New Professional and Business Services Hubs in Greater Manchester, Liverpool City Region, and **West Yorkshire**, and piloting Made Smarter for Professional and Business Services across the North of England." (p.110,142/3)
* WY financial services support (p.136)

**PBS sectors plan:**

Aleady mentioned, but a little more detail: "Delivering a new technology adoption programme for PBS SMEs, starting with a pilot in the North of England (North West, North East, Yorkshire and the Humber, excluding Lincolnshire authorities), supporting a corridor across our Northern city regions. This targeted and intensive support programme will equip high‑growth potential PBS SMEs with the skills, knowledge, and resources necessary to adopt AI and PBS-specific digital technologies." Including support for tradeable work.

**Creative industries sector plan:**

* Supporting "cutting-edge [createch](https://www.thecreativeindustries.co.uk/createch) in West Midlands, West Yorkshire and Bristol and Bath"
* Description of a WY cluster: "West Yorkshire: Music, Film & TV, Video Games, Createch, Fashion/ Textiles. Production Park in Wakefield, host of a COSTAR Live lab, deploys pioneering technology to develop arena tours, film and TV. Channel 4's Leeds HG equips regional talent with industry-leading training and opportunities. Home to Game Republic Ltd, which operates two leading games networks across the North of England." (p.62)
* (Bradford gets a photo in this one - Les Girafes, p.51)

**Digital and technologies sector plan:**

* Again: "Local Innovation Partnerships Fund, a new programme to grow high-potential innovation clusters across the UK. This will empower local leaders to decide how to target R&D investment in their region and unleash their full innovation potential. The fund earmarks at least £30 million each for 10 regions [including] West Yorkshire." (pp.60-61)
* And the cluster champions programme mentioned above. pp.60-61

**Life Sciences sector plan:**

* "Continued support for Life Sciences-focused Investment Zones in: West Yorkshire" (and Liverpool City Region). p.54




### Points from the IndStrat LQ plots for Bradford

The IndStrat LQ plots pick out for Bradford which IS-sectors are more concentrated job-wise, what their trajectory is and what the job counts are. This section can be read in conjunction with the more detailed data/sector analysis below, from which it draws.

There are three sectors in particular identified below as being important to Bradford but facing productivity challenges that could benefit from support flowing from the industrial strategy: **chemical production**, **head offices / management consultancy** and **architect/engineering/tech testing** services. See the analysis below for a description of the history of these sectors.

* Bradford has both manufacturing and service strengths that marry with the industrial strategy.
* From those examined below, Bradford's IS-8 overlaps are:
    - PBS: across a number of sectors, including **legal services, accounting, head offices and management consultancy**.
    - Finance: as well as **banks**, Companies House accounts picks out a very large number of **holding companies** in the finance category.
    - Advanced manufacturing: the **machinery / equipment not elsewhere classified** SIC covers a lot of advanced firms including space sectors, and is 3000 jobs in BRES. While **architect/engineering/tech testing** is technically PBS, it has important connections to production and faces challenges. **Chemical production** has already been mentioned as key to the industrial strategy but a nationally struggling sector.
    - Digital: **electronic instuments/navigation** isn't a large number of jobs relatively, but it is likely a part of the larger advanced manufacturing / digital tech crossover.




## West Yorkshire [Local Growth Plan](https://www.westyorks-ca.gov.uk/growing-the-economy/economic-strategy/local-growth-plan/)

The LGP's "priority one" is [unlocking growth in sectors](https://www.westyorks-ca.gov.uk/growing-the-economy/economic-strategy/local-growth-plan/priority-one-unlocking-growth-in-sectors/). 

That page has a detailed outline of subsectors across West Yorkshire, and it argues for building on those with the largest job presence, including **textiles** and **food manufacture**, as well as **chemicals**. Each of these sectors are identified below as facing long-standing productivity challenges; here, other advanced manufacturing sectors are also highlighted as vital for Bradford (especially **metals** and **machinery/transport manufacture** - the WY LGP cites **space tech** too, a lot of which will be in these categories). **Chemical production** is a focus of the UK Industrial Strategy - including as part of its "foundational inputs" category - and so could potentially benefit the most, out of all high-job-count / low-output sectors listed here.

The LGP also singles out **financial and professional services** as vital, and for Bradford, the analysis below absolutely bears this out - though again, with some high growth slope sectors facing productivity issues (**head offices**), other Bradford strengths are in a much stronger position (**legal/accounting**) while Bradford finance is productive but possibly static (though Companies House data suggests a dynamic finance sector).

The **high growth cluster** analysis in the LGP puts "engineering ... at the heart of these opportunities as a driver and foundation of growth." As mentioned, engineering is an historic Bradford strength that is facing a long term productivity issue; understanding this better could help support this 'cluster action' approach.




## [West Yorkshire LSIP](https://www.wnychamber.co.uk/app/uploads/2023/08/LSIP_West_Yorkshire_2023.pdf) (Local Skills Improvement Plan) published 2023

Page 4 of the West Yorkshire LSIP says it aims “ro enable employers and individuals to access the technical skills they need to support growth, transformation and increased productivity in our key sectors, defined as:”

* Health and Social Care
* Engineering and Advanced Manufacturing
* Financial and Professional Services
* Low Carbon
* Digital and Technology
* Creative Industries
* Education
* Construction
* Transport and Logistics

The analysis below digs further into **engineering, advanced manufacturing** and **financial and professional services** - all Bradford sector strengths, with differing productivity challenges, though these are especially notable in engineering.

Other sectors listed in LSIP could go into two categories:

1. Broad sectors where, with the data here, it is difficult to shed much light on productivity issues. Health and social care faces the same issues as other mainly-public sectors in the data - they have clearly grown to be the largest or one of the largest broad sectors, but public sector productivity analysis is difficult (partly because of how ONS calculate it at a regional level - they just use job count times regional job average to estimate GVA). This in no way negates the need to improve health and social care skills, it just makes it hard to analyse productivity data.
    - Companies House data *does* show that both health and social care have a very large private sector presence in Bradford (@fig-percentempfourplaces, @fig-percentempfourplaces_lessthan10 and plots from @fig-indivfirmsSY on to identify specific firms), so these could well benefit / be benefiting from from LSIP-related support.
2. Other sectors have not been picked out below for specific analysis, but are high job count with national-level productivity issues, and often not highly tradeable beyond Bradford. See for example the 2D GVA vs job plots for transport (@fig-gvajobsTr) and construction (@fig-gvajobsSC).

The following sections dig into what the data shows for Bradford's sectors; plots and analyses follow that text.

# HEADLINES FROM THE DATA ANALYSIS

## The data

Three main data sources are used here (and one other [smaller ONS source](https://www.ons.gov.uk/file?uri=/employmentandlabourmarket/peopleinwork/labourproductivity/datasets/subregionalproductivitylabourproductivitygvaperhourworkedandgvaperfilledjobindicesbyuknuts2andnuts3subregions/current/labourproductivityitls1.xlsx) for the hourly productivity plots). 

The main three data sources are:

* [ONS regional GVA data](https://www.ons.gov.uk/economy/grossvalueaddedgva/datasets/nominalandrealregionalgrossvalueaddedbalancedbyindustry) at ITL3 geographies, up to 2023 (the most recent available data).
* [BRES](https://www.nomisweb.co.uk/datasets/newbres6pub) (Business Register and Employment Survey) data for job counts by sector and place, again up to 2023 for the latest.
* Data extracted from [Companies House accounts](https://download.companieshouse.gov.uk/en_monthlyaccountsdata.html) for individual firms in the last year, between June 2024 and end of June 2025. This data only contains firms with a registered address in Bradford, so it misses all firms with ownership registered elsewhere (e.g. this includes [Produmax](https://find-and-update.company-information.service.gov.uk/company/00878102), registered in Manchester). However, it can provide a highly granular and up-to-date picture of sector economies, including employment change in the last year.

For some analyses, **regional GVA and BRES jobs data are linked**. To do this, BRES jobs are binned into the smaller list of 44 SIC groupings that the ONS uses for the GVA data, so e.g. GVA per full time job can be examined. Where possible, the most granular sector breakdowns are used.

Lastly, the UK government's Industrial Strategy [sector definition list](https://www.gov.uk/government/publications/industrial-strategy/industrial-strategy-sector-definitions-list) is used to examine how Bradford's economy meshes with that strategy.

All code for this analysis is available [here](https://github.com/DanOlner/RegionalEconomicTools), and [here](https://github.com/DanOlner/companieshouseopen) for processing Companies House accounts.





## Service sectors

### Strongest service sectors

* The **head offices and management consultancy** sector (SIC 70, 2.3% of FT jobs, 2.7% of GVA; @fig-otherLQs) jumps out in several of the plots below (looking at [5 digit job counts](https://danolner.github.io/RegionalEconomicTools/quarto_docs/Bradford_BRES_5digit_LQs_July2025.html) for this sector, this is mostly 'head office activities' in Bradford). While not a very productivity sector overall, it has the strongest GVA and jobs growth slopes of any sector, averaging 15% GVA and 16.7% GVA per FT job growth from 2015 to 2023 (@fig-growthdata_cv_andperFT), statistically separable from most other places, and becoming even stronger in more recent years (@fig-growthdata_cv_andperFT_later). Plotting its GVA percent change (2015/17 to 2021/23 averages) against the same for jobs (@fig-gvajobshoc) shows (in the lighter diagonal corner) how its productivity gains have come from GVA growth mostly, but including a jobs increases, and that Bradford's trajectory has been one of the strongest, percentage-wise. @fig-gvajobs1 (last panel) shows output per worker for the sector has been historically low everywhere, but Bradford has seen a large jump in trend since COVID, and it is now some way above the average for other ITL3 zones in the data (50K per FT versus 36).
    - Location quotients for the sector show a contrast (@fig-otherLQs): GVA continuing to increase its concentration, but more recently Bradford's jobs in this sector being less concentrated in Bradford than the GB average (though the jobs LQ has hovered between more and less over time). The last year's Companies House (CH) data, though, sees the sector growing its job count more than any other in Bradford, and more than other West Yorkshire authorities for firms with 10+ employees (@fig-percentempfourplaces) - only a small number of firms in CH (8 showing a 14.2% employee increase overall) but indicative of sector growth.
* **Finance** should possibly be at the top of this list, as it is 5.4% of Bradford's economy and 3.7% of its full time workforce (4900 jobs in BRES; @fig-otherLQs). However, it doesn't stand out in comparison to other Bradford sectors or the rest of GB. For example, in its 2D GVA vs jobs percent change plot (@fig-gvajobsFin), while its productivity has risen, it's only a little - though note, it is at least in the more positive half in that plot, compared to many other places with productivity drops (in the grey top left diagonal).
    - The 'finance' SIC category in the GVA data covers SICs 64 to 66 (finance, auxiliaries and insurance) - looking deeper into the BRES 2 and 5 digit lists, building societies are the largest job chunk (1200 / 0.85%) and have been concentrated in Bradford (near the top of UK concentration) for the full data time period. There are actually more jobs in the 'banks' category (2400, 1,8%) but that is only a little above the GB average. Financial auxiliary jobs have been shrinking in number, and statistically more than most other places (@fig-growthdata_bres2dig) and insurance is a very small number of jobs.
    - Companies House data also shows strong jobs growth in the last year (only Wakefield ahead in WY) - 14 firms with 10+ employees going from 1434 to 1558 employees (an 8.65% increase, @fig-percentempfourplaces). (CH data also shows over 300 "fewer than 10 employees") firms / lone traders, @fig-percentempfourplaces_lessthan10).
* **Legal and accounting** is a stand-out related sector in the PBS (Professional and Business Services) category (average 3.2% of jobs in 2021-23, 1.8% of total).  With historically very high output per FT worker in Bradford (@fig-gvajobs1 panel 5), it has dropped more recently but is still very high - 93.5K per FT versus 78.4K average for other ITL3 zones.
    - The sector has grown jobs only modestly over the data time period, but it started from a high GVA and jobs baseline. Digging into the 5 digit SICs for the sector, 'accounting/auditing' (1150 jobs in latest BRES year) has been growing its job concentration strongly - still below average GB concentration, but heading higher and very high output. BRES records 833 workers in 'solicitors', but the trajectory is not as strong as accounting. 
    - Companies House data for the last year shows growth in "fewer than 10 employee" firms of 7.4% (now up to 670 employees), though 10+ employee firms have been fairly static.



### Notes on other service sectors

* **Real estate** has a high output per worker everywhere (with imputed rent removed, so this is purely real estate itself, not a reflection of rising house prices). This should arguably be in "strongest service sectors", but it is difficult to discern from the data where the productivity is coming from. Exactly half of the job count in this sector is in 'housing association renting'. Jobs *are* increasing in each of the other subsectors (real estate, agencies, other letting) which are another 1000 jobs and, looking in Companies House data, there are firms like [Dacre, Son and Hartley](https://find-and-update.company-information.service.gov.uk/company/03090769) who employ 100 people and are clearly strong. Job numbers have risen since 2015/17 but output hasn't risen quite as quickly (see GVA vs jobs plot, @fig-gvajobsrealestate). The sector is also productive even by UK standards (195K per FT worker versus 167K). But in this data, it cannot be seen what role housing associations versus other real estate is playing in this.
* It is tricky to define **architecture/engineering/tech testing's** place in Bradford (SIC 71). It is a large chunk of Bradford's economy - more than 2000 jobs, 1.5% of full time workforce in latest BRES year, and has historically kept around the same number. However - while its GVA per worker in @fig-gvajobs1 was very high in 2015-17, it has dropped off constantly since (while other places in WY have slowly climbed), now approaching the same output per worker as head offices / consultancy (though it is still above the average of other ITL3s). Jobs in the sector have grown in most places, but Bradford is some way into the group where GVA has dropped (@fig-gvajobsarcheng). There has been a jump in job numbers in the most recent data but no GVA jump, suggesting productivity continues to decline. Companies House shows a large number of SMEs in the sector, with modest jobs growth. But this sector (despite technically being under a 'services' heading) is obviously important to manufacturing and other productive output.
    - What factors underpin this steady slope? This data isn't able to say directly, but identifying what's happening here could be a first step to addressing it.
* **Utilities and power** (5.7% of Bradford's GVA, 2500 jobs in BRES) are tricky sectors to analyse. While the sector appears highly productive in Bradford (187K per FT job versus 138K GB average), differences in this sector largely come down to the capital intensity of the sector. There can be innovative frontier firms in these SIC codes, but the fact that very few firms in these sectors appear in the Companies House data suggests this sector is largely external utilities in the area, and could be difficult to affect.
* **Research / advertising**, while apparently a high productivity sector, is fewer than 300 jobs, and this SIC misses other places where important research activity takes place.
* **Other professional, scientific and technical activities** ("Other PSTech", so any other apart from head offices, consultancy, legal, engineering - see the [BRES job treemap for 'professional/sci/techn'](https://danolner.github.io/RegionalEconomicTools/miscdocs/WYLAs_BRES_treemap.html)) appears productive, but is only a very small number of jobs. Bradford has been following the same trajectory as West Yorkshire, output per worker growing, and this is bucking the national trend where in places with more than 1% GVA in this sector, jobs have grown but productivity dropped. But it's a very small sector.
* **Office administration and business support activities** ("Admin" in the plots below; distinct from "public admin") is not highly productive but its trajectory is worth noting. As @fig-gvajobsadmin shows, its jobs/GVA combination has led to a low but steady GVA per FT job increase (where a lot of other places have seen declines). The fact that this sector is so many jobs (3200 average 2021-23), and its connections with other PBS sectors, makes it worth noting.


## Manufacturing / production sectors

Despite an overall decline in the proportion of Bradford's broad manufacturing economy (in line with the most of the rest of the UK and West Yorkshire, @fig-struct), there is important production sector activity to highlight.

### Strongest sectors

* The **machinery and transport manufacture** sector (SIC 28-30, ~2.5% of jobs in 2021-23 average) has been growing in productivity since 2015/17, on a quite different upward trajectory from the rest of West Yorkshire (@fig-gvajobs1 panel 4). It has been growing as a proportion of GVA and jobs in Bradford and its GVA has been rising steadily in the 'increasing productivity' zone (@fig-gvajobs_mtf) where most other places have seen productivity drops. It is now at or near the UK average sector productivity - a rare position for a Northern manufacturing sector (though Companies House data for the last year doesn't show quite as strong performance).
    - Two thirds of these jobs are in machinery manufacturing, and every 5 digit subsector in that is increasing in concentration (only a tiny proportion in motor vehicles, according to BRES). 
* **Metals** sectors (SIC 2 digit 13-15) has a high LQ and ~2.6% of Bradford jobs (@fig-productionLQs). It has had continuous GVA and jobs growth - an average of 4.1% yearly GVA growth (1.1-6.9%, @fig-growthdata_cv_andperFT) from 2015, increasing more since 2019. This covers many cutting-edge firms like Produmax, who classify themselves as 'manufacture of other fabricated metal products'. (The 5 digit SIC jobs breakdown for fabricated metals shows this category is where the largest chunk of the fab metals jobs are in Bradford; BRES doesn't record many jobs in basic metals).
    - However, the metals sector has low average output per worker. As the 2D GVA/jobs plots for metals shows (@fig-gvajobsmetals), the sector has maintained an output increase alongside its productivity - right in the centre of other places seeing productivity losses and gains. This mean its productivity (gva/job) hasn't actually changed much, while job and GVA values have both increased by more than 60% since 2015/17.
* **Textiles** continue to grow GVA and jobs concentration (GVA more than jobs, @fig-productionLQs) though the sector isn't quite 1% of total full time jobs. Companies House data, however, suggests the sector is probably larger than BRES' survey is picking up on - there are nearly 1000 jobs in apparel firms in CH (BRES finds only 20) - so it may well be more like 2% of total jobs.
    - Historically, it has been higher than average GVA per job for Bradford, with all West Yorkshire authorities very close in output - but this has declined since COVID (@fig-gvajobs1) and is now almost exactly average for Bradford but a little below UK average for the sector. The 2D GVA/job plot for textiles (@fig-textiles) shines some light on this. In the very few UK places where textiles are more than 1% of the economy, none have better productivity in 2021/23 than they did in 2015/17. Some (e.g. Calderdale/Kirklees) have seen jobs increase while GVA has *dropped*. Along with Blackburn, Bradford is doing relatively well on this metric - growth continues though with apparent productivity issues.


### Notes on other manufacturing/production sectors

* As Bradford's largest 2-digit sector by job count, **Food/beverage manufacturing** has 3.6% of full time employers according to BRES, and 2.5% of its GVA. Worker growth looks especially strong in the last year in the Companies House data: the 14 firms with 10+ employees have grown nearly 9% (@fig-percentempfourplaces) and 74 microfirms with fewer than 10 employees have gone from 212 to 281 employees (@fig-percentempfourplaces_lessthan10), both higher than the rest of West Yorkshire. However, the sector overall has seen declining productivity, and Bradford's GVA per worker is at the lower end for West Yorkshire (along with Leeds - the other authorities are higher; @fig-gvajobs1). As the GVA vs jobs 2D plot for the sector shows (@fig-gvajobsFM), most places have seen job increases with either GVA shrinkage or too slow GVA growth for productivity gains; Bradford's GVA growth in the sector has been static. * Bradford's **chemicals industry** (SIC 19-23) has a similar pattern to architecture/engineering/tech testing - except in chemicals, job numbers continue to grow while GVA drops. Jobs have yo-yo'd but have been mostly rising in recent years - but GVA has dropped, leading to worsening productivity. Demand issues in recent years accounts for some of this (see e.g. [this report](https://cefic.org/landscape-of-the-industry/united-kingdom/)) - but this pattern of GVA decline has been in progress from 2015-17. 5 digit chemical sectors are mostly seeing jobs growth - rubber/plastic jobs increasing strongly (700+ now in two subsectors) and chemical manufacturing jobs in most subsectors continue to grow too. But - see @fig-gvajobsCh - Bradford has seen one of the largest output drops in GB (while, as everywhere else, seeing job numbers rise). This is clearly a national issue, but Bradford seems especially vulnerable.
* **Construction of buildings** ("Buildings" in the plots, SIC 41) is worth noting for its improving productivity (GVA growing faster than jobs but both growing, @fig-gvajobsB). The West Yorkshire pattern in @fig-gvajobs1 (panel 3) for output per full time job shows the region moving together, though Bradford is at the bottom of the sector's GVA per worker here - but it is also following the recent rise seen across the region.


```{r gvaperhour_dataprep}
# ITL3 Hours % change v GVA % change on average----

#So need the number of hours and chained volume GVA
#Hours count is in the per hour productivity sheet
#GVA in the region by industry sheet

#hours worked first
url1 <- 'https://www.ons.gov.uk/file?uri=/employmentandlabourmarket/peopleinwork/labourproductivity/datasets/subregionalproductivitylabourproductivitygvaperhourworkedandgvaperfilledjobindicesbyuknuts2andnuts3subregions/current/labourproductivityitls1.xlsx'
p1f <- tempfile(fileext=".xlsx")
download.file(url1, p1f, mode="wb")

#"Productivity Hours Worked per Week; ITL2 and ITL3 subregions (constrained to ITL1), 2004 - 2023"
hoursworked <- readxl::read_excel(path = p1f,range = "Productivity Hours!A5:W247")

names(hoursworked) <- gsub(x = names(hoursworked), pattern = ' ', replacement = '_')

hoursworked = hoursworked %>%
  filter(ITL_level == 'ITL3') %>% 
  pivot_longer(Hours_2004:Hours_2023, names_to = 'year', values_to = 'hoursworkedperweek') %>% #get most recent year
  mutate(
    year = substr(year,7,10),
    year = as.numeric(year)
  ) %>% 
  select(-ITL_level)



#Then ITL3 CV for all industries (so has to include imputed rent as we can't sum CV)
url2 <- 'https://www.ons.gov.uk/file?uri=/economy/grossvalueaddedgva/datasets/nominalandrealregionalgrossvalueaddedbalancedbyindustry/current/regionalgrossvalueaddedbalancedbyindustryandallinternationalterritoriallevelsitlregions.xlsx'
p2f <- tempfile(fileext=".xlsx")
download.file(url2, p2f, mode="wb") 

#"Table 3b: ITL3, chained volume measures in 2019 money value, pounds million [note 2]"
itl3.all <- readxl::read_excel(path = p2f, range = "Table 3b!A2:AD11468")

names(itl3.all) <- gsub(x = names(itl3.all), pattern = ' ', replacement = '_')

itl3.all = itl3.all %>% 
  filter(SIC07_description == 'All industries') %>% 
  pivot_longer(`1998`:names(itl3.all)[length(names(itl3.all))], names_to = 'year', values_to = 'value') %>% #get most recent year
  mutate(year = as.numeric(year)) %>% 
  rename(gva_cv = value)



#Check link between two is valid... tick
# table(unique(hoursworked$ITL_code) %in% itl3.all$ITL_code)

#2004:2023
# unique(hoursworked$year)[unique(hoursworked$year) %in% unique(itl3.all$year)]

#Get list of corecities
corecities <- getdistinct('sheffield|Belfast|Birmingham|Bristol|Cardiff|Glasgow|Leeds|Liverpool|Manchester|Tyne|Nottingham', itl3.all$Region_name)

#Yes, all in there, need to remove a few...
corecities <- corecities[!grepl(x = corecities, pattern = 'Greater|shire', ignore.case = T)]
corecities <- corecities[order(corecities)]

#Check match also works in hours worked... tick
# table(corecities %in% hoursworked$Region_name)



#Get London ITL3s
itl_lookup = read_csv("../data/LAD_(December_2024)_to_LAU1_to_ITL3_to_ITL2_to_ITL1_(January_2025)_Lookup_in_the_UK.csv")
#itl_lookup = read_csv("data/LAD_(December_2024)_to_LAU1_to_ITL3_to_ITL2_to_ITL1_(January_2025)_Lookup_in_the_UK.csv")

#Get all London ITL2 zones then use to get unique ITL3 zones
itl3.lon = itl_lookup %>%
  filter(qg('london',ITL225NM)) %>% 
  select(ITL325CD,ITL325NM) %>% 
  distinct()

#Check itl3 code match... tick
# table(itl3.lon$ITL325CD %in% hoursworked$ITL_code)
# table(itl3.lon$ITL325CD %in% itl3.all$ITL_code)
#tick
# table(itl3.lon$ITL325NM %in% hoursworked$Region_name)
# table(itl3.lon$ITL325NM %in% itl3.all$Region_name)


```






# ONS GVA AND JOBS DATA

## Structural /  broad sector GVA change 

```{r, fig-struct}
#| fig-width: 8
#| fig-height: 16

# LOAD ITL3 GVA DATA----

#ITL3 zones, SIC sections, current prices, 2023 latest data (minus imputed rent)
df = read_csv('https://bit.ly/rtasteritl3noir')

#This one has imputed rent included if you want it
# df = read_csv('https://bit.ly/rtasteritl3')



#Finding location quotients for each year in the data 
#And then overwriting the df variable with the result
df = df %>%
  group_split(year) %>%
  map(add_location_quotient_and_proportions,
      regionvar = Region_name,
      lq_var = SIC07_description,
      valuevar = value) %>% 
  bind_rows()


#1. Set window for rolling average: 3 years
smoothband = 3

#2. Make new column with rolling average in (applying to all places and sectors)
#Years are already in the right order, so will roll apply across those correctly
df = df %>% 
  group_by(Region_name,SIC07_description) %>% 
  mutate(
    sector_regional_prop_movingav = rollapply(
      sector_regional_proportion,smoothband,mean,align='center',fill=NA
      )
  ) %>% 
  ungroup()


#Get set pastel colours for each SIC section
n <- length(unique(df$SIC07_description))
set.seed(12)
qual_col_pals = RColorBrewer::brewer.pal.info[RColorBrewer::brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(RColorBrewer::brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

randomcols <- col_vector[10:(10+(n-1))]

#Do once!
getplot <- function(place){
  
  df = df %>% 
    filter(!qg("Activities of households|agri",SIC07_description)) %>% 
    mutate(
      SIC07_description = reduceSICnames(names = SIC07_description, level = 'section')
    )
  
  top12sectors = df %>%
    filter(
      Region_name == place,
      year == max(year) - 1#to get 3 year moving average value
    ) %>%
    arrange(-sector_regional_prop_movingav) %>%
    slice_head(n= 12) %>%
    select(SIC07_description) %>%
    pull()
  
  
  ggplot(df %>% 
           filter(
             Region_name == place, 
             SIC07_description %in% top12sectors,
             !is.na(sector_regional_prop_movingav)#Remove NAs created by 3 year moving average
           ), 
         aes(x = year, y = sector_regional_prop_movingav * 100, colour = fct_reorder(SIC07_description,-sector_regional_prop_movingav) )) +
    geom_point() +
    geom_line() +
    scale_color_manual(values = setNames(randomcols,unique(df$SIC07_description))) +#set manual pastel colours matching sector name
    ylab('Sector: percent of regional economy') +
    theme(plot.title = element_text(face = 'bold')) +
    labs(colour = 'SIC section') +
    # scale_y_log10() +
    ggtitle(paste0(place, ' GVA % of economy over time\nTop 12 sectors in ',max(df$year),'\n (', smoothband,' year moving average)'))
  
  
}

places = getdistinct('bradford|kirklees|calderdale|leeds|wakefield',df$Region_name)

plots <- map(places, getplot)

patchwork::wrap_plots(plots, ncol = 1)


```




## GVA per full time job over time: Bradford v other ITL3 zones for 2 digit SIC sectors

These plots use **linked regional GVA and BRES jobs data** to examine productivity, output and job changes. To make this link, sector categories have to be reduced to those in the ONS GVA data (46 categories rather than the full 88 two digit SICs).

These are measured in GVA per full time job (GVA / jobs), using inflation-adjusted (chained volume) GVA.

BRES does contain more detail - more on that below and in the separate [BRES 5 digit breakdown](https://danolner.github.io/RegionalEconomicTools/quarto_docs/Bradford_BRES_5digit_LQs_July2025.html){target="_blank"}.

GVA per job as a measure needs some caution: output is also due to other productive factors, so high capital intensity sectors (power, telecoms, other ICT though not all) can appear overly productive per worker. It's more representative for service sectors where an employee's output more clearly correlates to total output, such as legal services.


```{r, fig-gvajobs1}
#| fig-height: 20
#| fig-width: 8.5

#For current price version: The plots below are measured in **GVA per job parts per million of UK economy** - equivalent to e.g. 1% being "1 in 100 parts of economy", 0.1 here is "1 in 10 millionth of whole economy". So it's again examining **relative change in output per FT worker** across all sectors.


#2015 to 2023, linked on the regional GVA bespoke SIC categories
#Note, can use lookup if need to also sum for sections (which we can do here cos current prices)

# gvabres <- readRDS('../data/regionalGVA_plus_BRESjobcounts/regionalGVA_currentprices_BRES_FT_jobcount_bespoke2digitSIC_nONLY_MATCHING_GEOGs_2015_2023.rds') 
#gvabres <- readRDS('data/regionalGVA_plus_BRESjobcounts/regionalGVA_currentprices_BRES_FT_jobcount_bespoke2digitSIC_nONLY_MATCHING_GEOGs_2015_2023.rds') 

gvabres <- readRDS('../data/regionalGVA_plus_BRESjobcounts/regionalGVA_chainedvolume_BRES_FT_jobcount_bespoke2digitSIC_nONLY_MATCHING_GEOGs_2015_2023.rds') 
#gvabres <- readRDS('data/regionalGVA_plus_BRESjobcounts/regionalGVA_chainedvolume_BRES_FT_jobcount_bespoke2digitSIC_nONLY_MATCHING_GEOGs_2015_2023.rds') 

#Make abbreviated sector names
shortsectornames = tibble(
  SIC07_description = unique(gvabres$SIC07_description),
  SIC07_description_shortened = c(
  "Furniture",
  "Machinery repair/install",
  "Buildings",
  "Civil eng",
  "Spec constr",
  "Motor",
  "Wholesale",
  "Retail",
  "Transport",
  "Warehousing",
  "Post",
  "Accommodation",
  "Food",
  "Media",
  "ICT",
  "Real estate",
  "Legal",
  "Head office/consult",
  "Arch/Eng",
  "Research/Ads",#This is a horrible combo that makes no sense! 72/73
  "Other PSTech",#"Prof/sci/tech",
  "Rental",
  "Employment",
  "Cleaning/landscapes",
  "Admin",
  "Health",
  "Care",
  "SocialWork",
  "Arts/ent",
  "Recreation",
  "Membership",
  "Repair personal/computer",
  "Personal",
  "Agri/mining",
  "Food manuf",
  "Textiles",
  "Wood/paper",
  "Chemicals",
  "Metals",
  "Electronics",
  "Mach/transp manuf",
  "Utilities",
  "Finance",
  "Public admin",
  "Education",
  "Households"
  )
)

#save that for use elsewhere!
# write_csv(shortsectornames, 'data/shortsectornames_for_regionalGVA_2digitSICs.csv')

gvabres <- gvabres %>% 
  left_join(
    shortsectornames, by = 'SIC07_description'
  )%>% 
  filter(!qg('households|membership', SIC07_description_shortened))

#Look at version including codes used in GVA
#Ah not a good source as it's filtered out <100 FTs (so e.g. missing sci, which is only 50)
# gvabres %>% select(SIC_2DIGIT_CODE_GVA2023,SIC07_description,SIC07_description_shortened) %>% distinct() %>% View

#Would also quite like to split manufacturing off
#Check in orig GVA file for the codes...
#This is the full list - some won't match but enough to label the correct ones at 2 digit for those that do
productionsectors <- c(
  'A-E',
  'AB (1-9)',
  'C (10-33)',
  'CA (10-12)',
  'CB (13-15)',
  'CC (16-18)',
  'CD-CG (19-23)',
  'CH (24-25)',
  'CI-CJ (26-27)',
  'CK-CL (28-30)',
  'CM (31-33)',
  '31-32',
  '33',
  'DE (35-39)',
  'F (41-43)',
  '41',
  '42',
  '43'
)

#Yep
# table(productionsectors %in% gvabres$SIC_2DIGIT_CODE_GVA2023)
# unique(gvabres$SIC_2DIGIT_CODE_GVA2023)[unique(gvabres$SIC_2DIGIT_CODE_GVA2023) %in% productionsectors]

#Label production sectors
gvabres <- gvabres %>% 
  mutate(
    productionsector = ifelse(
      SIC_2DIGIT_CODE_GVA2023 %in% productionsectors,
      'production','other'
    )
  )



#For current prices - what we can do:
#Find GVA as a proportion of total UK output for those years
#Then that proportion per FT job
#To see how productivity has shifted about internally / how sectors have changed relative to each other

#That version is missing some places so isn't entire UK
#Let's get entire UK's GVA per year
#Can just sum from any current price source
#2023:
allgva <- read_csv('../data/regionalGVA/regionalGVA_currentprices_ITL2_SIC_2DIGIT_LONG_2023.csv')
#allgva <- read_csv('data/regionalGVA/regionalGVA_currentprices_ITL2_SIC_2DIGIT_LONG_2023.csv')




allgva <- allgva %>% 
  group_by(year) %>% 
  summarise(totalUKGVA_CP = sum(value)) %>% 
  ungroup()

#Add that in...
gvabres <- gvabres %>% 
  left_join(allgva, by = c('DATE'= 'year')
  )

#Some NAs in jobs - I think the issue with the missing values is just early years with no agri data...
#filter
gvabres <- gvabres %>% filter(!is.na(JOBCOUNT))


#Find gva proportion of UK total then the job's proportion of that
gvabres <- gvabres %>%
  mutate(
    gva_proportionofUKtotal = GVA / totalUKGVA_CP,
    job_output_propofUKtotal = gva_proportionofUKtotal / JOBCOUNT,
    gvaperjob = GVA  / JOBCOUNT#assuming using chained volume
  )


#Get regional GVA data version of SIC sections 
#To plot each by
# siclookupfromreg.gva <- read_csv('../data/SIClookup_forBRES_REGIONALGVA_2023_JOIN.csv')
#siclookupfromreg.gva <- read_csv('data/SIClookup_forBRES_REGIONALGVA_2023_JOIN.csv')



smoothband = 3

gvabres <- gvabres %>%
  group_by(Region_name,SIC07_description) %>%
  mutate(
    jobcount_movingav = rollapply(JOBCOUNT,smoothband,mean,align='center',fill=NA),
    gva_movingav = rollapply(GVA,smoothband,mean,align='center',fill=NA),
    gvaperjob_movingav = rollapply(gvaperjob * 1000,smoothband,mean,align='center',fill=NA),
    job_output_propofUKtotal_movingav = rollapply(job_output_propofUKtotal,smoothband,mean,align='center',fill=NA),
    job_PPM_ofUKtotal_movingav = job_output_propofUKtotal_movingav * 1000000#1% is 1 part per 100. Which we find by multiplying up the prop by 100. PPM is just x * a million.
  ) %>% 
  ungroup()



#ggplot won't give control over order very well... let's add on top manually, more control

#CURRENT PRICES VERSION
# prod.data <- gvabres %>% 
#   mutate(
#     DATE = DATE - 2000,#Make dates 2 digit, more readable on axis
#     SIC07_description_shortened = fct_reorder(SIC07_description_shortened, job_PPM_ofUKtotal_movingav, .desc = T),
#     highlow = ifelse(SIC07_description_shortened %in% levels(SIC07_description_shortened)[1:23],'high','low'),#Break into two plots to stack, split by total av productivity
#     sectorgrouping = cut_number(as.integer(SIC07_description_shortened), 8) %>% as.integer
#   ) %>% 
#   # filter(!qg('hous',SIC07_description))
#   filter(!qg('hous',SIC07_description), !is.na(job_output_propofUKtotal_movingav))


#CHAINED VOL VERSION
prod.data <- gvabres %>% 
  mutate(
    DATE = DATE - 2000,#Make dates 2 digit, more readable on axis
    SIC07_description_shortened = fct_reorder(SIC07_description_shortened, gvaperjob_movingav, .desc = T),
    highlow = ifelse(SIC07_description_shortened %in% levels(SIC07_description_shortened)[1:23],'high','low'),#Break into two plots to stack, split by total av productivity
    sectorgrouping = cut_number(as.integer(SIC07_description_shortened), 8) %>% as.integer
  ) %>% 
  # filter(!qg('hous',SIC07_description))
  filter(!qg('hous',SIC07_description), !is.na(gvaperjob_movingav))

#Break into n groups, using factor order created in previous line
# sectorgrouping = ntile(as.integer(prod.data$SIC07_description_shortened), 5)

#Save this for use elsewhere...
# saveRDS(prod.data,'local/data/proddata.rds')


#table(prod.data$sectorgroupings)

#Test
# debugonce(plotprod)
# plotprod(prod.data %>% filter(sectorgrouping==6),job_PPM_ofUKtotal_movingav)#current prices
# plotprod(prod.data %>% filter(sectorgrouping==6),gvaperjob_movingav)#chained vol

# plotz <- map(prod.data %>% group_split(sectorgrouping),plotprod,job_PPM_ofUKtotal_movingav)
plotz <- map(prod.data %>% group_split(sectorgrouping),plotprod,gvaperjob_movingav)

patchwork::wrap_plots(plotz,ncol = 1)

#Quick look
#gvabres %>% filter(qg('architecture',SIC07_description_shortened), qg('bradford',Region_name)) %>% View


```



## Location quotients for GVA and jobs in 2 digit SIC sectors: most recent and change over time

These LQ plots show location quotients for linked GVA and jobs data for Bradford, overlaid on top of LQs for all other ITL3 zones, to give a sense of what the spread is across Great Britain.

A location quotient is the ratio of a place's sector proportion to its national proportion. So e.g. if manufacturing in Bradford made up 20% of its economy compared to maufacturing being 10% of the UK economy, its LQ in Bradford would be 2.

Here’s an explanation of how to read these location quotient plots:

* Location quotients for sectoral GVA in the selected region
* Sectors to the right of the blue line are more concentrated in this region when compared to the national average. (e.g. LQ = 10, it is 10x more concentrated). To the left are less concentrated than nationally.
* Green circles have been increasing in concentration over time (larger circles increasing the most)
* Red circles have been decreasing in concentration over time (larger circles decreasing the most)
* Black horizontal lines show the min and max of concentration in the data’s history (from 2015 to 2023) (so, often, large green circles will be at the right of this line because sector concentration has been steadily increasing)
* Per sector GVA money amounts and percent of that region’s entire economy are shown on the right.

### Production sectors


```{r, fig-productionLQs}
#| fig-width: 9
#| fig-height: 10

#Get two versions of LQ/proportions - job count and GVA
gva.props <- gvabres %>% 
  group_split(DATE) %>%
  map(add_location_quotient_and_proportions,
      regionvar = Region_name,
      lq_var = SIC07_description_shortened,
      valuevar = GVA) %>% 
  bind_rows()

job.props <- gvabres %>% 
  group_split(DATE) %>%
  map(add_location_quotient_and_proportions,
      regionvar = Region_name,
      lq_var = SIC07_description_shortened,
      valuevar = JOBCOUNT) %>% 
  bind_rows()

# LET'S SEPARATE GVA AND JOBS OUT FOR BRADFORD AND COMPARE TO EVERYWHERE ELSE TOO
place = "Bradford"

#FUNCTION
#Split both gva.props and job.props by subsector list
makegvajoblqplots <- function(keepthesesectors){
  
  # gva.sub <- gva.props %>% filter(SIC07_description_shortened %in% keepthesesectors)

  LQ_slopes <- compute_slope_or_zero(
    data = gva.props, 
    Region_name, SIC07_description_shortened,#slopes will be found within whatever grouping vars are added here
    y = LQ_log, x = DATE)
  
  
  #Filter down to a single year... we may want to smooth years, let's see
  yeartoplot <- gva.props %>% filter(DATE == max(DATE))#use latest year
  
  #Add slopes into data to get LQ plots
  yeartoplot <- yeartoplot %>% 
    left_join(
      LQ_slopes,
      by = c('Region_name', 'SIC07_description_shortened')
    )
  
  #Get min/max values for LQ over time as well, for each sector and place, to add as bars so range of sector is easy to see
  minmaxes <- gva.props %>% 
    group_by(Region_name, SIC07_description_shortened) %>% 
    summarise(
      min_LQ_all_time = min(LQ),
      max_LQ_all_time = max(LQ)
    )
  
  #Join min and max
  yeartoplot <- yeartoplot %>% 
    left_join(
      minmaxes,
      by = c('Region_name', 'SIC07_description_shortened')
    )
  
  
  #Shorten here to subset of sectors
  yeartoplot <- yeartoplot %>% 
    filter(SIC07_description_shortened %in% keepthesesectors)
  
  
  sectorLQorder <- gva.props %>% filter(
    Region_name == place,
    DATE == max(DATE)#use latest data
  ) %>% 
    arrange(-LQ) %>% 
    select(SIC07_description_shortened) %>% 
    pull()
  
  
  #Turn the sector column into a factor and order by LCR's LQs
  yeartoplot$SIC07_description_shortened <- factor(yeartoplot$SIC07_description_shortened, levels = sectorLQorder, ordered = T)
  
  #If I could plot both and space them out, that would be good (could get Bradford change showing too)
  p <- LQ_baseplot(df = yeartoplot, alpha = 0.1, shape = 0, sector_name = SIC07_description_shortened, 
                   LQ_column = LQ, change_over_time = slope)
  
  p <- addplacename_to_LQplot(df = yeartoplot, plot_to_addto = p, 
                              placename = place, shapenumber = 16,
                              min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                              value_column = GVA, sector_regional_proportion = sector_regional_proportion,
                              region_name = Region_name,
                              sector_name = SIC07_description_shortened, change_over_time = slope, LQ_column = LQ,
                              text = 7)
  
  p <- p + 
    # coord_cartesian(xlim = c(0.1,7)) +
    ggtitle("GVA")
  
  
  
  
  # job.sub <- job.props %>% filter(SIC07_description_shortened %in% keepthesesectors)
  
  # REPEAT FOR JOBS / BRADFORD (Using same factor order?)
  LQ_slopes <- compute_slope_or_zero(
    data = job.props, 
    Region_name, SIC07_description_shortened,#slopes will be found within whatever grouping vars are added here
    y = LQ_log, x = DATE)
  
  
  #Filter down to a single year... we may want to smooth years, let's see
  yeartoplot <- job.props %>% filter(DATE == max(DATE))#use latest year
  
  #Add slopes into data to get LQ plots
  yeartoplot <- yeartoplot %>% 
    left_join(
      LQ_slopes,
      by = c('Region_name', 'SIC07_description_shortened')
    )
  
  #Get min/max values for LQ over time as well, for each sector and place, to add as bars so range of sector is easy to see
  minmaxes <- job.props %>% 
    group_by(Region_name, SIC07_description_shortened) %>% 
    summarise(
      min_LQ_all_time = min(LQ),
      max_LQ_all_time = max(LQ)
    )
  
  #Join min and max
  yeartoplot <- yeartoplot %>% 
    left_join(
      minmaxes,
      by = c('Region_name', 'SIC07_description_shortened')
    )
  
  #Shorten here to subset of sectors
  yeartoplot <- yeartoplot %>% 
    filter(SIC07_description_shortened %in% keepthesesectors)
  
  
  
  #USE SECTOR ORDER FROM GVA LQS ABOVE
  yeartoplot$SIC07_description_shortened <- factor(yeartoplot$SIC07_description_shortened, levels = sectorLQorder, ordered = T)
  
  #If I could plot both and space them out, that would be good (could get Bradford change showing too)
  p2 <- LQ_baseplot(df = yeartoplot, alpha = 0.1, shape = 0, sector_name = SIC07_description_shortened, 
                   LQ_column = LQ, change_over_time = slope)
  
  p2 <- addplacename_to_LQplot(df = yeartoplot, plot_to_addto = p2, 
                              placename = place, shapenumber = 16,
                              min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                              value_column = JOBCOUNT, sector_regional_proportion = sector_regional_proportion,
                              region_name = Region_name,
                              sector_name = SIC07_description_shortened, change_over_time = slope, LQ_column = LQ,
                              text = 7, value_col_ismoney = F)
  
  p2 <- p2 + 
    # coord_cartesian(xlim = c(0.1,7)) +
    ggtitle("Jobs")
  
  # p2 <- p2 + scale_y_discrete(position = "right")
  p2 <- p2 +
    theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
  
  p + p2 

}

#production sector plot first
makegvajoblqplots(gva.props$SIC07_description_shortened[gva.props$productionsector == 'production'])


```


### Rest of sectors

```{r, fig-otherLQs}
#| fig-width: 9
#| fig-height: 14

makegvajoblqplots(gva.props$SIC07_description_shortened[gva.props$productionsector == 'other'])

```




## Av. percent change over time for gva vs hours worked, ITL3 zones

**Hover for place names**.

* West Yorkshire ITL3s in green - Bradford's outlined in black (95% CIs in lines).
* Core cities in blue  (95% CIs)
* London ITL3s in red/pink

And:

* Anything above/left of diagonal line is a **productivity increase**. 




### 2004 to 2013 av % change

```{r, fig-plot1}
# startyear = 2014
# endyear = 2023

#So e.g. in the plot below for 2014-2023, Rotherham's productivity increased with a **GVA increase** and a **number of hours worked *decrease***.

startyear = 2004
endyear = 2013

#Hours per week first
slopes.hours = get_slope_and_se_safely(
  hoursworked %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(hoursworkedperweek), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.hours = slopes.hours %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )


#Mark 4 SY LAs and plot
slopes.hours = slopes.hours %>% 
  ungroup() %>% 
  # mutate(
  #   SY_LAs = ifelse(qg('rotherh|doncaste|sheffie|barnsley', Region_name),'SY LA','Other'),
  #   corecities = ifelse(Region_name %in% corecities,'core city','Other'),
  #   london = ifelse(qg('london',Region_name), 'London','Other')
  # )
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'SY LA',#will overwrite sheffield as core city
      qg('bradford|kirklees|calderdale|leeds|wakefield', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
      ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','WY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.hours = slopes.hours %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )




#Then GVA CV
slopes.gva = get_slope_and_se_safely(
  itl3.all %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(gva_cv), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.gva = slopes.gva %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )



#Mark 4 WY LAs and plot
slopes.gva = slopes.gva %>% 
  ungroup() %>% 
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      qg('bradford|kirklees|calderdale|leeds|wakefield', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
    ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','WY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.gva = slopes.gva %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )


# table(slopes.hours$itl3label)
# table(slopes.gva$itl3label)


#Keep the bits we want and combine into single df
both <- slopes.gva %>% 
  select(Region_name,itl3label,
         slope_percentperyear_GVA = slope_percentperyear,
         ci95min_percentperyear_GVA = ci95min_percentperyear,
         ci95max_percentperyear_GVA = ci95max_percentperyear) %>% 
  inner_join(
    slopes.hours %>% 
      select(Region_name,
             slope_percentperyear_HOURS = slope_percentperyear,
             ci95min_percentperyear_HOURS = ci95min_percentperyear,
             ci95max_percentperyear_HOURS = ci95max_percentperyear),
    by = 'Region_name'
  )


#Poor control over visibility. Layer instead. Base plot then function.
addtogva_hours_plot = function(p,plotdata, showerrors = T, sizeval = 1, colourval = 'black', alphaval = 1){
  
  p = p + geom_point(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, group = Region_name), colour = colourval, size = sizeval, alpha = alphaval) 
  
  if(showerrors){
    
    p = p + geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, xmin = ci95min_percentperyear_HOURS, xmax = ci95max_percentperyear_HOURS),width = 0.1, size = 1, colour = colourval, alpha = alphaval) +
    geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, ymin = ci95min_percentperyear_GVA, ymax = ci95max_percentperyear_GVA),width = 0.1, size = 1, colour = colourval, alpha = alphaval) 
    
  }
  
    p  + geom_abline(slope = 1, intercept = 0, colour = 'black', alpha = 0.5, size = 2) +
      geom_hline(yintercept = 0, colour = 'black', alpha = 0.3, size = 1) +
      geom_vline(xintercept = 0, colour = 'black', alpha = 0.3, size = 1)
  
}




p = ggplot()

p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Other'), F, 1, 'black', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'London'), F, 2, 'red', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Core city'), T, 3, RColorBrewer::brewer.pal(3, "Set1")[2], 0.5)
p = addtogva_hours_plot(p, both %>% filter(Region_name == 'Bradford'), F, 10, 'black', 1)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'WY LA'), T, 7, RColorBrewer::brewer.pal(3, "Set1")[3], 1)

p = p + coord_fixed() +
  xlab("Average % change: hours worked per week") +
  ylab("Average % change: GVA") 


ggplotly(p, tooltip = 'Region_name')



```

### 2014 to 2023 av % change

```{r, fig-plot2}
startyear = 2014
endyear = 2023

# startyear = 2004
# endyear = 2013

#Hours per week first
slopes.hours = get_slope_and_se_safely(
  hoursworked %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(hoursworkedperweek), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.hours = slopes.hours %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )


#Mark 4 WY LAs and plot
slopes.hours = slopes.hours %>% 
  ungroup() %>% 
  # mutate(
  #   SY_LAs = ifelse(qg('rotherh|doncaste|sheffie|barnsley', Region_name),'WY LA','Other'),
  #   corecities = ifelse(Region_name %in% corecities,'core city','Other'),
  #   london = ifelse(qg('london',Region_name), 'London','Other')
  # )
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      qg('bradford|kirklees|calderdale|leeds|wakefield', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
      ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','WY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.hours = slopes.hours %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )




#Then GVA CV
slopes.gva = get_slope_and_se_safely(
  itl3.all %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(gva_cv), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.gva = slopes.gva %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )



#Mark 4 WY LAs and plot
slopes.gva = slopes.gva %>% 
  ungroup() %>% 
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      qg('bradford|kirklees|calderdale|leeds|wakefield', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
    ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','WY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.gva = slopes.gva %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )


# table(slopes.hours$itl3label)
# table(slopes.gva$itl3label)


#Keep the bits we want and combine into single df
both <- slopes.gva %>% 
  select(Region_name,itl3label,
         slope_percentperyear_GVA = slope_percentperyear,
         ci95min_percentperyear_GVA = ci95min_percentperyear,
         ci95max_percentperyear_GVA = ci95max_percentperyear) %>% 
  inner_join(
    slopes.hours %>% 
      select(Region_name,
             slope_percentperyear_HOURS = slope_percentperyear,
             ci95min_percentperyear_HOURS = ci95min_percentperyear,
             ci95max_percentperyear_HOURS = ci95max_percentperyear),
    by = 'Region_name'
  )


#Poor control over visibility. Layer instead. Base plot then function.
addtogva_hours_plot = function(p,plotdata, showerrors = T, sizeval = 1, colourval = 'black', alphaval = 1){
  
  p = p + geom_point(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, group = Region_name), colour = colourval, size = sizeval, alpha = alphaval) 
  
  if(showerrors){
    
    p = p + geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, xmin = ci95min_percentperyear_HOURS, xmax = ci95max_percentperyear_HOURS),width = 0.1, size = 1, colour = colourval, alpha = alphaval) +
    geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, ymin = ci95min_percentperyear_GVA, ymax = ci95max_percentperyear_GVA),width = 0.1, size = 1, colour = colourval, alpha = alphaval) 
    
  }
  
    p  + geom_abline(slope = 1, intercept = 0, colour = 'black', alpha = 0.5, size = 2) +
      geom_hline(yintercept = 0, colour = 'black', alpha = 0.3, size = 1) +
      geom_vline(xintercept = 0, colour = 'black', alpha = 0.3, size = 1)
  
}




p = ggplot()

p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Other'), F, 1, 'black', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'London'), F, 2, 'red', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Core city'), T, 3, RColorBrewer::brewer.pal(3, "Set1")[2], 0.5)
p = addtogva_hours_plot(p, both %>% filter(Region_name == 'Bradford'), F, 10, 'black', 1)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'WY LA'), T, 7, RColorBrewer::brewer.pal(3, "Set1")[3], 1)

p = p + coord_fixed() +
  xlab("Average % change: hours worked per week") +
  ylab("Average % change: GVA") 


ggplotly(p, tooltip = 'Region_name')



```









# COMPANIES HOUSE DATA


```{r, fig-chplot}

#See here for fig-responsiveness (non)issue: https://github.com/quarto-dev/quarto-cli/issues/4836

#30 seconds-ish
# x <- Sys.time()
ch = readRDS('../../companieshouseopen/local/PROCESSED_accountextracts_n_livelist_geocoded_combined_July2025.rds')
# ch = readRDS('../companieshouseopen/local/PROCESSED_accountextracts_n_livelist_geocoded_combined_July2025.rds')
# Sys.time() - x

#get short SIC names
shorts <- readRDS('../local/data/bradford_SICs_withshortnames.rds')
#shorts <- readRDS('local/data/bradford_SICs_withshortnames.rds')

ch = ch %>%
  left_join(
    shorts %>% select(SIC_5DIGIT_CODE,SIC_SECTION_NAME_SHORT:SIC_5DIGIT_NAME_SHORT),
    by = 'SIC_5DIGIT_CODE'
  )

#Create the base summary, save, reuse
#Taken from CH code here
#https://github.com/DanOlner/companieshouseopen/blob/2ddc26d4b89716b8755c02fc6626190ad624b888/testcode/initial_datadigging.R#L417
empchange.summary <- ch %>% 
  st_set_geometry(NULL) %>% 
  filter(Employees_lastyear >= 10) %>%#Need to load cos we need to filter here so we can accurately count employees in 10+ employee firms
  group_by(localauthority_name,SIC_2DIGIT_NAME_SHORT) %>%
  summarise(
    total_employmentlastyear = sum(Employees_lastyear, na.rm = T),
    total_employmentthisyear = sum(Employees_thisyear, na.rm = T),
    employment_percentchange = percent_change(total_employmentlastyear,total_employmentthisyear),
    firmcount = n()
  ) %>% 
  ungroup() %>% 
  filter(firmcount > 5)#Reduce to sectors/places with 10+ firms
# 
# saveRDS(empchange.summary,'local/data/empchangsummary.rds')
# empchange.summary <- readRDS('../local/data/empchangsummary.rds')
#empchange.summary <- readRDS('local/data/empchangsummary.rds')

# empchange10plus <- empchange.summary %>% 
#   filter(
#     firmcount > 5,
#     Employees_lastyear >= 10#Reduce to sectors/places with 10+ firms
#   )

#Order by percent change per sector in Bradford
orderby <- empchange.summary %>% 
  filter(qg('bradford',localauthority_name)) %>%
  select(SIC_2DIGIT_NAME_SHORT,employment_percentchange) %>% 
  arrange(employment_percentchange) %>% 
  pull(SIC_2DIGIT_NAME_SHORT)

#Order by those...
empchange.summary <- empchange.summary %>% 
  mutate(
    SIC_2DIGIT_NAME_SHORT = factor(SIC_2DIGIT_NAME_SHORT, levels = orderby),
    `Place: ` = paste0(localauthority_name, ', firm count: ', firmcount,', emp last/this yr: ', total_employmentlastyear,' > ',total_employmentthisyear,' = ',round(employment_percentchange,2),'%')#for labels
    # `Place: ` = paste0(SIC_2DIGIT_NAME_SHORT,',', localauthority_name, ', firm count: ', firmcount,', emp last/this yr: ', total_employmentlastyear,' > ',total_employmentthisyear)#for labels
  ) %>% 
  filter(!is.na(SIC_2DIGIT_NAME_SHORT))#Filter these here as some will be missing spefically from Bradford, will end up NA in the factor

# table(is.na(empchange.summary$SIC_2DIGIT_NAME_SHORT))

#Get values for Bradford firms - count/%/employee counts
firmcount <- ch %>% st_set_geometry(NULL) %>% filter(localauthority_name == 'Bradford', Employees_thisyear > 0)

firmcount <- firmcount %>% 
  mutate(
    sizecategory = case_when(
      Employees_thisyear == 1 ~ "1",
      between(Employees_thisyear,2,4) ~ "2-4",
      between(Employees_thisyear,5,9) ~ "5-9",
      between(Employees_thisyear,10,20) ~ "10-20",
      between(Employees_thisyear,21,50) ~ "21-50",
      between(Employees_thisyear,51,999999) ~ "51+"
    ),
    sizecategory = factor(sizecategory, levels = c('1','2-4','5-9','10-20','21-50','51+'))
    )

#table from that to plot
firmtable <- tibble(
  `Firm size` = levels(firmcount$sizecategory),
  `Count` = table(firmcount$sizecategory),
  `Percent of firms` = paste0(round(table(firmcount$sizecategory) %>% prop.table() * 100,2),"%")
)

```


```{r eval=F}
firmcounts.by.2digit <- firmcount %>% 
  group_by(sizecategory,SIC_2DIGIT_NAME_SHORT) %>% 
  summarise(
    count = n(),
    section = max(SIC_SECTION_NAME_SHORT)
    ) %>% 
  group_by(SIC_2DIGIT_NAME_SHORT) %>% 
  mutate(
    percent = (count / sum(count)) * 100
  )

ggplot(firmcounts.by.2digit, aes(y = SIC_2DIGIT_NAME_SHORT, x = percent, fill = sizecategory)) +
  geom_bar(stat = 'identity', position = position_dodge(width = 0.8)) +
  facet_wrap(~section,ncol = 2, scales = 'free')

```



The next few plots use data extracted from individual firms' accounts **filed with Companies House between June 2024 and end of June 2025**. 

The following table shows the count (and percentage) of firms in Bradford that filed with CH, broken down by employee count.


```{r bradch_summarytable}

#escape = FALSE so that markdown in the cells (the links) is rendered correctly.
knitr::kable(firmtable, escape = FALSE, caption = "Table: Bradford count/% of firms by employee band")

```



## Comparing Companies House to BRES: location quotient plots

Note: these plots use the **full 2 digit SIC list (85 sectors included here, compared to 44 in the regional GVA groupings)**. This means that some sectors are split into more than one - e.g. finance (SIC 64 to 66 in the GVA data) is split here into **three**: finance, finance auxiliaries and insurance.

### Production sectors

```{r, fig-CHvBRESprod}
#| fig-width: 9

#Produced here:
#https://github.com/DanOlner/RegionalEconomicTools/blob/927f1783bfbb66293dd75870e147c858f4761ef3/bits_of_code/BradfordExplore.R#L1062
plotz <- readRDS('../local/data/ch_bres_lqplotlist.rds')
#plotz <- readRDS('local/data/ch_bres_lqplotlist.rds')

plotz[[1]] + plotz[[3]]

```

### Other sectors

```{r, fig-CHvBRESother}
#| fig-width: 9

plotz[[2]] + plotz[[4]]

```




## From Companies House firm level data: in the last year (March 2024-March 2025), percent change in employment in firms, five West Yorkshire local authorities compared to 'core cities' (hover for place names)

### Firms with 10+ employees


```{r, fig-percentempfourplaces}
#| fig-height: 20
#| label: fig-percentempfourplaces
#| fig-cap: "Companies House firm level data: percent change in employment for firms with ten+ employees between most recent accounts and previous year - core cities, with West Yorkshire's five local authorities overlaid. Hover for place name."


p <- ggplot() +
  geom_point(
    position = position_nudge(y = 0.2),
    data = empchange.summary %>% filter(
      qg('Belfast|Birmingham|Bristol|Cardiff|Glasgow|Leeds|Liverpool|Manchester|Tyne|Sheffield|Nottingham', localauthority_name)#just core cities (minus four places below)
    ),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, group = `Place: `),
    alpha = 0.5)

#Add four places
p <- p + 
  geom_point(
    position = position_nudge(y = -0.2),
    data = empchange.summary %>% filter(qg('leeds|kirklees|wakefield|calderdale',localauthority_name)),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, colour = localauthority_name, group = `Place: `),
    size = 7, shape = 17) +
  scale_colour_brewer(palette = 'Paired', direction = 1, guide = guide_legend(title = NULL)) +
  # coord_cartesian(xlim = c(-25,25)) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") 

#Add Bradford over the top to make sure visible
p <- p + 
  geom_point(
    position = position_nudge(y = -0.2),
    data = empchange.summary %>% filter(qg('bradford',localauthority_name)),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, colour = localauthority_name, group = `Place: `),
    size = 8, shape = 17) +
  scale_colour_brewer(palette = 'Paired', direction = 1, guide = guide_legend(title = NULL)) +
  # coord_cartesian(xlim = c(-25,25)) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") 



#p

# ggplotly(p, width = 800, height = 900, tooltip = 'localauthority_name')
ggplotly(p, tooltip = 'Place: ')


```

###  Firms with fewer than 10 employees



```{r, fig-percentempfourplaces_lessthan10}
#| fig-height: 20
#| label: fig-percentempfourplaces_lessthan10
#| fig-cap: "From Companies House firm level data: percent change in employment for firms with fewer than 10 employees, between most recent accounts and previous year - core cities, with West Yorkshire's five local authorities overlaid. Hover for place name."

empchange.summary <- ch %>% 
  st_set_geometry(NULL) %>% 
  filter(Employees_lastyear < 10) %>%#Need to load cos we need to filter here so we can accurately count employees in 10+ employee firms
  group_by(localauthority_name,SIC_2DIGIT_NAME_SHORT) %>%
  summarise(
    total_employmentlastyear = sum(Employees_lastyear, na.rm = T),
    total_employmentthisyear = sum(Employees_thisyear, na.rm = T),
    employment_percentchange = percent_change(total_employmentlastyear,total_employmentthisyear),
    firmcount = n()
  ) %>% 
  ungroup() %>% 
  filter(firmcount > 5)#Reduce to sectors/places with 10+ firms
# 
# saveRDS(empchange.summary,'local/data/empchangsummary.rds')
# empchange.summary <- readRDS('../local/data/empchangsummary.rds')
#empchange.summary <- readRDS('local/data/empchangsummary.rds')

# empchange10plus <- empchange.summary %>% 
#   filter(
#     firmcount > 5,
#     Employees_lastyear >= 10#Reduce to sectors/places with 10+ firms
#   )

#Order by percent change per sector in Bradford
orderby <- empchange.summary %>% 
  filter(qg('bradford',localauthority_name)) %>%
  select(SIC_2DIGIT_NAME_SHORT,employment_percentchange) %>% 
  arrange(employment_percentchange) %>% 
  pull(SIC_2DIGIT_NAME_SHORT)

#Order by those...
empchange.summary <- empchange.summary %>% 
  mutate(
    SIC_2DIGIT_NAME_SHORT = factor(SIC_2DIGIT_NAME_SHORT, levels = orderby),
    `Place: ` = paste0(localauthority_name, ', firm count: ', firmcount,', emp last/this yr: ', total_employmentlastyear,' > ',total_employmentthisyear,' = ',round(employment_percentchange,2),'%')#for labels
    # `Place: ` = paste0(SIC_2DIGIT_NAME_SHORT,',', localauthority_name, ', firm count: ', firmcount,', emp last/this yr: ', total_employmentlastyear,' > ',total_employmentthisyear)#for labels
  ) %>% 
  filter(!is.na(SIC_2DIGIT_NAME_SHORT))#Filter these here as some will be missing spefically from Bradford, will end up NA in the factor

# table(is.na(empchange.summary$SIC_2DIGIT_NAME_SHORT))




p <- ggplot() +
  geom_point(
    position = position_nudge(y = 0.2),
    data = empchange.summary %>% filter(
      qg('Belfast|Birmingham|Bristol|Cardiff|Glasgow|Leeds|Liverpool|Manchester|Tyne|Sheffield|Nottingham', localauthority_name)#just core cities (minus four places below)
    ),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, group = `Place: `),
    alpha = 0.5)

#Add four places
p <- p + 
  geom_point(
    position = position_nudge(y = -0.2),
    data = empchange.summary %>% filter(qg('leeds|kirklees|wakefield|calderdale',localauthority_name)),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, colour = localauthority_name, group = `Place: `),
    size = 7, shape = 17) +
  scale_colour_brewer(palette = 'Paired', direction = 1, guide = guide_legend(title = NULL)) +
  # coord_cartesian(xlim = c(-25,25)) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") 

#Add Bradford over the top to make sure visible
p <- p + 
  geom_point(
    position = position_nudge(y = -0.2),
    data = empchange.summary %>% filter(qg('bradford',localauthority_name)),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, colour = localauthority_name, group = `Place: `),
    size = 8, shape = 17) +
  scale_colour_brewer(palette = 'Paired', direction = 1, guide = guide_legend(title = NULL)) +
  # coord_cartesian(xlim = c(-25,25)) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") 

p <- p + coord_cartesian(xlim = c(-40,80))

#p

# ggplotly(p, width = 800, height = 900, tooltip = 'localauthority_name')
ggplotly(p, tooltip = 'Place: ')


```




## Just for Bradford, individual firms with 10+ employees, percent employee change from last accounts (hover for firm names)

**Hover over points** to identify individual Bradford firms (with 10+ employees). Each is plotted by what percentage their employee number grew between their latest accounts and a year prior.

```{r, fig-indivfirmsSY}
#| fig-height: 10
#| fig-width: 9

# fig-cap: "Percent change in employment between most recent accounts and previous year, ALL Bradford firms with 10+ employees, hover for firm name."

firm.change <- ch %>% 
  filter(localauthority_name == 'Bradford', Employees_thisyear >= 10 & Employees_lastyear >= 10) %>% #10+ in both time periods
  mutate(
    employee_percentchange = percent_change(Employees_lastyear,Employees_thisyear)
  )

#As well as a minimum of ten, let's break down by firm size...
#What's the spread?
# ggplot(firm.change, aes(x = Employees_thisyear)) +
#   geom_density()

firm.change <- firm.change %>% 
  mutate(
    SIC_2DIGIT_NAME_SHORT = fct_reorder(SIC_2DIGIT_NAME_SHORT, employee_percentchange),
    `Firm: ` = paste0(CompanyName,', ',Employees_lastyear,' >> ',Employees_thisyear),
    production = ifelse(SIC_2DIGIT_CODE_NUMERIC < 44, 'production','other'),
    production = factor(production,levels = c('production','other')),
    sizecategory = case_when(
      between(Employees_thisyear,10,20) ~ "10-20",
      between(Employees_thisyear,21,50) ~ "21-50",
      between(Employees_thisyear,51,999999) ~ "51+"
    )
    ) %>% 
  filter(!is.na(SIC_2DIGIT_NAME_SHORT))

# table(firm.change$sizecategory)
 # table(firm.change$production)

#FACET doesn't work well with plotly - things overlap
#Plot them separately.
p <- ggplot(firm.change %>% filter(sizecategory == "10-20"),
            aes(y = SIC_2DIGIT_NAME_SHORT, x = employee_percentchange, group = `Firm: `)) +
  geom_jitter(height = 0.1, alpha = 0.3) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") +
  facet_wrap(~production, ncol = 2, scales = 'free') +
  theme(
    # panel.grid.major = element_line(linetype = "dotted"),
    # axis.title.x = element_text( size = 10, margin=margin(80,80,80,80)),
    axis.title.y = element_text( size = 10, margin=margin(30,30,30,30))
  ) +
  ggtitle("Firms with 10-20 employees")

ggplotly(p, tooltip = 'Firm: ')

p <- ggplot(firm.change %>% filter(sizecategory == "21-50"),
            aes(y = SIC_2DIGIT_NAME_SHORT, x = employee_percentchange, group = `Firm: `)) +
  geom_jitter(height = 0.1, alpha = 0.3) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") +
  facet_wrap(~production, ncol = 2, scales = 'free') +
  theme(
    # panel.grid.major = element_line(linetype = "dotted"),
    # axis.title.x = element_text( size = 10, margin=margin(80,80,80,80)),
    axis.title.y = element_text( size = 10, margin=margin(30,30,30,30))
  ) +
  ggtitle("Firms with 21-50 employees")


ggplotly(p, tooltip = 'Firm: ')

p <- ggplot(firm.change %>% filter(sizecategory == "51+"),
            aes(y = SIC_2DIGIT_NAME_SHORT, x = employee_percentchange, group = `Firm: `)) +
  geom_jitter(height = 0.1, alpha = 0.3) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") +
  facet_wrap(~production, ncol = 2, scales = 'free') +
  theme(
    # panel.grid.major = element_line(linetype = "dotted"),
    # axis.title.x = element_text( size = 10, margin=margin(80,80,80,80)),
    axis.title.y = element_text( size = 10, margin=margin(30,30,30,30))
  ) +
  ggtitle("Firms with 50+ employees")


ggplotly(p, tooltip = 'Firm: ')




```





# GROWTH TRENDS IN GVA, JOBS AND GVA PER JOB

This section examines **growth trends in Bradford** for **GVA and GVA per full time job (a productivity measure)** (both using the 46 sector categories available in the GVA data), as well as **growth in jobs using BRES data** (a larger list of 88 SIC 2 digit sectors).

**NOTE: There's a section right at the end that digs more deeply into some selected sectors from this section ("Closer look at jobs vs GVA percent change in selected sectors"), taken from the [GVA vs job percent change 2D plots](https://danolner.github.io/RegionalEconomicTools/quarto_docs/Bradford_2D_gvajobs_percentchangeplots_July2025.html){target="_blank"} catalogue page.**

First, this is done for **the full range of data - 2015 to 2023**. Second, it's **repeated for the more recent time period from 2019 to 2023**, to look for trends changes across the pandemic and more recently.

Here's an explanation of what's in the plots below. 

* Linear slopes are found for each sector in Bradford, and each is compared to (a) all other sectors in Bradford ("Bradford internal", light blue) and (b) the same sector in all other ITL3s zones in the data ("Bradford to other places", dark blue).
* For example - looking at the most obvious stand-out sector in the first plot is **head offices / consultancy**. Its GVA has grown significantly more than around 90% of other sectors in Bradford, and significantly more than around 70% of the head office / consultancy sector in other ITL3 zones. Its productivity (measured in GVA per full time job) has grown more than around 43% of other ITL3 zones.
* The x axis gives the **percent of slopes that are significantly different to others** (i.e. 95% confidence intervals don't overlap).
* The vertical axis lists all sectors in alphabetical order. These labels contain the following information:
    - Green text: slope rises over time - growth is on average going up
    - Red: slope drops over time - growth is on average going down
    - Bold green and red: the up or down slope is significantly different from zero (faded red and green can't be statitically separated from no change over time).
    - Text after the sector name is **percent change per year on average** with 95% confidence intervals.

So:

* To see **whether a sector has been growing or shrinking overall**, use the sector text (green for growth, red for shrinkage, values given for average % change per year)
* To **compare that sector's trend with other sectors in Bradford or the same sector in other places**, use the bars in the plot. Note - these are relative: a slope could be relatively more positive than another, but the sector overall could still be shrinking. For example, ICT has been growing strongly in most places in the UK - but is growth in one place significantly stronger than others? That's what the place and sector slope comparisons try to answer.


## Full data range: 2015 to 2023


```{r, fig-growthdata_cv_andperFT}
#| fig-height: 12 
#| fig-width: 9.5

#Counting sigs. Statisticians having kittens. Let's do it.
itl3.2digit.cv <- read_csv('../data/regionalGVA/regionalGVA_chainedvolume_ITL3_SIC_2DIGIT_LONG_2023.csv') %>% filter(!qg('agri|Owner-occupiers|membership|activities of households',SIC07_description))

#Add in shorter regional GVA SIC text
shortsectornames <- read_csv('../data/shortsectornames_for_regionalGVA_2digitSICs.csv')

#check match, tick
# table(unique(itl3.2digit.cv$SIC07_description) %in% shortsectornames$SIC07_description)
itl3.2digit.cv <- itl3.2digit.cv %>% 
  left_join(
    shortsectornames, by = 'SIC07_description'
  )

#CV first. No NeweyWest
bradford.2dig <- plotSlopeCounts(
  df = itl3.2digit.cv,
  placename = 'Bradford',
  startdate = 2014,
  enddate = 2023,
  date_colname = year,
  region_colname = Region_name,
  sector_colname = SIC07_description_shortened,
  value_colname = value,
  conf_interval = 95
)




#And for GVA per job - for the places we're able to combined data for these (done in miscchecks.R)
bres.gva.2digit.2023 <- readRDS('../data/regionalGVA_plus_BRESjobcounts/regionalGVA_chainedvolume_BRES_FT_jobcount_bespoke2digitSIC_nONLY_MATCHING_GEOGs_2015_2023.rds') %>% 
  filter(JOBCOUNT > 0, !qg('membership|activities of households|agri', SIC07_description)) %>% 
  mutate(gvaperjob = GVA / JOBCOUNT)

#check match, tick
# table(unique(bres.gva.2digit.2023$SIC07_description) %in% shortsectornames$SIC07_description)
bres.gva.2digit.2023 <- bres.gva.2digit.2023 %>% 
  left_join(
    shortsectornames, by = 'SIC07_description'
  )


#CHeck this has same list of sectors as the CV data above... tick
#(With imputed rent removed from CV)
# table(unique(bres.gva.2digit.2023$SIC07_description) %in% itl3.2digit.cv$SIC07_description)
# unique(bres.gva.2digit.2023$SIC07_description)[!unique(bres.gva.2digit.2023$SIC07_description) %in% itl3.2digit.cv$SIC07_description]

bradford.gvaperjob.2dig <- plotSlopeCounts(
  df = bres.gva.2digit.2023,
  placename = 'Bradford',
  startdate = 2015,
  enddate = 2023,
  date_colname = DATE,
  region_colname = Region_name,
  sector_colname = SIC07_description_shortened,
  value_colname = gvaperjob,#Note, this gets log'd in the function, don't do it here
  conf_interval = 95,
  includesectorname_on_axis = T
)

#Check sector match before alterations
# bradford.2dig$plot + bradford.gvaperjob.2dig$plot


#Axes match - remove from RHS
bradford.gvaperjob.2dig$plot <- bradford.gvaperjob.2dig$plot +
  theme(axis.title.y=element_blank(),
        # axis.text.y=element_blank(),
        axis.ticks.y=element_blank(), 
        axis.title.x=element_blank(),
        legend.position = "bottom"
        ) +
        # axis.text.x=element_blank(),
        # axis.ticks.x=element_blank()) +
  guides(fill = F) +
  ggtitle("GVA/FT") 

#And only need the one legend!
bradford.2dig$plot <- bradford.2dig$plot +
  ggtitle("GVA") +
  theme(
    legend.position = "bottom",
    axis.title.y=element_blank()#add this explanation in figure text
  )

bradford.2dig$plot + bradford.gvaperjob.2dig$plot


```




```{r, fig-growthdata_bres2dig}
#| fig-height: 14 

bres <- read_csv("../local/data/BRES/separate_SIC_types_summedfrom5digitSIC/BRES_ALLYEARSWITHDATA_NUTS3_n_ITL321_stacked_2_Fulltimeemployees_2015_2023_SIC_5DIGIT.csv")

#This comes with ALL NUTS and ITL3 data
#Which means 2022 is doubled up
#E.g. see here: 
#bres %>% filter(qg('bradford', GEOGRAPHY_NAME),qg('Museum activities',SIC_5DIGIT_NAME)) %>% View

#Remove one of them...
#Need to filter NOT year == 2022 NOT GEOGRAPHY_CODE_NUTS2018 is NA, keep the rest
#(2023 we need to keep!)
bres <- bres %>% 
  filter(!(DATE == 2022 & is.na(GEOGRAPHY_CODE_NUTS2018)))

#unique(bres$DATE)#tick

SIClookup <- read_csv('../data/SIClookup.csv')

#Join SIC lookup on 5 digit name
#Keep 2 digit and section codes
#May shorten names in a mo...
bres <- bres %>%
  left_join(
    SIClookup %>% select(SIC_5DIGIT_NAME,SIC_2DIGIT_NAME,SIC_SECTION_NAME),
    by = 'SIC_5DIGIT_NAME'
  )

#Make shorter names, use those.
#Make lookup so can be merged in.
names.sections = unique(bres$SIC_SECTION_NAME)
shortnames.sections = reduceSICnames(unique(bres$SIC_SECTION_NAME),'section')

chk.sections = data.frame(names = names.sections, shortnames = shortnames.sections)

names.2digit = unique(bres$SIC_2DIGIT_NAME)
shortnames.2digit = reduceSICnames(unique(bres$SIC_SECTION_NAME),'2 digit')

chk.2digit = data.frame(names = names.2digit, shortnames = shortnames.2digit)


#And 729 categories in the 5 digit...
names.5digit = unique(bres$SIC_5DIGIT_NAME)
#NOTE: THE ORDER HERE IS BASED ON THE ORDER IN bres
#We should make an actual lookup to make sure we don't lose that order
shortnames.5digit = reduceSICnames(unique(bres$SIC_5DIGIT_NAME),'5 digit')

chk.5digit = data.frame(names = names.5digit, shortnames = shortnames.5digit)


#Merge into bres
bres <- bres %>% 
  left_join(chk.sections %>% rename(SIC_SECTION_NAME_SHORT = shortnames), by = c('SIC_SECTION_NAME' = 'names')) %>% 
  left_join(chk.2digit %>% rename(SIC_2DIGIT_NAME_SHORT = shortnames), by = c('SIC_2DIGIT_NAME' = 'names')) %>% 
  left_join(chk.5digit %>% rename(SIC_5DIGIT_NAME_SHORT = shortnames), by = c('SIC_5DIGIT_NAME' = 'names')) 


#Drop some sectors
bres <- bres %>% 
  filter(!qg('activities of households|agri|membership', SIC_2DIGIT_NAME))

#Good!
# table(unique(bres$SIC_2DIGIT_NAME_SHORT) %in% ch.2digitsums$SIC_2DIGIT_NAME_SHORT)


# AND ON WITH LQS 
#OK so getting on with LQs and thinking about what those mean...
#The LQ boundaries won't differ because they're both for Bradford
#And the national sums will be matching quantities across both
bres.2dig <- bres %>% 
  group_by(DATE,GEOGRAPHY_NAME,SIC_2DIGIT_NAME_SHORT) %>% 
  summarise(jobcount = sum(JOBCOUNT)) %>% 
  ungroup()

bradford.bres.2dig <- plotSlopeCounts(
  df = bres.2dig,
  placename = 'Bradford',
  startdate = 2015,
  enddate = 2023,
  date_colname = DATE,
  region_colname = GEOGRAPHY_NAME,
  sector_colname = SIC_2DIGIT_NAME_SHORT,#add shortened version in
  value_colname = jobcount,#Note, this gets log'd in the function, don't do it here
  conf_interval = 95
)

#GVA and BRES
# bradford.2dig2$plot + bradford.bres.2dig2$plot
bradford.bres.2dig$plot <- bradford.bres.2dig$plot +
theme(
  axis.title.y=element_blank(),
  legend.position = "bottom") +
  ggtitle("BRES FT")

bradford.bres.2dig$plot

```





## More recent years: 2019 to 2023

```{r, fig-growthdata_cv_andperFT_later}
#| fig-height: 12 
#| fig-width: 9.5

#Counting sigs. Statisticians having kittens. Let's do it.
itl3.2digit.cv <- read_csv('../data/regionalGVA/regionalGVA_chainedvolume_ITL3_SIC_2DIGIT_LONG_2023.csv') %>% filter(!qg('agri|Owner-occupiers|membership|activities of households',SIC07_description))

#Add in shorter regional GVA SIC text
shortsectornames <- read_csv('../data/shortsectornames_for_regionalGVA_2digitSICs.csv')

#check match, tick
# table(unique(itl3.2digit.cv$SIC07_description) %in% shortsectornames$SIC07_description)
itl3.2digit.cv <- itl3.2digit.cv %>% 
  left_join(
    shortsectornames, by = 'SIC07_description'
  )

#CV first. No NeweyWest
bradford.2dig <- plotSlopeCounts(
  df = itl3.2digit.cv,
  placename = 'Bradford',
  startdate = 2019,
  enddate = 2023,
  date_colname = year,
  region_colname = Region_name,
  sector_colname = SIC07_description_shortened,
  value_colname = value,
  conf_interval = 95
)




#And for GVA per job - for the places we're able to combined data for these (done in miscchecks.R)
bres.gva.2digit.2023 <- readRDS('../data/regionalGVA_plus_BRESjobcounts/regionalGVA_chainedvolume_BRES_FT_jobcount_bespoke2digitSIC_nONLY_MATCHING_GEOGs_2015_2023.rds') %>% 
  filter(JOBCOUNT > 0, !qg('membership|activities of households|agri', SIC07_description)) %>% 
  mutate(gvaperjob = GVA / JOBCOUNT)

#check match, tick
# table(unique(bres.gva.2digit.2023$SIC07_description) %in% shortsectornames$SIC07_description)
bres.gva.2digit.2023 <- bres.gva.2digit.2023 %>% 
  left_join(
    shortsectornames, by = 'SIC07_description'
  )


#CHeck this has same list of sectors as the CV data above... tick
#(With imputed rent removed from CV)
# table(unique(bres.gva.2digit.2023$SIC07_description) %in% itl3.2digit.cv$SIC07_description)
# unique(bres.gva.2digit.2023$SIC07_description)[!unique(bres.gva.2digit.2023$SIC07_description) %in% itl3.2digit.cv$SIC07_description]

bradford.gvaperjob.2dig <- plotSlopeCounts(
  df = bres.gva.2digit.2023,
  placename = 'Bradford',
  startdate = 2019,
  enddate = 2023,
  date_colname = DATE,
  region_colname = Region_name,
  sector_colname = SIC07_description_shortened,
  value_colname = gvaperjob,#Note, this gets log'd in the function, don't do it here
  conf_interval = 95,
  includesectorname_on_axis = T
)

#Check sector match before alterations
# bradford.2dig$plot + bradford.gvaperjob.2dig$plot


#Axes match - remove from RHS
bradford.gvaperjob.2dig$plot <- bradford.gvaperjob.2dig$plot +
  theme(axis.title.y=element_blank(),
        # axis.text.y=element_blank(),
        axis.ticks.y=element_blank(), 
        axis.title.x=element_blank(),
        legend.position = "bottom"
        ) +
        # axis.text.x=element_blank(),
        # axis.ticks.x=element_blank()) +
  guides(fill = F) +
  ggtitle("GVA/FT") 

#And only need the one legend!
bradford.2dig$plot <- bradford.2dig$plot +
  ggtitle("GVA") +
  theme(
    legend.position = "bottom",
    axis.title.y=element_blank()#add this explanation in figure text
  )

bradford.2dig$plot + bradford.gvaperjob.2dig$plot


```




```{r, fig-growthdata_bres2dig_later}
#| fig-height: 14 

bres <- read_csv("../local/data/BRES/separate_SIC_types_summedfrom5digitSIC/BRES_ALLYEARSWITHDATA_NUTS3_n_ITL321_stacked_2_Fulltimeemployees_2015_2023_SIC_5DIGIT.csv")

#This comes with ALL NUTS and ITL3 data
#Which means 2022 is doubled up
#E.g. see here: 
#bres %>% filter(qg('bradford', GEOGRAPHY_NAME),qg('Museum activities',SIC_5DIGIT_NAME)) %>% View

#Remove one of them...
#Need to filter NOT year == 2022 NOT GEOGRAPHY_CODE_NUTS2018 is NA, keep the rest
#(2023 we need to keep!)
bres <- bres %>% 
  filter(!(DATE == 2022 & is.na(GEOGRAPHY_CODE_NUTS2018)))

#unique(bres$DATE)#tick

SIClookup <- read_csv('../data/SIClookup.csv')

#Join SIC lookup on 5 digit name
#Keep 2 digit and section codes
#May shorten names in a mo...
bres <- bres %>%
  left_join(
    SIClookup %>% select(SIC_5DIGIT_NAME,SIC_2DIGIT_NAME,SIC_SECTION_NAME),
    by = 'SIC_5DIGIT_NAME'
  )

#Make shorter names, use those.
#Make lookup so can be merged in.
names.sections = unique(bres$SIC_SECTION_NAME)
shortnames.sections = reduceSICnames(unique(bres$SIC_SECTION_NAME),'section')

chk.sections = data.frame(names = names.sections, shortnames = shortnames.sections)

names.2digit = unique(bres$SIC_2DIGIT_NAME)
shortnames.2digit = reduceSICnames(unique(bres$SIC_SECTION_NAME),'2 digit')

chk.2digit = data.frame(names = names.2digit, shortnames = shortnames.2digit)


#And 729 categories in the 5 digit...
names.5digit = unique(bres$SIC_5DIGIT_NAME)
#NOTE: THE ORDER HERE IS BASED ON THE ORDER IN bres
#We should make an actual lookup to make sure we don't lose that order
shortnames.5digit = reduceSICnames(unique(bres$SIC_5DIGIT_NAME),'5 digit')

chk.5digit = data.frame(names = names.5digit, shortnames = shortnames.5digit)


#Merge into bres
bres <- bres %>% 
  left_join(chk.sections %>% rename(SIC_SECTION_NAME_SHORT = shortnames), by = c('SIC_SECTION_NAME' = 'names')) %>% 
  left_join(chk.2digit %>% rename(SIC_2DIGIT_NAME_SHORT = shortnames), by = c('SIC_2DIGIT_NAME' = 'names')) %>% 
  left_join(chk.5digit %>% rename(SIC_5DIGIT_NAME_SHORT = shortnames), by = c('SIC_5DIGIT_NAME' = 'names')) 


#Drop some sectors
bres <- bres %>% 
  filter(!qg('activities of households|agri|membership', SIC_2DIGIT_NAME))

#Good!
# table(unique(bres$SIC_2DIGIT_NAME_SHORT) %in% ch.2digitsums$SIC_2DIGIT_NAME_SHORT)


# AND ON WITH LQS 
#OK so getting on with LQs and thinking about what those mean...
#The LQ boundaries won't differ because they're both for Bradford
#And the national sums will be matching quantities across both
bres.2dig <- bres %>% 
  group_by(DATE,GEOGRAPHY_NAME,SIC_2DIGIT_NAME_SHORT) %>% 
  summarise(jobcount = sum(JOBCOUNT)) %>% 
  ungroup()

bradford.bres.2dig <- plotSlopeCounts(
  df = bres.2dig,
  placename = 'Bradford',
  startdate = 2019,
  enddate = 2023,
  date_colname = DATE,
  region_colname = GEOGRAPHY_NAME,
  sector_colname = SIC_2DIGIT_NAME_SHORT,#add shortened version in
  value_colname = jobcount,#Note, this gets log'd in the function, don't do it here
  conf_interval = 95
)

#GVA and BRES
# bradford.2dig2$plot + bradford.bres.2dig2$plot
bradford.bres.2dig$plot <- bradford.bres.2dig$plot +
theme(
  axis.title.y=element_blank(),
  legend.position = "bottom") +
  ggtitle("BRES FT")

bradford.bres.2dig$plot

```








# UK INDUSTRIAL STRATEGY LINKAGE

The UK Industrial Strategy includes a [sector definition list](https://www.gov.uk/government/publications/industrial-strategy/industrial-strategy-sector-definitions-list) that lists sectors (of differing specificity / SIC code number) that the government considers part of its target sectors. As that page says, most of these SIC codes cannot capture the full range of possible firms the strategy aims to cover, and it also contains entire sections that SIC codes can't cover at all - clean energy and certain digital sectors in particular.

However, it does allow an analysis of which sectors are *definitely* within those industrial strategy categories. This section gathers job details for those specific sectors - a distinct range of types and breadth - and compares Bradford to the rest of the UK where possible, to see which sectors align with the industrial strategy and other analysis above. This is done for both BRES and Companies House job counts, to pick up on different job types.

The industrial strategy drills down below its top level sectors into 'frontier' sectors, usually (but not always) using more granular SIC codes. The plots below break the data down into these frontier and other sectors using location quotients. 

LQs are useful here as they show whether a sector has a stronger presence in a place regardlesss of the sector size or SIC digit level.

Only industrial strategy sectors with 100+ jobs in Bradford are shown (average from 2021 to 2023 for BRES, and in the latest accounts for Companies House).

Note also: job percents for Bradford show **percent of *just* total industrial strategy jobs**, not percent of all jobs.


## "Frontier" SIC sectors 

### BRES frontier sectors

```{r, fig-bresindstrat}
#| fig-width: 9

place = "Bradford"

#Blue peter this up
#Done here: https://github.com/DanOlner/RegionalEconomicTools/blob/1d5df72210586d0570cd687534bcfb3c98827ece/prepcode/misc_checks.R#L1991
#save a copy of that to save processing time in quarto
yeartoplot.lqfiltered <- readRDS('../local/data/bres_indstrat_LQs.rds')
#yeartoplot.lqfiltered <- readRDS('local/data/bres_indstrat_LQs.rds')

#If I could plot both and space them out, that would be good (could get Bradford change showing too)
p <- LQ_baseplot(df = yeartoplot.lqfiltered %>% filter(is_frontier == 1), alpha = 0.1, shape = 0, sector_name = sic_namefrom_indstrat_combo, LQ_column = LQ, change_over_time = slope)

p <- addplacename_to_LQplot(df = yeartoplot.lqfiltered %>% filter(is_frontier == 1), plot_to_addto = p, 
                            placename = place, shapenumber = 16,
                            min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                            value_column = jobcount_movingav, sector_regional_proportion = sector_regional_proportion,
                            region_name = GEOGRAPHY_NAME,
                            sector_name = sic_namefrom_indstrat_combo, change_over_time = slope, LQ_column = LQ,
                            text = 7, value_col_ismoney = F)

p <- p + 
  # coord_cartesian(xlim = c(0.1,7)) +
  ggtitle("IndStrat frontier sector")

p

```

### Companies house frontier sectors


```{r, fig-chindstrat}
#| fig-width: 9

#Blue peter this up
#Done here: https://github.com/DanOlner/RegionalEconomicTools/blob/1d5df72210586d0570cd687534bcfb3c98827ece/prepcode/misc_checks.R#L1991
#save a copy of that to save processing time in quarto
yeartoplot.lqfiltered.ch <- readRDS('../local/data/companieshouse_indstrat_LQs.rds')

p1 <- LQ_baseplot(df = yeartoplot.lqfiltered.ch %>% filter(is_frontier == 1), alpha = 0.03, sector_name = sic_namefrom_indstrat_combo, 
                  LQ_column = LQ, change_over_time = slope)

# debugonce(addplacename_to_LQplot)
p1 <- addplacename_to_LQplot(df = yeartoplot.lqfiltered.ch %>% filter(is_frontier == 1), plot_to_addto = p1, 
                             placename = place, shapenumber = 16,
                             # min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                             value_column = total_jobs, sector_regional_proportion = sector_regional_proportion,
                             region_name = localauthority_name,
                             sector_name = sic_namefrom_indstrat_combo, change_over_time = slope, LQ_column = LQ,
                             value_col_ismoney = F, text = 7)
                             # value_col_ismoney = F, text = 7, maxLQvalmultiplier = 2,useplacenameforminmaxdisplay = T, overridetextpos = 14)

p1 <- p1 + 
  # coord_cartesian(xlim = c(0.1,7)) +
  ggtitle("IndStrat frontier sector")

p1


```

## Other industrial strategy SIC sectors 

### BRES other sectors

```{r, fig-bresindstrat2}
#| fig-width: 9

p <- LQ_baseplot(df = yeartoplot.lqfiltered %>% filter(is_frontier == 0), alpha = 0.1, shape = 0, sector_name = sic_namefrom_indstrat_combo, LQ_column = LQ, change_over_time = slope)

p <- addplacename_to_LQplot(df = yeartoplot.lqfiltered %>% filter(is_frontier == 0), plot_to_addto = p, 
                            placename = place, shapenumber = 16,
                            min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                            value_column = jobcount_movingav, sector_regional_proportion = sector_regional_proportion,
                            region_name = GEOGRAPHY_NAME,
                            sector_name = sic_namefrom_indstrat_combo, change_over_time = slope, LQ_column = LQ,
                            text = 7, value_col_ismoney = F)

p <- p + 
  # coord_cartesian(xlim = c(0.1,7)) +
  ggtitle("IndStrat other sector")

p


```

### Companies House other sectors


```{r, fig-chindstrat2}
#| fig-width: 9
 
p2 <- LQ_baseplot(df = yeartoplot.lqfiltered.ch %>% filter(is_frontier == 0), alpha = 0.03, sector_name = sic_namefrom_indstrat_combo, 
                  LQ_column = LQ, change_over_time = slope)

# debugonce(addplacename_to_LQplot)
p2 <- addplacename_to_LQplot(df = yeartoplot.lqfiltered.ch %>% filter(is_frontier == 0), plot_to_addto = p2, 
                             placename = place, shapenumber = 16,
                             # min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                             value_column = total_jobs, sector_regional_proportion = sector_regional_proportion,
                             region_name = localauthority_name,
                             sector_name = sic_namefrom_indstrat_combo, change_over_time = slope, LQ_column = LQ,
                             value_col_ismoney = F, text = 7)
                             # value_col_ismoney = F, text = 7, maxLQvalmultiplier = 2,useplacenameforminmaxdisplay = T, overridetextpos = 14)

p2 <- p2 + 
  # coord_cartesian(xlim = c(0.1,7)) +
  ggtitle("IndStrat other sector")

p2



```



# ANNEX: Closer look at jobs vs GVA percent change in selected sectors.

These plots show **percent change in chained volume GVA and job count** on different axes for some selected sectors, with **Bradford's change overlaid in blue** and job count plus percent of regional economy included in title (for the most recent year). It uses a 3 year moving average to smooth out jumps.

Some are chosen to illustrate their growth relative to other places, others to highlight productivity or job change issues.

(See the [full catalogue here](https://danolner.github.io/RegionalEconomicTools/quarto_docs/Bradford_2D_gvajobs_percentchangeplots_July2025.html){target="_blank"} for every sector).

Only places where a sector makes up 1.5% or more of the economy are included - except for Bradford, which is included in all plots.

This is done for the shorter 2 digit SIC categories in the regional GVA data, joined to BRES jobs data.

Any arrows in the **top left grey diagonal section** will have seen productivity drops (measured as GVA per FT job) over that time. In the bottom left diagonal, productivity went up.

Arrows in the bottom right quadrant have seen **a combination of productivity increases with job decreases**.

**NOTE: All sectors are included even if they're very small proportions of Bradford's economy - so percent changes could appear large for very tiny sectors. Look at the job count in the title to check. Including them all for completeness.**


### Head offices / consultancy

```{r, fig-gvajobshoc}

gva.jobs.ITL3.for2dplot <- readRDS('../local/data/gvajobsfor2dplot.rds')

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('head office',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
    # coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```


### Architecture/Engineering/Tech testing

```{r, fig-gvajobsarcheng}

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('arch/eng',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") +
    coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```


### Metals

```{r, fig-gvajobsmetals}

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('metals',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") +
    coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```

### Machinery/transport manufacture

```{r, fig-gvajobs_mtf}

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('transp manuf',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  #Check... 
  # gvabres %>% filter(Region_name == "Bradford", SIC07_description_shortened == sector, DATE %in% c(2021:2023)) %>% View
  
  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
    # coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```



### Textiles

```{r, fig-textiles}

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('textiles',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
    # coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```

### Other professional / sci / tech

```{r, fig-gvajobsPST}

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('other ps',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
    # coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```


### ICT

```{r, fig-gvajobsict}

gva.jobs.ITL3.for2dplot <- readRDS('../local/data/gvajobsfor2dplot.rds')

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('ict',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") + 
    coord_cartesian(xlim = c(-50,130), ylim = c(-30,150))
  
  print(p)
  


```




### Buildings

```{r, fig-gvajobsB}

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('buildings',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
    # coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```


### Finance

```{r, fig-gvajobsFin}

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('finance',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
    # coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```

### Real estate

```{r, fig-gvajobsrealestate}

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('real est',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
    # coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```





### Office administration and business support activities

```{r, fig-gvajobsadmin}

sector <- "Admin"

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
    # coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```


### Chemicals

```{r, fig-gvajobsCh}

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('chemicals',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
    # coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```

### Food manufacture

```{r, fig-gvajobsFM}

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('food manuf',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
    # coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```


### Research / advertising

```{r, fig-gvajobsRA}

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('research',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
    # coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```

### Specialised construction

```{r, fig-gvajobsSC}

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('spec con',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
    # coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```


### Transport

```{r, fig-gvajobsTr}

sector <-  gva.jobs.ITL3.for2dplot$SIC07_description_shortened[qg('transport',gva.jobs.ITL3.for2dplot$SIC07_description_shortened)] %>% unique

placestokeep <- gva.jobs.ITL3.for2dplot %>% 
  filter(!is.na(jobcount_movingav)) %>% 
  filter(DATE == max(DATE), SIC07_description_shortened == sector) %>% 
  filter(JOBS_sector_regional_percent_movingav > 1.5) %>%#Keep only places where this sector makes up 1%+ of reg econ
  select(Region_name) %>% 
  distinct() %>% 
  pull

#Always plot Bradford! Add in if not there
if(!"Bradford" %in% placestokeep) placestokeep = c(placestokeep,"Bradford")


p <- twod_percentplot(
    df = gva.jobs.ITL3.for2dplot %>% filter(SIC07_description_shortened == sector, Region_name %in% placestokeep),
    category_var = Region_name,
    x_var = gva_movingav,
    y_var = jobcount_movingav,#
    # y_var = JOBS_sector_regional_percent_movingav,#this shows structural change better - jobs have grown nominally in most sectors (but breaks GVA/job diagonal)
    timevar = DATE,
    label_var = `gva/job moving av`,
    category_var_value_to_highlight = 'Bradford',
    start_time = 2016,
    end_time = 2022
  )
  
  bradjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(jobcount_movingav) %>% pull 
  bradpercentjobs <- gva.jobs.ITL3.for2dplot %>% filter(!is.na(jobcount_movingav), Region_name == "Bradford", SIC07_description_shortened == sector, DATE == 2022) %>% select(JOBS_sector_regional_percent_movingav) %>% pull %>% round(2)

  
  p <- p + 
    ggtitle(paste0(sector,': ', round(bradjobs/1000,1), 'K jobs in Bradford, ',bradpercentjobs,'% of tot')) +
    xlab("GVA % change 2015/17 av to 2021/23 av") +
    ylab("JOB COUNT % change 2015/17 av to 2021/23 av") 
    # coord_cartesian(xlim = c(-50,130), ylim = c(-30,100))
  
  print(p)
  


```


