---
title: "Bradford sectors and clusters: past, present"
author:
  - name: "Dan Olner"
format: 
  html:
    output-file: "Bradford_sectorclusters_July2025"
    fig-width: 8
    fig-height: 8
execute:
  echo: false
  message: false
  warning: false
fig-responsive: false
collapse: false
---

```{r setup}
library(tidyverse)
# library(spdep)
library(sf)#for geo stuff
# library(tmap)#for mapping 
library(zoo)#for making smoothed moving averages
library(patchwork)#combine ggplots easily
library(plotly)#For interactive plots
# # library(ggdist)
# library(tidyr)
# library(distributional)
library(RColorBrewer)

source('../functions/misc_functions.R')
source('../functions/adhoc_functions.R')
#source('functions/misc_functions.R')
#source('functions/adhoc_functions.R')

options(scipen = 999)

#Set ggplot theme
theme_set(theme_light())


```

# SUMMARY





# HEADLINES FROM THE DATA ANALYSIS



## Explanation of the data used



## Manufacturing / production sectors



## Service / other sectors






```{r gvaperhour_dataprep}
# ITL3 Hours % change v GVA % change on average----

#So need the number of hours and chained volume GVA
#Hours count is in the per hour productivity sheet
#GVA in the region by industry sheet

#hours worked first
url1 <- 'https://www.ons.gov.uk/file?uri=/employmentandlabourmarket/peopleinwork/labourproductivity/datasets/subregionalproductivitylabourproductivitygvaperhourworkedandgvaperfilledjobindicesbyuknuts2andnuts3subregions/current/labourproductivityitls1.xlsx'
p1f <- tempfile(fileext=".xlsx")
download.file(url1, p1f, mode="wb")

#"Productivity Hours Worked per Week; ITL2 and ITL3 subregions (constrained to ITL1), 2004 - 2023"
hoursworked <- readxl::read_excel(path = p1f,range = "Productivity Hours!A5:W247")

names(hoursworked) <- gsub(x = names(hoursworked), pattern = ' ', replacement = '_')

hoursworked = hoursworked %>%
  filter(ITL_level == 'ITL3') %>% 
  pivot_longer(Hours_2004:Hours_2023, names_to = 'year', values_to = 'hoursworkedperweek') %>% #get most recent year
  mutate(
    year = substr(year,7,10),
    year = as.numeric(year)
  ) %>% 
  select(-ITL_level)



#Then ITL3 CV for all industries (so has to include imputed rent as we can't sum CV)
url2 <- 'https://www.ons.gov.uk/file?uri=/economy/grossvalueaddedgva/datasets/nominalandrealregionalgrossvalueaddedbalancedbyindustry/current/regionalgrossvalueaddedbalancedbyindustryandallinternationalterritoriallevelsitlregions.xlsx'
p2f <- tempfile(fileext=".xlsx")
download.file(url2, p2f, mode="wb") 

#"Table 3b: ITL3, chained volume measures in 2019 money value, pounds million [note 2]"
itl3.all <- readxl::read_excel(path = p2f, range = "Table 3b!A2:AD11468")

names(itl3.all) <- gsub(x = names(itl3.all), pattern = ' ', replacement = '_')

itl3.all = itl3.all %>% 
  filter(SIC07_description == 'All industries') %>% 
  pivot_longer(`1998`:names(itl3.all)[length(names(itl3.all))], names_to = 'year', values_to = 'value') %>% #get most recent year
  mutate(year = as.numeric(year)) %>% 
  rename(gva_cv = value)



#Check link between two is valid... tick
# table(unique(hoursworked$ITL_code) %in% itl3.all$ITL_code)

#2004:2023
# unique(hoursworked$year)[unique(hoursworked$year) %in% unique(itl3.all$year)]

#Get list of corecities
corecities <- getdistinct('sheffield|Belfast|Birmingham|Bristol|Cardiff|Glasgow|Leeds|Liverpool|Manchester|Tyne|Nottingham', itl3.all$Region_name)

#Yes, all in there, need to remove a few...
corecities <- corecities[!grepl(x = corecities, pattern = 'Greater|shire', ignore.case = T)]
corecities <- corecities[order(corecities)]

#Check match also works in hours worked... tick
# table(corecities %in% hoursworked$Region_name)



#Get London ITL3s
itl_lookup = read_csv("../data/LAD_(December_2024)_to_LAU1_to_ITL3_to_ITL2_to_ITL1_(January_2025)_Lookup_in_the_UK.csv")
#itl_lookup = read_csv("data/LAD_(December_2024)_to_LAU1_to_ITL3_to_ITL2_to_ITL1_(January_2025)_Lookup_in_the_UK.csv")

#Get all London ITL2 zones then use to get unique ITL3 zones
itl3.lon = itl_lookup %>%
  filter(qg('london',ITL225NM)) %>% 
  select(ITL325CD,ITL325NM) %>% 
  distinct()

#Check itl3 code match... tick
# table(itl3.lon$ITL325CD %in% hoursworked$ITL_code)
# table(itl3.lon$ITL325CD %in% itl3.all$ITL_code)
#tick
# table(itl3.lon$ITL325NM %in% hoursworked$Region_name)
# table(itl3.lon$ITL325NM %in% itl3.all$Region_name)


```






# ONS GVA AND JOBS DATA

## Structural /  broad sector GVA change 

```{r struct}
#| fig-width: 8
#| fig-height: 16

# LOAD ITL3 GVA DATA----

#ITL3 zones, SIC sections, current prices, 2023 latest data (minus imputed rent)
df = read_csv('https://bit.ly/rtasteritl3noir')

#This one has imputed rent included if you want it
# df = read_csv('https://bit.ly/rtasteritl3')



#Finding location quotients for each year in the data 
#And then overwriting the df variable with the result
df = df %>%
  group_split(year) %>%
  map(add_location_quotient_and_proportions,
      regionvar = Region_name,
      lq_var = SIC07_description,
      valuevar = value) %>% 
  bind_rows()


#1. Set window for rolling average: 3 years
smoothband = 3

#2. Make new column with rolling average in (applying to all places and sectors)
#Years are already in the right order, so will roll apply across those correctly
df = df %>% 
  group_by(Region_name,SIC07_description) %>% 
  mutate(
    sector_regional_prop_movingav = rollapply(
      sector_regional_proportion,smoothband,mean,align='center',fill=NA
      )
  ) %>% 
  ungroup()


#Get set pastel colours for each SIC section
n <- length(unique(df$SIC07_description))
set.seed(12)
qual_col_pals = RColorBrewer::brewer.pal.info[RColorBrewer::brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(RColorBrewer::brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

randomcols <- col_vector[10:(10+(n-1))]

#Do once!
getplot <- function(place){
  
  df = df %>% 
    filter(!qg("Activities of households|agri",SIC07_description)) %>% 
    mutate(
      SIC07_description = reduceSICnames(names = SIC07_description, level = 'section')
    )
  
  top12sectors = df %>%
    filter(
      Region_name == place,
      year == max(year) - 1#to get 3 year moving average value
    ) %>%
    arrange(-sector_regional_prop_movingav) %>%
    slice_head(n= 12) %>%
    select(SIC07_description) %>%
    pull()
  
  
  ggplot(df %>% 
           filter(
             Region_name == place, 
             SIC07_description %in% top12sectors,
             !is.na(sector_regional_prop_movingav)#Remove NAs created by 3 year moving average
           ), 
         aes(x = year, y = sector_regional_prop_movingav * 100, colour = fct_reorder(SIC07_description,-sector_regional_prop_movingav) )) +
    geom_point() +
    geom_line() +
    scale_color_manual(values = setNames(randomcols,unique(df$SIC07_description))) +#set manual pastel colours matching sector name
    ylab('Sector: percent of regional economy') +
    theme(plot.title = element_text(face = 'bold')) +
    labs(colour = 'SIC section') +
    # scale_y_log10() +
    ggtitle(paste0(place, ' GVA % of economy over time\nTop 12 sectors in ',max(df$year),'\n (', smoothband,' year moving average)'))
  
  
}

places = getdistinct('bradford|kirklees|calderdale|leeds|wakefield',df$Region_name)

plots <- map(places, getplot)

patchwork::wrap_plots(plots, ncol = 1)


```




## GVA per full time job over time: Bradford v other ITL3 zones for 2 digit SIC sectors

These plots use **linked regional GVA and BRES jobs data** to examine productivity, output and job changes. To make this link, sector categories have to be reduced to those in the ONS GVA data (46 categories rather than the full 88 two digit SICs).

BRES does contain more detail - more on that below and in the separate [BRES 5 digit breakdown](https://danolner.github.io/RegionalEconomicTools/quarto_docs/Bradford_BRES_5digit_LQs_July2025.html){target="_blank"}.

The plots below are measured in **GVA per job parts per million of UK economy** - equivalent to e.g. 1% being "1 in 100 parts of economy", 0.1 here is "1 in 10 millionth of whole economy". So it's again examining **relative change in output per FT worker** across all sectors.

GVA per job as a measure needs some caution: output is also due to other productive factors, so high capital intensity sectors (power, telecoms, other ICT though not all) can appear overly productive per worker. It's more representative for service sectors where an employee's output more clearly correlates to total output, such as legal services.


```{r gvajobs1}
#| fig-height: 20
#| fig-width: 8.5

#2015 to 2023, linked on the regional GVA bespoke SIC categories
#Note, can use lookup if need to also sum for sections (which we can do here cos current prices)
gvabres <- readRDS('../data/regionalGVA_plus_BRESjobcounts/regionalGVA_currentprices_BRES_FT_jobcount_bespoke2digitSIC_nONLY_MATCHING_GEOGs_2015_2023.rds') 
#gvabres <- readRDS('data/regionalGVA_plus_BRESjobcounts/regionalGVA_currentprices_BRES_FT_jobcount_bespoke2digitSIC_nONLY_MATCHING_GEOGs_2015_2023.rds') 

#Make abbreviated sector names
shortsectornames = tibble(
  SIC07_description = unique(gvabres$SIC07_description),
  SIC07_description_shortened = c(
  "Furniture",
  "Machinery repair/install",
  "Buildings",
  "Civil eng",
  "Spec constr",
  "Motor",
  "Wholesale",
  "Retail",
  "Transport",
  "Warehousing",
  "Post",
  "Accommodation",
  "Food",
  "Media",
  "ICT",
  "Real estate",
  "Legal",
  "Head office/consult",
  "Arch/Eng",
  "Research/Ads",#This is a horrible combo that makes no sense! 72/73
  "Other PSTech",#"Prof/sci/tech",
  "Rental",
  "Employment",
  "Cleaning/landscapes",
  "Admin",
  "Health",
  "Care",
  "SocialWork",
  "Arts/ent",
  "Recreation",
  "Membership",
  "Repair personal/computer",
  "Personal",
  "Agri/mining",
  "Food manuf",
  "Textiles",
  "Wood/paper",
  "Chemicals",
  "Metals",
  "Electronics",
  "Mach/transp manuf",
  "Utilities",
  "Finance",
  "Public admin",
  "Education",
  "Households"
  )
)

#save that for use elsewhere!
# write_csv(shortsectornames, 'data/shortsectornames_for_regionalGVA_2digitSICs.csv')

gvabres <- gvabres %>% 
  left_join(
    shortsectornames, by = 'SIC07_description'
  )%>% 
  filter(!qg('households|membership', SIC07_description_shortened))

#Look at version including codes used in GVA
#Ah not a good source as it's filtered out <100 FTs (so e.g. missing sci, which is only 50)
# gvabres %>% select(SIC_2DIGIT_CODE_GVA2023,SIC07_description,SIC07_description_shortened) %>% distinct() %>% View

#Would also quite like to split manufacturing off
#Check in orig GVA file for the codes...
#This is the full list - some won't match but enough to label the correct ones at 2 digit for those that do
productionsectors <- c(
  'A-E',
  'AB (1-9)',
  'C (10-33)',
  'CA (10-12)',
  'CB (13-15)',
  'CC (16-18)',
  'CD-CG (19-23)',
  'CH (24-25)',
  'CI-CJ (26-27)',
  'CK-CL (28-30)',
  'CM (31-33)',
  '31-32',
  '33',
  'DE (35-39)',
  'F (41-43)',
  '41',
  '42',
  '43'
)

#Yep
# table(productionsectors %in% gvabres$SIC_2DIGIT_CODE_GVA2023)
# unique(gvabres$SIC_2DIGIT_CODE_GVA2023)[unique(gvabres$SIC_2DIGIT_CODE_GVA2023) %in% productionsectors]

#Label production sectors
gvabres <- gvabres %>% 
  mutate(
    productionsector = ifelse(
      SIC_2DIGIT_CODE_GVA2023 %in% productionsectors,
      'production','other'
    )
  )



#For current prices - what we can do:
#Find GVA as a proportion of total UK output for those years
#Then that proportion per FT job
#To see how productivity has shifted about internally / how sectors have changed relative to each other

#That version is missing some places so isn't entire UK
#Let's get entire UK's GVA per year
#Can just sum from any current price source
#2023:
allgva <- read_csv('../data/regionalGVA/regionalGVA_currentprices_ITL2_SIC_2DIGIT_LONG_2023.csv')
#allgva <- read_csv('data/regionalGVA/regionalGVA_currentprices_ITL2_SIC_2DIGIT_LONG_2023.csv')




allgva <- allgva %>% 
  group_by(year) %>% 
  summarise(totalUKGVA_CP = sum(value)) %>% 
  ungroup()

#Add that in...
gvabres <- gvabres %>% 
  left_join(allgva, by = c('DATE'= 'year')
  )

#Some NAs in jobs - I think the issue with the missing values is just early years with no agri data...
#filter
gvabres <- gvabres %>% filter(!is.na(JOBCOUNT))


#Find gva proportion of UK total then the job's proportion of that
gvabres <- gvabres %>%
  mutate(
    gva_proportionofUKtotal = GVA / totalUKGVA_CP,
    job_output_propofUKtotal = gva_proportionofUKtotal / JOBCOUNT
  )


#Get regional GVA data version of SIC sections 
#To plot each by
# siclookupfromreg.gva <- read_csv('../data/SIClookup_forBRES_REGIONALGVA_2023_JOIN.csv')
#siclookupfromreg.gva <- read_csv('data/SIClookup_forBRES_REGIONALGVA_2023_JOIN.csv')



smoothband = 3

gvabres <- gvabres %>%
  group_by(Region_name,SIC07_description) %>%
  mutate(
    jobcount_movingav = rollapply(JOBCOUNT,smoothband,mean,align='center',fill=NA),
    gva_movingav = rollapply(GVA,smoothband,mean,align='center',fill=NA),
    job_output_propofUKtotal_movingav = rollapply(job_output_propofUKtotal,smoothband,mean,align='center',fill=NA),
    job_PPM_ofUKtotal_movingav = job_output_propofUKtotal_movingav * 1000000#1% is 1 part per 100. Which we find by multiplying up the prop by 100. PPM is just x * a million.
  ) %>% 
  ungroup()



#ggplot won't give control over order very well... let's add on top manually, more control
prod.data <- gvabres %>% 
  mutate(
    DATE = DATE - 2000,#Make dates 2 digit, more readable on axis
    SIC07_description_shortened = fct_reorder(SIC07_description_shortened, job_PPM_ofUKtotal_movingav, .desc = T),
    highlow = ifelse(SIC07_description_shortened %in% levels(SIC07_description_shortened)[1:23],'high','low'),#Break into two plots to stack, split by total av productivity
    sectorgrouping = cut_number(as.integer(SIC07_description_shortened), 8) %>% as.integer
  ) %>% 
  # filter(!qg('hous',SIC07_description))
  filter(!qg('hous',SIC07_description), !is.na(job_output_propofUKtotal_movingav))

#Break into n groups, using factor order created in previous line
# sectorgrouping = ntile(as.integer(prod.data$SIC07_description_shortened), 5)

  
#table(prod.data$sectorgroupings)

#Test
# debugonce(plotprod)
# plotprod(prod.data %>% filter(sectorgrouping==6))

plotz <- map(prod.data %>% group_split(sectorgrouping),plotprod)

patchwork::wrap_plots(plotz,ncol = 1)

#Quick look
#gvabres %>% filter(qg('architecture',SIC07_description_shortened), qg('bradford',Region_name)) %>% View


```



## Location quotients for GVA and jobs in 2 digit SIC sectors: most recent and change over time

These LQ plots show location quotients for linked GVA and jobs data for Bradford, overlaid on top of LQs for all other ITL3 zones, to give a sense of what the spread is across Great Britain.

A location quotient is the ratio of a place's sector proportion to its national proportion. So e.g. if manufacturing in Bradford made up 20% of its economy compared to maufacturing being 10% of the UK economy, its LQ in Bradford would be 2.

Here’s an explanation of how to read these location quotient plots:

* Location quotients for sectoral GVA in the selected region
* Sectors to the right of the blue line are more concentrated in this region when compared to the national average. (e.g. LQ = 10, it is 10x more concentrated). To the left are less concentrated than nationally.
* Green circles have been increasing in concentration over time (larger circles increasing the most)
* Red circles have been decreasing in concentration over time (larger circles decreasing the most)
* Black horizontal lines show the min and max of concentration in the data’s history (from 2015 to 2023) (so, often, large green circles will be at the right of this line because sector concentration has been steadily increasing)
* Per sector GVA money amounts and percent of that region’s entire economy are shown on the right.

### Production sectors


```{r productionLQs}
#| fig-width: 9
#| fig-height: 10

#Get two versions of LQ/proportions - job count and GVA
gva.props <- gvabres %>% 
  group_split(DATE) %>%
  map(add_location_quotient_and_proportions,
      regionvar = Region_name,
      lq_var = SIC07_description_shortened,
      valuevar = GVA) %>% 
  bind_rows()

job.props <- gvabres %>% 
  group_split(DATE) %>%
  map(add_location_quotient_and_proportions,
      regionvar = Region_name,
      lq_var = SIC07_description_shortened,
      valuevar = JOBCOUNT) %>% 
  bind_rows()

# LET'S SEPARATE GVA AND JOBS OUT FOR BRADFORD AND COMPARE TO EVERYWHERE ELSE TOO
place = "Bradford"

#FUNCTION
#Split both gva.props and job.props by subsector list
makegvajoblqplots <- function(keepthesesectors){
  
  # gva.sub <- gva.props %>% filter(SIC07_description_shortened %in% keepthesesectors)

  LQ_slopes <- compute_slope_or_zero(
    data = gva.props, 
    Region_name, SIC07_description_shortened,#slopes will be found within whatever grouping vars are added here
    y = LQ_log, x = DATE)
  
  
  #Filter down to a single year... we may want to smooth years, let's see
  yeartoplot <- gva.props %>% filter(DATE == max(DATE))#use latest year
  
  #Add slopes into data to get LQ plots
  yeartoplot <- yeartoplot %>% 
    left_join(
      LQ_slopes,
      by = c('Region_name', 'SIC07_description_shortened')
    )
  
  #Get min/max values for LQ over time as well, for each sector and place, to add as bars so range of sector is easy to see
  minmaxes <- gva.props %>% 
    group_by(Region_name, SIC07_description_shortened) %>% 
    summarise(
      min_LQ_all_time = min(LQ),
      max_LQ_all_time = max(LQ)
    )
  
  #Join min and max
  yeartoplot <- yeartoplot %>% 
    left_join(
      minmaxes,
      by = c('Region_name', 'SIC07_description_shortened')
    )
  
  
  #Shorten here to subset of sectors
  yeartoplot <- yeartoplot %>% 
    filter(SIC07_description_shortened %in% keepthesesectors)
  
  
  sectorLQorder <- gva.props %>% filter(
    Region_name == place,
    DATE == max(DATE)#use latest data
  ) %>% 
    arrange(-LQ) %>% 
    select(SIC07_description_shortened) %>% 
    pull()
  
  
  #Turn the sector column into a factor and order by LCR's LQs
  yeartoplot$SIC07_description_shortened <- factor(yeartoplot$SIC07_description_shortened, levels = sectorLQorder, ordered = T)
  
  #If I could plot both and space them out, that would be good (could get Bradford change showing too)
  p <- LQ_baseplot(df = yeartoplot, alpha = 0.1, shape = 0, sector_name = SIC07_description_shortened, 
                   LQ_column = LQ, change_over_time = slope)
  
  p <- addplacename_to_LQplot(df = yeartoplot, plot_to_addto = p, 
                              placename = place, shapenumber = 16,
                              min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                              value_column = GVA, sector_regional_proportion = sector_regional_proportion,
                              region_name = Region_name,
                              sector_name = SIC07_description_shortened, change_over_time = slope, LQ_column = LQ,
                              text = 7)
  
  p <- p + 
    # coord_cartesian(xlim = c(0.1,7)) +
    ggtitle("GVA")
  
  
  
  
  # job.sub <- job.props %>% filter(SIC07_description_shortened %in% keepthesesectors)
  
  # REPEAT FOR JOBS / BRADFORD (Using same factor order?)
  LQ_slopes <- compute_slope_or_zero(
    data = job.props, 
    Region_name, SIC07_description_shortened,#slopes will be found within whatever grouping vars are added here
    y = LQ_log, x = DATE)
  
  
  #Filter down to a single year... we may want to smooth years, let's see
  yeartoplot <- job.props %>% filter(DATE == max(DATE))#use latest year
  
  #Add slopes into data to get LQ plots
  yeartoplot <- yeartoplot %>% 
    left_join(
      LQ_slopes,
      by = c('Region_name', 'SIC07_description_shortened')
    )
  
  #Get min/max values for LQ over time as well, for each sector and place, to add as bars so range of sector is easy to see
  minmaxes <- job.props %>% 
    group_by(Region_name, SIC07_description_shortened) %>% 
    summarise(
      min_LQ_all_time = min(LQ),
      max_LQ_all_time = max(LQ)
    )
  
  #Join min and max
  yeartoplot <- yeartoplot %>% 
    left_join(
      minmaxes,
      by = c('Region_name', 'SIC07_description_shortened')
    )
  
  #Shorten here to subset of sectors
  yeartoplot <- yeartoplot %>% 
    filter(SIC07_description_shortened %in% keepthesesectors)
  
  
  
  #USE SECTOR ORDER FROM GVA LQS ABOVE
  yeartoplot$SIC07_description_shortened <- factor(yeartoplot$SIC07_description_shortened, levels = sectorLQorder, ordered = T)
  
  #If I could plot both and space them out, that would be good (could get Bradford change showing too)
  p2 <- LQ_baseplot(df = yeartoplot, alpha = 0.1, shape = 0, sector_name = SIC07_description_shortened, 
                   LQ_column = LQ, change_over_time = slope)
  
  p2 <- addplacename_to_LQplot(df = yeartoplot, plot_to_addto = p2, 
                              placename = place, shapenumber = 16,
                              min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                              value_column = JOBCOUNT, sector_regional_proportion = sector_regional_proportion,
                              region_name = Region_name,
                              sector_name = SIC07_description_shortened, change_over_time = slope, LQ_column = LQ,
                              text = 7, value_col_ismoney = F)
  
  p2 <- p2 + 
    # coord_cartesian(xlim = c(0.1,7)) +
    ggtitle("Jobs")
  
  # p2 <- p2 + scale_y_discrete(position = "right")
  p2 <- p2 +
    theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
  
  p + p2 

}

#production sector plot first
makegvajoblqplots(gva.props$SIC07_description_shortened[gva.props$productionsector == 'production'])


```


### Rest of sectors

```{r otherLQs}
#| fig-width: 9
#| fig-height: 14

makegvajoblqplots(gva.props$SIC07_description_shortened[gva.props$productionsector == 'other'])

```




## Av. percent change over time for gva vs hours worked, ITL3 zones

**Hover for place names**.

* West Yorkshire ITL3s in green - Bradford's outlined in black (95% CIs in lines).
* Core cities in blue  (95% CIs)
* London ITL3s in red/pink

And:

* Anything above/left of diagonal line is a **productivity increase**. 




### 2004 to 2013 av % change

```{r plot1}
# startyear = 2014
# endyear = 2023

#So e.g. in the plot below for 2014-2023, Rotherham's productivity increased with a **GVA increase** and a **number of hours worked *decrease***.

startyear = 2004
endyear = 2013

#Hours per week first
slopes.hours = get_slope_and_se_safely(
  hoursworked %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(hoursworkedperweek), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.hours = slopes.hours %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )


#Mark 4 SY LAs and plot
slopes.hours = slopes.hours %>% 
  ungroup() %>% 
  # mutate(
  #   SY_LAs = ifelse(qg('rotherh|doncaste|sheffie|barnsley', Region_name),'SY LA','Other'),
  #   corecities = ifelse(Region_name %in% corecities,'core city','Other'),
  #   london = ifelse(qg('london',Region_name), 'London','Other')
  # )
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'SY LA',#will overwrite sheffield as core city
      qg('bradford|kirklees|calderdale|leeds|wakefield', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
      ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','WY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.hours = slopes.hours %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )




#Then GVA CV
slopes.gva = get_slope_and_se_safely(
  itl3.all %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(gva_cv), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.gva = slopes.gva %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )



#Mark 4 WY LAs and plot
slopes.gva = slopes.gva %>% 
  ungroup() %>% 
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      qg('bradford|kirklees|calderdale|leeds|wakefield', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
    ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','WY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.gva = slopes.gva %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )


# table(slopes.hours$itl3label)
# table(slopes.gva$itl3label)


#Keep the bits we want and combine into single df
both <- slopes.gva %>% 
  select(Region_name,itl3label,
         slope_percentperyear_GVA = slope_percentperyear,
         ci95min_percentperyear_GVA = ci95min_percentperyear,
         ci95max_percentperyear_GVA = ci95max_percentperyear) %>% 
  inner_join(
    slopes.hours %>% 
      select(Region_name,
             slope_percentperyear_HOURS = slope_percentperyear,
             ci95min_percentperyear_HOURS = ci95min_percentperyear,
             ci95max_percentperyear_HOURS = ci95max_percentperyear),
    by = 'Region_name'
  )


#Poor control over visibility. Layer instead. Base plot then function.
addtogva_hours_plot = function(p,plotdata, showerrors = T, sizeval = 1, colourval = 'black', alphaval = 1){
  
  p = p + geom_point(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, group = Region_name), colour = colourval, size = sizeval, alpha = alphaval) 
  
  if(showerrors){
    
    p = p + geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, xmin = ci95min_percentperyear_HOURS, xmax = ci95max_percentperyear_HOURS),width = 0.1, size = 1, colour = colourval, alpha = alphaval) +
    geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, ymin = ci95min_percentperyear_GVA, ymax = ci95max_percentperyear_GVA),width = 0.1, size = 1, colour = colourval, alpha = alphaval) 
    
  }
  
    p  + geom_abline(slope = 1, intercept = 0, colour = 'black', alpha = 0.5, size = 2) +
      geom_hline(yintercept = 0, colour = 'black', alpha = 0.3, size = 1) +
      geom_vline(xintercept = 0, colour = 'black', alpha = 0.3, size = 1)
  
}




p = ggplot()

p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Other'), F, 1, 'black', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'London'), F, 2, 'red', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Core city'), T, 3, RColorBrewer::brewer.pal(3, "Set1")[2], 0.5)
p = addtogva_hours_plot(p, both %>% filter(Region_name == 'Bradford'), F, 10, 'black', 1)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'WY LA'), T, 7, RColorBrewer::brewer.pal(3, "Set1")[3], 1)

p = p + coord_fixed() +
  xlab("Average % change: hours worked per week") +
  ylab("Average % change: GVA") 


ggplotly(p, tooltip = 'Region_name')



```

### 2014 to 2023 av % change

```{r plot2}
startyear = 2014
endyear = 2023

# startyear = 2004
# endyear = 2013

#Hours per week first
slopes.hours = get_slope_and_se_safely(
  hoursworked %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(hoursworkedperweek), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.hours = slopes.hours %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )


#Mark 4 WY LAs and plot
slopes.hours = slopes.hours %>% 
  ungroup() %>% 
  # mutate(
  #   SY_LAs = ifelse(qg('rotherh|doncaste|sheffie|barnsley', Region_name),'WY LA','Other'),
  #   corecities = ifelse(Region_name %in% corecities,'core city','Other'),
  #   london = ifelse(qg('london',Region_name), 'London','Other')
  # )
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      qg('bradford|kirklees|calderdale|leeds|wakefield', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
      ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','WY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.hours = slopes.hours %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )




#Then GVA CV
slopes.gva = get_slope_and_se_safely(
  itl3.all %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(gva_cv), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.gva = slopes.gva %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )



#Mark 4 WY LAs and plot
slopes.gva = slopes.gva %>% 
  ungroup() %>% 
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      qg('bradford|kirklees|calderdale|leeds|wakefield', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
    ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','WY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.gva = slopes.gva %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )


# table(slopes.hours$itl3label)
# table(slopes.gva$itl3label)


#Keep the bits we want and combine into single df
both <- slopes.gva %>% 
  select(Region_name,itl3label,
         slope_percentperyear_GVA = slope_percentperyear,
         ci95min_percentperyear_GVA = ci95min_percentperyear,
         ci95max_percentperyear_GVA = ci95max_percentperyear) %>% 
  inner_join(
    slopes.hours %>% 
      select(Region_name,
             slope_percentperyear_HOURS = slope_percentperyear,
             ci95min_percentperyear_HOURS = ci95min_percentperyear,
             ci95max_percentperyear_HOURS = ci95max_percentperyear),
    by = 'Region_name'
  )


#Poor control over visibility. Layer instead. Base plot then function.
addtogva_hours_plot = function(p,plotdata, showerrors = T, sizeval = 1, colourval = 'black', alphaval = 1){
  
  p = p + geom_point(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, group = Region_name), colour = colourval, size = sizeval, alpha = alphaval) 
  
  if(showerrors){
    
    p = p + geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, xmin = ci95min_percentperyear_HOURS, xmax = ci95max_percentperyear_HOURS),width = 0.1, size = 1, colour = colourval, alpha = alphaval) +
    geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, ymin = ci95min_percentperyear_GVA, ymax = ci95max_percentperyear_GVA),width = 0.1, size = 1, colour = colourval, alpha = alphaval) 
    
  }
  
    p  + geom_abline(slope = 1, intercept = 0, colour = 'black', alpha = 0.5, size = 2) +
      geom_hline(yintercept = 0, colour = 'black', alpha = 0.3, size = 1) +
      geom_vline(xintercept = 0, colour = 'black', alpha = 0.3, size = 1)
  
}




p = ggplot()

p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Other'), F, 1, 'black', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'London'), F, 2, 'red', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Core city'), T, 3, RColorBrewer::brewer.pal(3, "Set1")[2], 0.5)
p = addtogva_hours_plot(p, both %>% filter(Region_name == 'Bradford'), F, 10, 'black', 1)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'WY LA'), T, 7, RColorBrewer::brewer.pal(3, "Set1")[3], 1)

p = p + coord_fixed() +
  xlab("Average % change: hours worked per week") +
  ylab("Average % change: GVA") 


ggplotly(p, tooltip = 'Region_name')



```









# COMPANIES HOUSE DATA


```{r, chplot}

#See here for fig-responsiveness (non)issue: https://github.com/quarto-dev/quarto-cli/issues/4836

#30 seconds-ish
# x <- Sys.time()
ch = readRDS('../../companieshouseopen/local/PROCESSED_accountextracts_n_livelist_geocoded_combined_July2025.rds')
# ch = readRDS('../companieshouseopen/local/PROCESSED_accountextracts_n_livelist_geocoded_combined_July2025.rds')
# Sys.time() - x

#get short SIC names
shorts <- readRDS('../local/data/bradford_SICs_withshortnames.rds')
#shorts <- readRDS('local/data/bradford_SICs_withshortnames.rds')

ch = ch %>%
  left_join(
    shorts %>% select(SIC_5DIGIT_CODE,SIC_SECTION_NAME_SHORT:SIC_5DIGIT_NAME_SHORT),
    by = 'SIC_5DIGIT_CODE'
  )

#Create the base summary, save, reuse
#Taken from CH code here
#https://github.com/DanOlner/companieshouseopen/blob/2ddc26d4b89716b8755c02fc6626190ad624b888/testcode/initial_datadigging.R#L417
empchange.summary <- ch %>% 
  st_set_geometry(NULL) %>% 
  filter(Employees_lastyear >= 10) %>%#Need to load cos we need to filter here so we can accurately count employees in 10+ employee firms
  group_by(localauthority_name,SIC_2DIGIT_NAME_SHORT) %>%
  summarise(
    total_employmentlastyear = sum(Employees_lastyear, na.rm = T),
    total_employmentthisyear = sum(Employees_thisyear, na.rm = T),
    employment_percentchange = percent_change(total_employmentlastyear,total_employmentthisyear),
    firmcount = n()
  ) %>% 
  ungroup() %>% 
  filter(firmcount > 5)#Reduce to sectors/places with 10+ firms
# 
# saveRDS(empchange.summary,'local/data/empchangsummary.rds')
# empchange.summary <- readRDS('../local/data/empchangsummary.rds')
#empchange.summary <- readRDS('local/data/empchangsummary.rds')

# empchange10plus <- empchange.summary %>% 
#   filter(
#     firmcount > 5,
#     Employees_lastyear >= 10#Reduce to sectors/places with 10+ firms
#   )

#Order by percent change per sector in Bradford
orderby <- empchange.summary %>% 
  filter(qg('bradford',localauthority_name)) %>%
  select(SIC_2DIGIT_NAME_SHORT,employment_percentchange) %>% 
  arrange(employment_percentchange) %>% 
  pull(SIC_2DIGIT_NAME_SHORT)

#Order by those...
empchange.summary <- empchange.summary %>% 
  mutate(
    SIC_2DIGIT_NAME_SHORT = factor(SIC_2DIGIT_NAME_SHORT, levels = orderby),
    `Place: ` = paste0(localauthority_name, ', firm count: ', firmcount,', emp last/this yr: ', total_employmentlastyear,' > ',total_employmentthisyear,' = ',round(employment_percentchange,2),'%')#for labels
    # `Place: ` = paste0(SIC_2DIGIT_NAME_SHORT,',', localauthority_name, ', firm count: ', firmcount,', emp last/this yr: ', total_employmentlastyear,' > ',total_employmentthisyear)#for labels
  ) %>% 
  filter(!is.na(SIC_2DIGIT_NAME_SHORT))#Filter these here as some will be missing spefically from Bradford, will end up NA in the factor

# table(is.na(empchange.summary$SIC_2DIGIT_NAME_SHORT))

#Get values for Bradford firms - count/%/employee counts
firmcount <- ch %>% st_set_geometry(NULL) %>% filter(localauthority_name == 'Bradford', Employees_thisyear > 0)

firmcount <- firmcount %>% 
  mutate(
    sizecategory = case_when(
      Employees_thisyear == 1 ~ "1",
      between(Employees_thisyear,2,4) ~ "2-4",
      between(Employees_thisyear,5,9) ~ "5-9",
      between(Employees_thisyear,10,20) ~ "10-20",
      between(Employees_thisyear,21,50) ~ "21-50",
      between(Employees_thisyear,51,999999) ~ "51+"
    ),
    sizecategory = factor(sizecategory, levels = c('1','2-4','5-9','10-20','21-50','51+'))
    )

#table from that to plot
firmtable <- tibble(
  `Firm size` = levels(firmcount$sizecategory),
  `Count` = table(firmcount$sizecategory),
  `Percent of firms` = paste0(round(table(firmcount$sizecategory) %>% prop.table() * 100,2),"%")
)

```


```{r eval=F}
firmcounts.by.2digit <- firmcount %>% 
  group_by(sizecategory,SIC_2DIGIT_NAME_SHORT) %>% 
  summarise(
    count = n(),
    section = max(SIC_SECTION_NAME_SHORT)
    ) %>% 
  group_by(SIC_2DIGIT_NAME_SHORT) %>% 
  mutate(
    percent = (count / sum(count)) * 100
  )

ggplot(firmcounts.by.2digit, aes(y = SIC_2DIGIT_NAME_SHORT, x = percent, fill = sizecategory)) +
  geom_bar(stat = 'identity', position = position_dodge(width = 0.8)) +
  facet_wrap(~section,ncol = 2, scales = 'free')

```



The next few plots use data extracted from individual firms' accounts **filed with Companies House between June 2024 and end of June 2025**. 

The following table shows the count (and percentage) of firms in Bradford that filed with CH, broken down by employee count.


```{r bradch_summarytable}

#escape = FALSE so that markdown in the cells (the links) is rendered correctly.
knitr::kable(firmtable, escape = FALSE, caption = "Table: Bradford count/% of firms by employee band")

```



## Comparing Companies House to BRES: location quotient plots

### Production sectors

```{r CHvBRESprod}
#| fig-width: 9

#Produced here:
#https://github.com/DanOlner/RegionalEconomicTools/blob/927f1783bfbb66293dd75870e147c858f4761ef3/bits_of_code/BradfordExplore.R#L1062
plotz <- readRDS('../local/data/ch_bres_lqplotlist.rds')
#plotz <- readRDS('local/data/ch_bres_lqplotlist.rds')

plotz[[1]] + plotz[[3]]

```

### Other sectors

```{r CHvBRESother}
#| fig-width: 9

plotz[[2]] + plotz[[4]]

```




## From Companies House firm level data: in the last year (March 2024-March 2025), percent change in employment in firms, five West Yorkshire local authorities compared to 'core cities' (hover for place names)

### Firms with 10+ employees


```{r, fig-percentempfourplaces}
#| fig-height: 20
#| label: fig-percentempfourplaces
#| fig-cap: "Companies House firm level data: percent change in employment for firms with ten+ employees between most recent accounts and previous year - core cities, with West Yorkshire's five local authorities overlaid. Hover for place name."


p <- ggplot() +
  geom_point(
    position = position_nudge(y = 0.2),
    data = empchange.summary %>% filter(
      qg('Belfast|Birmingham|Bristol|Cardiff|Glasgow|Leeds|Liverpool|Manchester|Tyne|Sheffield|Nottingham', localauthority_name)#just core cities (minus four places below)
    ),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, group = `Place: `),
    alpha = 0.5)

#Add four places
p <- p + 
  geom_point(
    position = position_nudge(y = -0.2),
    data = empchange.summary %>% filter(qg('leeds|kirklees|wakefield|calderdale',localauthority_name)),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, colour = localauthority_name, group = `Place: `),
    size = 7, shape = 17) +
  scale_colour_brewer(palette = 'Paired', direction = 1, guide = guide_legend(title = NULL)) +
  # coord_cartesian(xlim = c(-25,25)) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") 

#Add Bradford over the top to make sure visible
p <- p + 
  geom_point(
    position = position_nudge(y = -0.2),
    data = empchange.summary %>% filter(qg('bradford',localauthority_name)),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, colour = localauthority_name, group = `Place: `),
    size = 8, shape = 17) +
  scale_colour_brewer(palette = 'Paired', direction = 1, guide = guide_legend(title = NULL)) +
  # coord_cartesian(xlim = c(-25,25)) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") 



#p

# ggplotly(p, width = 800, height = 900, tooltip = 'localauthority_name')
ggplotly(p, tooltip = 'Place: ')


```

###  Firms with fewer than 10 employees



```{r, fig-percentempfourplaces_lessthan10}
#| fig-height: 20
#| label: fig-percentempfourplaces_lessthan10
#| fig-cap: "From Companies House firm level data: percent change in employment for firms with fewer than 10 employees, between most recent accounts and previous year - core cities, with West Yorkshire's five local authorities overlaid. Hover for place name."

empchange.summary <- ch %>% 
  st_set_geometry(NULL) %>% 
  filter(Employees_lastyear < 10) %>%#Need to load cos we need to filter here so we can accurately count employees in 10+ employee firms
  group_by(localauthority_name,SIC_2DIGIT_NAME_SHORT) %>%
  summarise(
    total_employmentlastyear = sum(Employees_lastyear, na.rm = T),
    total_employmentthisyear = sum(Employees_thisyear, na.rm = T),
    employment_percentchange = percent_change(total_employmentlastyear,total_employmentthisyear),
    firmcount = n()
  ) %>% 
  ungroup() %>% 
  filter(firmcount > 5)#Reduce to sectors/places with 10+ firms
# 
# saveRDS(empchange.summary,'local/data/empchangsummary.rds')
# empchange.summary <- readRDS('../local/data/empchangsummary.rds')
#empchange.summary <- readRDS('local/data/empchangsummary.rds')

# empchange10plus <- empchange.summary %>% 
#   filter(
#     firmcount > 5,
#     Employees_lastyear >= 10#Reduce to sectors/places with 10+ firms
#   )

#Order by percent change per sector in Bradford
orderby <- empchange.summary %>% 
  filter(qg('bradford',localauthority_name)) %>%
  select(SIC_2DIGIT_NAME_SHORT,employment_percentchange) %>% 
  arrange(employment_percentchange) %>% 
  pull(SIC_2DIGIT_NAME_SHORT)

#Order by those...
empchange.summary <- empchange.summary %>% 
  mutate(
    SIC_2DIGIT_NAME_SHORT = factor(SIC_2DIGIT_NAME_SHORT, levels = orderby),
    `Place: ` = paste0(localauthority_name, ', firm count: ', firmcount,', emp last/this yr: ', total_employmentlastyear,' > ',total_employmentthisyear,' = ',round(employment_percentchange,2),'%')#for labels
    # `Place: ` = paste0(SIC_2DIGIT_NAME_SHORT,',', localauthority_name, ', firm count: ', firmcount,', emp last/this yr: ', total_employmentlastyear,' > ',total_employmentthisyear)#for labels
  ) %>% 
  filter(!is.na(SIC_2DIGIT_NAME_SHORT))#Filter these here as some will be missing spefically from Bradford, will end up NA in the factor

# table(is.na(empchange.summary$SIC_2DIGIT_NAME_SHORT))




p <- ggplot() +
  geom_point(
    position = position_nudge(y = 0.2),
    data = empchange.summary %>% filter(
      qg('Belfast|Birmingham|Bristol|Cardiff|Glasgow|Leeds|Liverpool|Manchester|Tyne|Sheffield|Nottingham', localauthority_name)#just core cities (minus four places below)
    ),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, group = `Place: `),
    alpha = 0.5)

#Add four places
p <- p + 
  geom_point(
    position = position_nudge(y = -0.2),
    data = empchange.summary %>% filter(qg('leeds|kirklees|wakefield|calderdale',localauthority_name)),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, colour = localauthority_name, group = `Place: `),
    size = 7, shape = 17) +
  scale_colour_brewer(palette = 'Paired', direction = 1, guide = guide_legend(title = NULL)) +
  # coord_cartesian(xlim = c(-25,25)) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") 

#Add Bradford over the top to make sure visible
p <- p + 
  geom_point(
    position = position_nudge(y = -0.2),
    data = empchange.summary %>% filter(qg('bradford',localauthority_name)),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, colour = localauthority_name, group = `Place: `),
    size = 8, shape = 17) +
  scale_colour_brewer(palette = 'Paired', direction = 1, guide = guide_legend(title = NULL)) +
  # coord_cartesian(xlim = c(-25,25)) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") 

p <- p + coord_cartesian(xlim = c(-40,80))

#p

# ggplotly(p, width = 800, height = 900, tooltip = 'localauthority_name')
ggplotly(p, tooltip = 'Place: ')


```




## Just for Bradford, individual firms with 10+ employees, percent employee change from last accounts (hover for firm names)

This section is included to be able to identify individual firms (with 10+ employees) by hovering over the points. Each is plotted by what percentage their employee number grew between their latest accounts and a year prior.

```{r, fig-indivfirmsSY}
#| fig-height: 10
#| fig-width: 9

# fig-cap: "Percent change in employment between most recent accounts and previous year, ALL Bradford firms with 10+ employees, hover for firm name."

firm.change <- ch %>% 
  filter(localauthority_name == 'Bradford', Employees_thisyear >= 10 & Employees_lastyear >= 10) %>% #10+ in both time periods
  mutate(
    employee_percentchange = percent_change(Employees_lastyear,Employees_thisyear)
  )

#As well as a minimum of ten, let's break down by firm size...
#What's the spread?
# ggplot(firm.change, aes(x = Employees_thisyear)) +
#   geom_density()

firm.change <- firm.change %>% 
  mutate(
    SIC_2DIGIT_NAME_SHORT = fct_reorder(SIC_2DIGIT_NAME_SHORT, employee_percentchange),
    `Firm: ` = paste0(CompanyName,', ',Employees_lastyear,' >> ',Employees_thisyear),
    production = ifelse(SIC_2DIGIT_CODE_NUMERIC < 44, 'production','other'),
    production = factor(production,levels = c('production','other')),
    sizecategory = case_when(
      between(Employees_thisyear,10,20) ~ "10-20",
      between(Employees_thisyear,21,50) ~ "21-50",
      between(Employees_thisyear,51,999999) ~ "51+"
    )
    ) %>% 
  filter(!is.na(SIC_2DIGIT_NAME_SHORT))

# table(firm.change$sizecategory)
 # table(firm.change$production)

#FACET doesn't work well with plotly - things overlap
#Plot them separately.
p <- ggplot(firm.change %>% filter(sizecategory == "10-20"),
            aes(y = SIC_2DIGIT_NAME_SHORT, x = employee_percentchange, group = `Firm: `)) +
  geom_jitter(height = 0.1, alpha = 0.3) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") +
  facet_wrap(~production, ncol = 2, scales = 'free') +
  theme(
    # panel.grid.major = element_line(linetype = "dotted"),
    # axis.title.x = element_text( size = 10, margin=margin(80,80,80,80)),
    axis.title.y = element_text( size = 10, margin=margin(30,30,30,30))
  ) +
  ggtitle("Firms with 10-20 employees")

ggplotly(p, tooltip = 'Firm: ')

p <- ggplot(firm.change %>% filter(sizecategory == "21-50"),
            aes(y = SIC_2DIGIT_NAME_SHORT, x = employee_percentchange, group = `Firm: `)) +
  geom_jitter(height = 0.1, alpha = 0.3) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") +
  facet_wrap(~production, ncol = 2, scales = 'free') +
  theme(
    # panel.grid.major = element_line(linetype = "dotted"),
    # axis.title.x = element_text( size = 10, margin=margin(80,80,80,80)),
    axis.title.y = element_text( size = 10, margin=margin(30,30,30,30))
  ) +
  ggtitle("Firms with 21-50 employees")


ggplotly(p, tooltip = 'Firm: ')

p <- ggplot(firm.change %>% filter(sizecategory == "51+"),
            aes(y = SIC_2DIGIT_NAME_SHORT, x = employee_percentchange, group = `Firm: `)) +
  geom_jitter(height = 0.1, alpha = 0.3) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") +
  facet_wrap(~production, ncol = 2, scales = 'free') +
  theme(
    # panel.grid.major = element_line(linetype = "dotted"),
    # axis.title.x = element_text( size = 10, margin=margin(80,80,80,80)),
    axis.title.y = element_text( size = 10, margin=margin(30,30,30,30))
  ) +
  ggtitle("Firms with 50+ employees")


ggplotly(p, tooltip = 'Firm: ')




```





## GROWTH TRENDS IN GVA, JOBS AND GVA PER JOB

This section examines **growth trends in Bradford** for **GVA and GVA per full time job (a productivity measure)** (both using the 46 sector categories available in the GVA data), as well as **growth in jobs using BRES data** (a larger list of 88 SIC 2 digit sectors).

First, this is done for **the full range of data - 2015 to 2023**. Second, it's **repeated for the more recent time period from 2019 to 2023**, to look for trends changes across the pandemic and more recently.

Here's an explanation of what's in the plots below. 

* Linear slopes are found for each sector in Bradford, and each is compared to (a) all other sectors in Bradford ("Bradford internal", light blue) and (b) the same sector in all other ITL3s zones in the data ("Bradford to other places", dark blue).
* For example - looking at the most obvious stand-out sector in the first plot is **head offices / consultancy**. Its GVA has grown significantly more than around 90% of other sectors in Bradford, and significantly more than around 70% of the head office / consultancy sector in other ITL3 zones. Its productivity (measured in GVA per full time job) has grown more than around 43% of other ITL3 zones.
* The x axis gives the **percent of slopes that are significantly different to others** (i.e. 95% confidence intervals don't overlap).
* The vertical axis lists all sectors in alphabetical order. These labels contain the following information:
    - Green text: slope rises over time - growth is on average going up
    - Red: slope drops over time - growth is on average going down
    - Bold green and red: the up or down slope is significantly different from zero (faded red and green can't be statitically separated from no change over time).
    - Text after the sector name is **percent change per year on average** with 95% confidence intervals.

So:
* To see **whether a sector has been growing or shrinking overall**, use the sector text (green for growth, red for shrinkage, values given for average % change per year)
* To **compare that sector's trend with other sectors in Bradford or the same sector in other places**, use the bars in the plot. Note - these are relative: a slope could be relatively more positive than another, but the sector overall could still be shrinking. For example, ICT has been growing strongly in most places in the UK - but is growth in one place significantly stronger than others? That's what the place and sector slope comparisons try to answer.


### Full data range: 2015 to 2023


```{r, growthdata_cv_andperFT}
#| fig-height: 12 
#| fig-width: 9.5

#Counting sigs. Statisticians having kittens. Let's do it.
itl3.2digit.cv <- read_csv('../data/regionalGVA/regionalGVA_chainedvolume_ITL3_SIC_2DIGIT_LONG_2023.csv') %>% filter(!qg('agri|Owner-occupiers|membership|activities of households',SIC07_description))

#Add in shorter regional GVA SIC text
shortsectornames <- read_csv('../data/shortsectornames_for_regionalGVA_2digitSICs.csv')

#check match, tick
# table(unique(itl3.2digit.cv$SIC07_description) %in% shortsectornames$SIC07_description)
itl3.2digit.cv <- itl3.2digit.cv %>% 
  left_join(
    shortsectornames, by = 'SIC07_description'
  )

#CV first. No NeweyWest
bradford.2dig <- plotSlopeCounts(
  df = itl3.2digit.cv,
  placename = 'Bradford',
  startdate = 2014,
  enddate = 2023,
  date_colname = year,
  region_colname = Region_name,
  sector_colname = SIC07_description_shortened,
  value_colname = value,
  conf_interval = 95
)




#And for GVA per job - for the places we're able to combined data for these (done in miscchecks.R)
bres.gva.2digit.2023 <- readRDS('../data/regionalGVA_plus_BRESjobcounts/regionalGVA_chainedvolume_BRES_FT_jobcount_bespoke2digitSIC_nONLY_MATCHING_GEOGs_2015_2023.rds') %>% 
  filter(JOBCOUNT > 0, !qg('membership|activities of households|agri', SIC07_description)) %>% 
  mutate(gvaperjob = GVA / JOBCOUNT)

#check match, tick
# table(unique(bres.gva.2digit.2023$SIC07_description) %in% shortsectornames$SIC07_description)
bres.gva.2digit.2023 <- bres.gva.2digit.2023 %>% 
  left_join(
    shortsectornames, by = 'SIC07_description'
  )


#CHeck this has same list of sectors as the CV data above... tick
#(With imputed rent removed from CV)
# table(unique(bres.gva.2digit.2023$SIC07_description) %in% itl3.2digit.cv$SIC07_description)
# unique(bres.gva.2digit.2023$SIC07_description)[!unique(bres.gva.2digit.2023$SIC07_description) %in% itl3.2digit.cv$SIC07_description]

bradford.gvaperjob.2dig <- plotSlopeCounts(
  df = bres.gva.2digit.2023,
  placename = 'Bradford',
  startdate = 2015,
  enddate = 2023,
  date_colname = DATE,
  region_colname = Region_name,
  sector_colname = SIC07_description_shortened,
  value_colname = gvaperjob,#Note, this gets log'd in the function, don't do it here
  conf_interval = 95,
  includesectorname_on_axis = T
)

#Check sector match before alterations
# bradford.2dig$plot + bradford.gvaperjob.2dig$plot


#Axes match - remove from RHS
bradford.gvaperjob.2dig$plot <- bradford.gvaperjob.2dig$plot +
  theme(axis.title.y=element_blank(),
        # axis.text.y=element_blank(),
        axis.ticks.y=element_blank(), 
        axis.title.x=element_blank(),
        legend.position = "bottom"
        ) +
        # axis.text.x=element_blank(),
        # axis.ticks.x=element_blank()) +
  guides(fill = F) +
  ggtitle("GVA/FT") 

#And only need the one legend!
bradford.2dig$plot <- bradford.2dig$plot +
  ggtitle("GVA") +
  theme(
    legend.position = "bottom",
    axis.title.y=element_blank()#add this explanation in figure text
  )

bradford.2dig$plot + bradford.gvaperjob.2dig$plot


```




```{r, growthdata_bres2dig}
#| fig-height: 14 

bres <- read_csv("../local/data/BRES/separate_SIC_types_summedfrom5digitSIC/BRES_ALLYEARSWITHDATA_NUTS3_n_ITL321_stacked_2_Fulltimeemployees_2015_2023_SIC_5DIGIT.csv")

#This comes with ALL NUTS and ITL3 data
#Which means 2022 is doubled up
#E.g. see here: 
#bres %>% filter(qg('bradford', GEOGRAPHY_NAME),qg('Museum activities',SIC_5DIGIT_NAME)) %>% View

#Remove one of them...
#Need to filter NOT year == 2022 NOT GEOGRAPHY_CODE_NUTS2018 is NA, keep the rest
#(2023 we need to keep!)
bres <- bres %>% 
  filter(!(DATE == 2022 & is.na(GEOGRAPHY_CODE_NUTS2018)))

#unique(bres$DATE)#tick

SIClookup <- read_csv('../data/SIClookup.csv')

#Join SIC lookup on 5 digit name
#Keep 2 digit and section codes
#May shorten names in a mo...
bres <- bres %>%
  left_join(
    SIClookup %>% select(SIC_5DIGIT_NAME,SIC_2DIGIT_NAME,SIC_SECTION_NAME),
    by = 'SIC_5DIGIT_NAME'
  )

#Make shorter names, use those.
#Make lookup so can be merged in.
names.sections = unique(bres$SIC_SECTION_NAME)
shortnames.sections = reduceSICnames(unique(bres$SIC_SECTION_NAME),'section')

chk.sections = data.frame(names = names.sections, shortnames = shortnames.sections)

names.2digit = unique(bres$SIC_2DIGIT_NAME)
shortnames.2digit = reduceSICnames(unique(bres$SIC_SECTION_NAME),'2 digit')

chk.2digit = data.frame(names = names.2digit, shortnames = shortnames.2digit)


#And 729 categories in the 5 digit...
names.5digit = unique(bres$SIC_5DIGIT_NAME)
#NOTE: THE ORDER HERE IS BASED ON THE ORDER IN bres
#We should make an actual lookup to make sure we don't lose that order
shortnames.5digit = reduceSICnames(unique(bres$SIC_5DIGIT_NAME),'5 digit')

chk.5digit = data.frame(names = names.5digit, shortnames = shortnames.5digit)


#Merge into bres
bres <- bres %>% 
  left_join(chk.sections %>% rename(SIC_SECTION_NAME_SHORT = shortnames), by = c('SIC_SECTION_NAME' = 'names')) %>% 
  left_join(chk.2digit %>% rename(SIC_2DIGIT_NAME_SHORT = shortnames), by = c('SIC_2DIGIT_NAME' = 'names')) %>% 
  left_join(chk.5digit %>% rename(SIC_5DIGIT_NAME_SHORT = shortnames), by = c('SIC_5DIGIT_NAME' = 'names')) 


#Drop some sectors
bres <- bres %>% 
  filter(!qg('activities of households|agri|membership', SIC_2DIGIT_NAME))

#Good!
# table(unique(bres$SIC_2DIGIT_NAME_SHORT) %in% ch.2digitsums$SIC_2DIGIT_NAME_SHORT)


# AND ON WITH LQS 
#OK so getting on with LQs and thinking about what those mean...
#The LQ boundaries won't differ because they're both for Bradford
#And the national sums will be matching quantities across both
bres.2dig <- bres %>% 
  group_by(DATE,GEOGRAPHY_NAME,SIC_2DIGIT_NAME_SHORT) %>% 
  summarise(jobcount = sum(JOBCOUNT)) %>% 
  ungroup()

bradford.bres.2dig <- plotSlopeCounts(
  df = bres.2dig,
  placename = 'Bradford',
  startdate = 2015,
  enddate = 2023,
  date_colname = DATE,
  region_colname = GEOGRAPHY_NAME,
  sector_colname = SIC_2DIGIT_NAME_SHORT,#add shortened version in
  value_colname = jobcount,#Note, this gets log'd in the function, don't do it here
  conf_interval = 95
)

#GVA and BRES
# bradford.2dig2$plot + bradford.bres.2dig2$plot
bradford.bres.2dig$plot <- bradford.bres.2dig$plot +
theme(
  axis.title.y=element_blank(),
  legend.position = "bottom") +
  ggtitle("BRES FT")

bradford.bres.2dig$plot

```





### More recent years: 2019 to 2023

```{r, growthdata_cv_andperFT_later}
#| fig-height: 12 
#| fig-width: 9.5

#Counting sigs. Statisticians having kittens. Let's do it.
itl3.2digit.cv <- read_csv('../data/regionalGVA/regionalGVA_chainedvolume_ITL3_SIC_2DIGIT_LONG_2023.csv') %>% filter(!qg('agri|Owner-occupiers|membership|activities of households',SIC07_description))

#Add in shorter regional GVA SIC text
shortsectornames <- read_csv('../data/shortsectornames_for_regionalGVA_2digitSICs.csv')

#check match, tick
# table(unique(itl3.2digit.cv$SIC07_description) %in% shortsectornames$SIC07_description)
itl3.2digit.cv <- itl3.2digit.cv %>% 
  left_join(
    shortsectornames, by = 'SIC07_description'
  )

#CV first. No NeweyWest
bradford.2dig <- plotSlopeCounts(
  df = itl3.2digit.cv,
  placename = 'Bradford',
  startdate = 2019,
  enddate = 2023,
  date_colname = year,
  region_colname = Region_name,
  sector_colname = SIC07_description_shortened,
  value_colname = value,
  conf_interval = 95
)




#And for GVA per job - for the places we're able to combined data for these (done in miscchecks.R)
bres.gva.2digit.2023 <- readRDS('../data/regionalGVA_plus_BRESjobcounts/regionalGVA_chainedvolume_BRES_FT_jobcount_bespoke2digitSIC_nONLY_MATCHING_GEOGs_2015_2023.rds') %>% 
  filter(JOBCOUNT > 0, !qg('membership|activities of households|agri', SIC07_description)) %>% 
  mutate(gvaperjob = GVA / JOBCOUNT)

#check match, tick
# table(unique(bres.gva.2digit.2023$SIC07_description) %in% shortsectornames$SIC07_description)
bres.gva.2digit.2023 <- bres.gva.2digit.2023 %>% 
  left_join(
    shortsectornames, by = 'SIC07_description'
  )


#CHeck this has same list of sectors as the CV data above... tick
#(With imputed rent removed from CV)
# table(unique(bres.gva.2digit.2023$SIC07_description) %in% itl3.2digit.cv$SIC07_description)
# unique(bres.gva.2digit.2023$SIC07_description)[!unique(bres.gva.2digit.2023$SIC07_description) %in% itl3.2digit.cv$SIC07_description]

bradford.gvaperjob.2dig <- plotSlopeCounts(
  df = bres.gva.2digit.2023,
  placename = 'Bradford',
  startdate = 2019,
  enddate = 2023,
  date_colname = DATE,
  region_colname = Region_name,
  sector_colname = SIC07_description_shortened,
  value_colname = gvaperjob,#Note, this gets log'd in the function, don't do it here
  conf_interval = 95,
  includesectorname_on_axis = T
)

#Check sector match before alterations
# bradford.2dig$plot + bradford.gvaperjob.2dig$plot


#Axes match - remove from RHS
bradford.gvaperjob.2dig$plot <- bradford.gvaperjob.2dig$plot +
  theme(axis.title.y=element_blank(),
        # axis.text.y=element_blank(),
        axis.ticks.y=element_blank(), 
        axis.title.x=element_blank(),
        legend.position = "bottom"
        ) +
        # axis.text.x=element_blank(),
        # axis.ticks.x=element_blank()) +
  guides(fill = F) +
  ggtitle("GVA/FT") 

#And only need the one legend!
bradford.2dig$plot <- bradford.2dig$plot +
  ggtitle("GVA") +
  theme(
    legend.position = "bottom",
    axis.title.y=element_blank()#add this explanation in figure text
  )

bradford.2dig$plot + bradford.gvaperjob.2dig$plot


```




```{r, growthdata_bres2dig_later}
#| fig-height: 14 

bres <- read_csv("../local/data/BRES/separate_SIC_types_summedfrom5digitSIC/BRES_ALLYEARSWITHDATA_NUTS3_n_ITL321_stacked_2_Fulltimeemployees_2015_2023_SIC_5DIGIT.csv")

#This comes with ALL NUTS and ITL3 data
#Which means 2022 is doubled up
#E.g. see here: 
#bres %>% filter(qg('bradford', GEOGRAPHY_NAME),qg('Museum activities',SIC_5DIGIT_NAME)) %>% View

#Remove one of them...
#Need to filter NOT year == 2022 NOT GEOGRAPHY_CODE_NUTS2018 is NA, keep the rest
#(2023 we need to keep!)
bres <- bres %>% 
  filter(!(DATE == 2022 & is.na(GEOGRAPHY_CODE_NUTS2018)))

#unique(bres$DATE)#tick

SIClookup <- read_csv('../data/SIClookup.csv')

#Join SIC lookup on 5 digit name
#Keep 2 digit and section codes
#May shorten names in a mo...
bres <- bres %>%
  left_join(
    SIClookup %>% select(SIC_5DIGIT_NAME,SIC_2DIGIT_NAME,SIC_SECTION_NAME),
    by = 'SIC_5DIGIT_NAME'
  )

#Make shorter names, use those.
#Make lookup so can be merged in.
names.sections = unique(bres$SIC_SECTION_NAME)
shortnames.sections = reduceSICnames(unique(bres$SIC_SECTION_NAME),'section')

chk.sections = data.frame(names = names.sections, shortnames = shortnames.sections)

names.2digit = unique(bres$SIC_2DIGIT_NAME)
shortnames.2digit = reduceSICnames(unique(bres$SIC_SECTION_NAME),'2 digit')

chk.2digit = data.frame(names = names.2digit, shortnames = shortnames.2digit)


#And 729 categories in the 5 digit...
names.5digit = unique(bres$SIC_5DIGIT_NAME)
#NOTE: THE ORDER HERE IS BASED ON THE ORDER IN bres
#We should make an actual lookup to make sure we don't lose that order
shortnames.5digit = reduceSICnames(unique(bres$SIC_5DIGIT_NAME),'5 digit')

chk.5digit = data.frame(names = names.5digit, shortnames = shortnames.5digit)


#Merge into bres
bres <- bres %>% 
  left_join(chk.sections %>% rename(SIC_SECTION_NAME_SHORT = shortnames), by = c('SIC_SECTION_NAME' = 'names')) %>% 
  left_join(chk.2digit %>% rename(SIC_2DIGIT_NAME_SHORT = shortnames), by = c('SIC_2DIGIT_NAME' = 'names')) %>% 
  left_join(chk.5digit %>% rename(SIC_5DIGIT_NAME_SHORT = shortnames), by = c('SIC_5DIGIT_NAME' = 'names')) 


#Drop some sectors
bres <- bres %>% 
  filter(!qg('activities of households|agri|membership', SIC_2DIGIT_NAME))

#Good!
# table(unique(bres$SIC_2DIGIT_NAME_SHORT) %in% ch.2digitsums$SIC_2DIGIT_NAME_SHORT)


# AND ON WITH LQS 
#OK so getting on with LQs and thinking about what those mean...
#The LQ boundaries won't differ because they're both for Bradford
#And the national sums will be matching quantities across both
bres.2dig <- bres %>% 
  group_by(DATE,GEOGRAPHY_NAME,SIC_2DIGIT_NAME_SHORT) %>% 
  summarise(jobcount = sum(JOBCOUNT)) %>% 
  ungroup()

bradford.bres.2dig <- plotSlopeCounts(
  df = bres.2dig,
  placename = 'Bradford',
  startdate = 2019,
  enddate = 2023,
  date_colname = DATE,
  region_colname = GEOGRAPHY_NAME,
  sector_colname = SIC_2DIGIT_NAME_SHORT,#add shortened version in
  value_colname = jobcount,#Note, this gets log'd in the function, don't do it here
  conf_interval = 95
)

#GVA and BRES
# bradford.2dig2$plot + bradford.bres.2dig2$plot
bradford.bres.2dig$plot <- bradford.bres.2dig$plot +
theme(
  axis.title.y=element_blank(),
  legend.position = "bottom") +
  ggtitle("BRES FT")

bradford.bres.2dig$plot

```


# UK INDUSTRIAL STRATEGY LINKAGE

The UK Industrial Strategy includes a [sector definition list](https://www.gov.uk/government/publications/industrial-strategy/industrial-strategy-sector-definitions-list) that lists sectors (of differing specificity / SIC code number) that the government considers part of its target sectors. As that page says, most of these SIC codes cannot capture the full range of possible firms the strategy aims to cover, and it also contains entire sections that SIC codes can't cover at all - clean energy and certain digital sectors in particular.

However, it does allow an analysis of which sectors are *definitely* within those industrial strategy categories. This section gathers job details for those specific sectors - a distinct range of types and breadth - and compares Bradford to the rest of the UK where possible, to see which sectors align with the industrial strategy and other analysis above. This is done for both BRES and Companies House job counts, to pick up on different job types.

The industrial strategy drills down below its top level sectors into 'frontier' sectors, usually (but not always) using more granular SIC codes. The plots below break the data down into these frontier and other sectors using location quotients. 

LQs are useful here as they show whether a sector has a stronger presence in a place regardlesss of the sector size or SIC digit level.

Only industrial strategy sectors with 100+ jobs in Bradford are shown (average from 2021 to 2023 for BRES, and in the latest accounts for Companies House).

Note also: job percents for Bradford show **percent of *just* total industrial strategy jobs**, not percent of all jobs.


## "Frontier" SIC sectors 

### BRES frontier sectors

```{r, bresindstrat}
#| fig-width: 9

place = "Bradford"

#Blue peter this up
#Done here: https://github.com/DanOlner/RegionalEconomicTools/blob/1d5df72210586d0570cd687534bcfb3c98827ece/prepcode/misc_checks.R#L1991
#save a copy of that to save processing time in quarto
yeartoplot.lqfiltered <- readRDS('../local/data/bres_indstrat_LQs.rds')
#yeartoplot.lqfiltered <- readRDS('local/data/bres_indstrat_LQs.rds')

#If I could plot both and space them out, that would be good (could get Bradford change showing too)
p <- LQ_baseplot(df = yeartoplot.lqfiltered %>% filter(is_frontier == 1), alpha = 0.1, shape = 0, sector_name = sic_namefrom_indstrat_combo, LQ_column = LQ, change_over_time = slope)

p <- addplacename_to_LQplot(df = yeartoplot.lqfiltered %>% filter(is_frontier == 1), plot_to_addto = p, 
                            placename = place, shapenumber = 16,
                            min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                            value_column = jobcount_movingav, sector_regional_proportion = sector_regional_proportion,
                            region_name = GEOGRAPHY_NAME,
                            sector_name = sic_namefrom_indstrat_combo, change_over_time = slope, LQ_column = LQ,
                            text = 7, value_col_ismoney = F)

p <- p + 
  # coord_cartesian(xlim = c(0.1,7)) +
  ggtitle("IndStrat frontier sector")

p

```

### Companies house frontier sectors


```{r, chindstrat}
#| fig-width: 9

#Blue peter this up
#Done here: https://github.com/DanOlner/RegionalEconomicTools/blob/1d5df72210586d0570cd687534bcfb3c98827ece/prepcode/misc_checks.R#L1991
#save a copy of that to save processing time in quarto
yeartoplot.lqfiltered.ch <- readRDS('../local/data/companieshouse_indstrat_LQs.rds')

p1 <- LQ_baseplot(df = yeartoplot.lqfiltered.ch %>% filter(is_frontier == 1), alpha = 0.03, sector_name = sic_namefrom_indstrat_combo, 
                  LQ_column = LQ, change_over_time = slope)

# debugonce(addplacename_to_LQplot)
p1 <- addplacename_to_LQplot(df = yeartoplot.lqfiltered.ch %>% filter(is_frontier == 1), plot_to_addto = p1, 
                             placename = place, shapenumber = 16,
                             # min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                             value_column = total_jobs, sector_regional_proportion = sector_regional_proportion,
                             region_name = localauthority_name,
                             sector_name = sic_namefrom_indstrat_combo, change_over_time = slope, LQ_column = LQ,
                             value_col_ismoney = F, text = 7)
                             # value_col_ismoney = F, text = 7, maxLQvalmultiplier = 2,useplacenameforminmaxdisplay = T, overridetextpos = 14)

p1 <- p1 + 
  # coord_cartesian(xlim = c(0.1,7)) +
  ggtitle("IndStrat frontier sector")

p1


```

## Other industrial strategy SIC sectors 

### BRES other sectors

```{r, bresindstrat2}
#| fig-width: 9

p <- LQ_baseplot(df = yeartoplot.lqfiltered %>% filter(is_frontier == 0), alpha = 0.1, shape = 0, sector_name = sic_namefrom_indstrat_combo, LQ_column = LQ, change_over_time = slope)

p <- addplacename_to_LQplot(df = yeartoplot.lqfiltered %>% filter(is_frontier == 0), plot_to_addto = p, 
                            placename = place, shapenumber = 16,
                            min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                            value_column = jobcount_movingav, sector_regional_proportion = sector_regional_proportion,
                            region_name = GEOGRAPHY_NAME,
                            sector_name = sic_namefrom_indstrat_combo, change_over_time = slope, LQ_column = LQ,
                            text = 7, value_col_ismoney = F)

p <- p + 
  # coord_cartesian(xlim = c(0.1,7)) +
  ggtitle("IndStrat other sector")

p


```

### Companies House other sectors


```{r,chindstrat2}
#| fig-width: 9
 
p2 <- LQ_baseplot(df = yeartoplot.lqfiltered.ch %>% filter(is_frontier == 0), alpha = 0.03, sector_name = sic_namefrom_indstrat_combo, 
                  LQ_column = LQ, change_over_time = slope)

# debugonce(addplacename_to_LQplot)
p2 <- addplacename_to_LQplot(df = yeartoplot.lqfiltered.ch %>% filter(is_frontier == 0), plot_to_addto = p2, 
                             placename = place, shapenumber = 16,
                             # min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                             value_column = total_jobs, sector_regional_proportion = sector_regional_proportion,
                             region_name = localauthority_name,
                             sector_name = sic_namefrom_indstrat_combo, change_over_time = slope, LQ_column = LQ,
                             value_col_ismoney = F, text = 7)
                             # value_col_ismoney = F, text = 7, maxLQvalmultiplier = 2,useplacenameforminmaxdisplay = T, overridetextpos = 14)

p2 <- p2 + 
  # coord_cartesian(xlim = c(0.1,7)) +
  ggtitle("IndStrat other sector")

p2



```



