---
title: "Bradford sectors and clusters: past, present"
author:
  - name: "Dan Olner (Yorkshire & Humber Policy Engagement & Research Network, University of Sheffield)"
format: 
  html:
    output-file: "Bradford_sectorclusters_July2025"
    fig-width: 8
    fig-height: 8
execute:
  echo: false
  message: false
  warning: false
fig-responsive: false
collapse: false
---

```{r setup}
library(tidyverse)
# library(spdep)
library(sf)#for geo stuff
# library(tmap)#for mapping 
library(zoo)#for making smoothed moving averages
library(patchwork)#combine ggplots easily
library(plotly)#For interactive plots
# # library(ggdist)
# library(tidyr)
# library(distributional)
library(RColorBrewer)

source('../functions/misc_functions.R')
source('../functions/adhoc_functions.R')
#source('functions/misc_functions.R')
#source('functions/adhoc_functions.R')

options(scipen = 999)

#Set ggplot theme
theme_set(theme_light())


```

```{r gvaperhour_dataprep}
# ITL3 Hours % change v GVA % change on average----

#So need the number of hours and chained volume GVA
#Hours count is in the per hour productivity sheet
#GVA in the region by industry sheet

#hours worked first
url1 <- 'https://www.ons.gov.uk/file?uri=/employmentandlabourmarket/peopleinwork/labourproductivity/datasets/subregionalproductivitylabourproductivitygvaperhourworkedandgvaperfilledjobindicesbyuknuts2andnuts3subregions/current/labourproductivityitls1.xlsx'
p1f <- tempfile(fileext=".xlsx")
download.file(url1, p1f, mode="wb")

#"Productivity Hours Worked per Week; ITL2 and ITL3 subregions (constrained to ITL1), 2004 - 2023"
hoursworked <- readxl::read_excel(path = p1f,range = "Productivity Hours!A5:W247")

names(hoursworked) <- gsub(x = names(hoursworked), pattern = ' ', replacement = '_')

hoursworked = hoursworked %>%
  filter(ITL_level == 'ITL3') %>% 
  pivot_longer(Hours_2004:Hours_2023, names_to = 'year', values_to = 'hoursworkedperweek') %>% #get most recent year
  mutate(
    year = substr(year,7,10),
    year = as.numeric(year)
  ) %>% 
  select(-ITL_level)



#Then ITL3 CV for all industries (so has to include imputed rent as we can't sum CV)
url2 <- 'https://www.ons.gov.uk/file?uri=/economy/grossvalueaddedgva/datasets/nominalandrealregionalgrossvalueaddedbalancedbyindustry/current/regionalgrossvalueaddedbalancedbyindustryandallinternationalterritoriallevelsitlregions.xlsx'
p2f <- tempfile(fileext=".xlsx")
download.file(url2, p2f, mode="wb") 

#"Table 3b: ITL3, chained volume measures in 2019 money value, pounds million [note 2]"
itl3.all <- readxl::read_excel(path = p2f, range = "Table 3b!A2:AD11468")

names(itl3.all) <- gsub(x = names(itl3.all), pattern = ' ', replacement = '_')

itl3.all = itl3.all %>% 
  filter(SIC07_description == 'All industries') %>% 
  pivot_longer(`1998`:names(itl3.all)[length(names(itl3.all))], names_to = 'year', values_to = 'value') %>% #get most recent year
  mutate(year = as.numeric(year)) %>% 
  rename(gva_cv = value)



#Check link between two is valid... tick
# table(unique(hoursworked$ITL_code) %in% itl3.all$ITL_code)

#2004:2023
# unique(hoursworked$year)[unique(hoursworked$year) %in% unique(itl3.all$year)]

#Get list of corecities
corecities <- getdistinct('sheffield|Belfast|Birmingham|Bristol|Cardiff|Glasgow|Leeds|Liverpool|Manchester|Tyne|Nottingham', itl3.all$Region_name)

#Yes, all in there, need to remove a few...
corecities <- corecities[!grepl(x = corecities, pattern = 'Greater|shire', ignore.case = T)]
corecities <- corecities[order(corecities)]

#Check match also works in hours worked... tick
# table(corecities %in% hoursworked$Region_name)



#Get London ITL3s
itl_lookup = read_csv("../data/LAD_(December_2024)_to_LAU1_to_ITL3_to_ITL2_to_ITL1_(January_2025)_Lookup_in_the_UK.csv")
#itl_lookup = read_csv("data/LAD_(December_2024)_to_LAU1_to_ITL3_to_ITL2_to_ITL1_(January_2025)_Lookup_in_the_UK.csv")

#Get all London ITL2 zones then use to get unique ITL3 zones
itl3.lon = itl_lookup %>%
  filter(qg('london',ITL225NM)) %>% 
  select(ITL325CD,ITL325NM) %>% 
  distinct()

#Check itl3 code match... tick
# table(itl3.lon$ITL325CD %in% hoursworked$ITL_code)
# table(itl3.lon$ITL325CD %in% itl3.all$ITL_code)
#tick
# table(itl3.lon$ITL325NM %in% hoursworked$Region_name)
# table(itl3.lon$ITL325NM %in% itl3.all$Region_name)


```






# ONS GVA AND JOBS DATA

## Structural /  broad sector GVA change 

```{r struct}
#| fig-width: 8
#| fig-height: 16

# LOAD ITL3 GVA DATA----

#ITL3 zones, SIC sections, current prices, 2023 latest data (minus imputed rent)
df = read_csv('https://bit.ly/rtasteritl3noir')

#This one has imputed rent included if you want it
# df = read_csv('https://bit.ly/rtasteritl3')



#Finding location quotients for each year in the data 
#And then overwriting the df variable with the result
df = df %>%
  group_split(year) %>%
  map(add_location_quotient_and_proportions,
      regionvar = Region_name,
      lq_var = SIC07_description,
      valuevar = value) %>% 
  bind_rows()


#1. Set window for rolling average: 3 years
smoothband = 3

#2. Make new column with rolling average in (applying to all places and sectors)
#Years are already in the right order, so will roll apply across those correctly
df = df %>% 
  group_by(Region_name,SIC07_description) %>% 
  mutate(
    sector_regional_prop_movingav = rollapply(
      sector_regional_proportion,smoothband,mean,align='center',fill=NA
      )
  ) %>% 
  ungroup()


#Get set pastel colours for each SIC section
n <- length(unique(df$SIC07_description))
set.seed(12)
qual_col_pals = RColorBrewer::brewer.pal.info[RColorBrewer::brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(RColorBrewer::brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

randomcols <- col_vector[10:(10+(n-1))]

#Do once!
getplot <- function(place){
  
  df = df %>% 
    filter(!qg("Activities of households|agri",SIC07_description)) %>% 
    mutate(
      SIC07_description = reduceSICnames(names = SIC07_description, level = 'section')
    )
  
  top12sectors = df %>%
    filter(
      Region_name == place,
      year == max(year) - 1#to get 3 year moving average value
    ) %>%
    arrange(-sector_regional_prop_movingav) %>%
    slice_head(n= 12) %>%
    select(SIC07_description) %>%
    pull()
  
  
  ggplot(df %>% 
           filter(
             Region_name == place, 
             SIC07_description %in% top12sectors,
             !is.na(sector_regional_prop_movingav)#Remove NAs created by 3 year moving average
           ), 
         aes(x = year, y = sector_regional_prop_movingav * 100, colour = fct_reorder(SIC07_description,-sector_regional_prop_movingav) )) +
    geom_point() +
    geom_line() +
    scale_color_manual(values = setNames(randomcols,unique(df$SIC07_description))) +#set manual pastel colours matching sector name
    ylab('Sector: percent of regional economy') +
    theme(plot.title = element_text(face = 'bold')) +
    labs(colour = 'SIC section') +
    # scale_y_log10() +
    ggtitle(paste0(place, ' GVA % of economy over time\nTop 12 sectors in ',max(df$year),'\n (', smoothband,' year moving average)'))
  
  
}

places = getdistinct('bradford|kirklees|calderdale|leeds|wakefield',df$Region_name)

plots <- map(places, getplot)

patchwork::wrap_plots(plots, ncol = 1)


```




## GVA per full time job over time: Bradford v other ITL3 zones for 2 digit SIC sectors

These plots use **linked regional GVA and BRES jobs data** to examine productivity, output and job changes. To make this link, sector categories have to be reduced to those in the ONS GVA data (46 categories rather than the full 88 two digit SICs).

BRES does contain more detail - more on that below and in the separate [BRES 5 digit breakdown](https://danolner.github.io/RegionalEconomicTools/quarto_docs/Bradford_BRES_5digit_LQs_July2025.html){target="_blank"}.

The plots below are measured in **GVA per job parts per million of UK economy** - equivalent to e.g. 1% being "1 in 100 parts of economy", 0.1 here is "1 in 10 millionth of whole economy". So it's again examining **relative change in output per FT worker** across all sectors.

GVA per job as a measure needs some caution: output is also due to other productive factors, so high capital intensity sectors (power, telecoms, other ICT though not all) can appear overly productive per worker. It's more representative for service sectors where an employee's output more clearly correlates to total output, such as legal services.


```{r gvajobs1}
#| fig-height: 20
#| fig-width: 8.5

#2015 to 2023, linked on the regional GVA bespoke SIC categories
#Note, can use lookup if need to also sum for sections (which we can do here cos current prices)
gvabres <- readRDS('../data/regionalGVA_plus_BRESjobcounts/regionalGVA_currentprices_BRES_FT_jobcount_bespoke2digitSIC_nONLY_MATCHING_GEOGs_2015_2023.rds') 
#gvabres <- readRDS('data/regionalGVA_plus_BRESjobcounts/regionalGVA_currentprices_BRES_FT_jobcount_bespoke2digitSIC_nONLY_MATCHING_GEOGs_2015_2023.rds') 

#Make abbreviated sector names
shortsectornames = tibble(
  SIC07_description = unique(gvabres$SIC07_description),
  SIC07_description_shortened = c(
  "Furniture",
  "Machinery repair/install",
  "Buildings",
  "Civil eng",
  "Spec constr",
  "Motor",
  "Wholesale",
  "Retail",
  "Transport",
  "Warehousing",
  "Post",
  "Accommodation",
  "Food",
  "Media",
  "ICT",
  "Real estate",
  "Legal",
  "Head office/consult",
  "Arch/Eng",
  "Research",
  "Other PSTech",#"Prof/sci/tech",
  "Rental",
  "Employment",
  "Cleaning/landscapes",
  "Admin",
  "Health",
  "Care",
  "SocialWork",
  "Arts/ent",
  "Recreation",
  "Membership",
  "Repair personal/computer",
  "Personal",
  "Agri/mining",
  "Food manuf",
  "Textiles",
  "Wood/paper",
  "Chemicals",
  "Metals",
  "Electronics",
  "Mach/transp manuf",
  "Utilities",
  "Finance",
  "Public admin",
  "Education",
  "Households"
  )
)

#save that for use elsewhere!
# write_csv(shortsectornames, 'data/shortsectornames_for_regionalGVA_2digitSICs.csv')

gvabres <- gvabres %>% 
  left_join(
    shortsectornames, by = 'SIC07_description'
  )%>% 
  filter(!qg('households|membership', SIC07_description_shortened))

#Would also quite like to split manufacturing off
#Check in orig GVA file for the codes...
#This is the full list - some won't match but enough to label the correct ones at 2 digit for those that do
productionsectors <- c(
  'A-E',
  'AB (1-9)',
  'C (10-33)',
  'CA (10-12)',
  'CB (13-15)',
  'CC (16-18)',
  'CD-CG (19-23)',
  'CH (24-25)',
  'CI-CJ (26-27)',
  'CK-CL (28-30)',
  'CM (31-33)',
  '31-32',
  '33',
  'DE (35-39)',
  'F (41-43)',
  '41',
  '42',
  '43'
)

#Yep
# table(productionsectors %in% gvabres$SIC_2DIGIT_CODE_GVA2023)
# unique(gvabres$SIC_2DIGIT_CODE_GVA2023)[unique(gvabres$SIC_2DIGIT_CODE_GVA2023) %in% productionsectors]

#Label production sectors
gvabres <- gvabres %>% 
  mutate(
    productionsector = ifelse(
      SIC_2DIGIT_CODE_GVA2023 %in% productionsectors,
      'production','other'
    )
  )



#For current prices - what we can do:
#Find GVA as a proportion of total UK output for those years
#Then that proportion per FT job
#To see how productivity has shifted about internally / how sectors have changed relative to each other

#That version is missing some places so isn't entire UK
#Let's get entire UK's GVA per year
#Can just sum from any current price source
#2023:
allgva <- read_csv('../data/regionalGVA/regionalGVA_currentprices_ITL2_SIC_2DIGIT_LONG_2023.csv')
#allgva <- read_csv('data/regionalGVA/regionalGVA_currentprices_ITL2_SIC_2DIGIT_LONG_2023.csv')




allgva <- allgva %>% 
  group_by(year) %>% 
  summarise(totalUKGVA_CP = sum(value)) %>% 
  ungroup()

#Add that in...
gvabres <- gvabres %>% 
  left_join(allgva, by = c('DATE'= 'year')
  )

#Some NAs in jobs - I think the issue with the missing values is just early years with no agri data...
#filter
gvabres <- gvabres %>% filter(!is.na(JOBCOUNT))


#Find gva proportion of UK total then the job's proportion of that
gvabres <- gvabres %>%
  mutate(
    gva_proportionofUKtotal = GVA / totalUKGVA_CP,
    job_output_propofUKtotal = gva_proportionofUKtotal / JOBCOUNT
  )


#Get regional GVA data version of SIC sections 
#To plot each by
# siclookupfromreg.gva <- read_csv('../data/SIClookup_forBRES_REGIONALGVA_2023_JOIN.csv')
#siclookupfromreg.gva <- read_csv('data/SIClookup_forBRES_REGIONALGVA_2023_JOIN.csv')



smoothband = 3

gvabres <- gvabres %>%
  group_by(Region_name,SIC07_description) %>%
  mutate(
    jobcount_movingav = rollapply(JOBCOUNT,smoothband,mean,align='center',fill=NA),
    gva_movingav = rollapply(GVA,smoothband,mean,align='center',fill=NA),
    job_output_propofUKtotal_movingav = rollapply(job_output_propofUKtotal,smoothband,mean,align='center',fill=NA),
    job_PPM_ofUKtotal_movingav = job_output_propofUKtotal_movingav * 1000000#1% is 1 part per 100. Which we find by multiplying up the prop by 100. PPM is just x * a million.
  ) %>% 
  ungroup()



#ggplot won't give control over order very well... let's add on top manually, more control
prod.data <- gvabres %>% 
  mutate(
    DATE = DATE - 2000,#Make dates 2 digit, more readable on axis
    SIC07_description_shortened = fct_reorder(SIC07_description_shortened, job_PPM_ofUKtotal_movingav, .desc = T),
    highlow = ifelse(SIC07_description_shortened %in% levels(SIC07_description_shortened)[1:23],'high','low'),#Break into two plots to stack, split by total av productivity
    sectorgrouping = cut_number(as.integer(SIC07_description_shortened), 8) %>% as.integer
  ) %>% 
  # filter(!qg('hous',SIC07_description))
  filter(!qg('hous',SIC07_description), !is.na(job_output_propofUKtotal_movingav))

#Break into n groups, using factor order created in previous line
# sectorgrouping = ntile(as.integer(prod.data$SIC07_description_shortened), 5)

  
#table(prod.data$sectorgroupings)

#Test
# debugonce(plotprod)
# plotprod(prod.data %>% filter(sectorgrouping==6))

plotz <- map(prod.data %>% group_split(sectorgrouping),plotprod)

patchwork::wrap_plots(plotz,ncol = 1)

#Quick look
#gvabres %>% filter(qg('architecture',SIC07_description_shortened), qg('bradford',Region_name)) %>% View


```



## Location quotients for GVA and jobs in 2 digit SIC sectors: most recent and change over time

These LQ plots show location quotients for linked GVA and jobs data for Bradford, overlaid on top of LQs for all other ITL3 zones, to give a sense of what the spread is across Great Britain.

A location quotient is the ratio of a place's sector proportion to its national proportion. So e.g. if manufacturing in Bradford made up 20% of its economy compared to maufacturing being 10% of the UK economy, its LQ in Bradford would be 2.

Here’s an explanation of how to read these location quotient plots:

* Location quotients for sectoral GVA in the selected region
* Sectors to the right of the blue line are more concentrated in this region when compared to the national average. (e.g. LQ = 10, it is 10x more concentrated). To the left are less concentrated than nationally.
* Green circles have been increasing in concentration over time (larger circles increasing the most)
* Red circles have been decreasing in concentration over time (larger circles decreasing the most)
* Black horizontal lines show the min and max of concentration in the data’s history (from 2015 to 2023) (so, often, large green circles will be at the right of this line because sector concentration has been steadily increasing)
* Per sector GVA money amounts and percent of that region’s entire economy are shown on the right.

### Production sectors


```{r productionLQs}
#| fig-width: 9
#| fig-height: 10

#Get two versions of LQ/proportions - job count and GVA
gva.props <- gvabres %>% 
  group_split(DATE) %>%
  map(add_location_quotient_and_proportions,
      regionvar = Region_name,
      lq_var = SIC07_description_shortened,
      valuevar = GVA) %>% 
  bind_rows()

job.props <- gvabres %>% 
  group_split(DATE) %>%
  map(add_location_quotient_and_proportions,
      regionvar = Region_name,
      lq_var = SIC07_description_shortened,
      valuevar = JOBCOUNT) %>% 
  bind_rows()

# LET'S SEPARATE GVA AND JOBS OUT FOR BRADFORD AND COMPARE TO EVERYWHERE ELSE TOO
place = "Bradford"

#FUNCTION
#Split both gva.props and job.props by subsector list
makegvajoblqplots <- function(keepthesesectors){
  
  # gva.sub <- gva.props %>% filter(SIC07_description_shortened %in% keepthesesectors)

  LQ_slopes <- compute_slope_or_zero(
    data = gva.props, 
    Region_name, SIC07_description_shortened,#slopes will be found within whatever grouping vars are added here
    y = LQ_log, x = DATE)
  
  
  #Filter down to a single year... we may want to smooth years, let's see
  yeartoplot <- gva.props %>% filter(DATE == max(DATE))#use latest year
  
  #Add slopes into data to get LQ plots
  yeartoplot <- yeartoplot %>% 
    left_join(
      LQ_slopes,
      by = c('Region_name', 'SIC07_description_shortened')
    )
  
  #Get min/max values for LQ over time as well, for each sector and place, to add as bars so range of sector is easy to see
  minmaxes <- gva.props %>% 
    group_by(Region_name, SIC07_description_shortened) %>% 
    summarise(
      min_LQ_all_time = min(LQ),
      max_LQ_all_time = max(LQ)
    )
  
  #Join min and max
  yeartoplot <- yeartoplot %>% 
    left_join(
      minmaxes,
      by = c('Region_name', 'SIC07_description_shortened')
    )
  
  
  #Shorten here to subset of sectors
  yeartoplot <- yeartoplot %>% 
    filter(SIC07_description_shortened %in% keepthesesectors)
  
  
  sectorLQorder <- gva.props %>% filter(
    Region_name == place,
    DATE == max(DATE)#use latest data
  ) %>% 
    arrange(-LQ) %>% 
    select(SIC07_description_shortened) %>% 
    pull()
  
  
  #Turn the sector column into a factor and order by LCR's LQs
  yeartoplot$SIC07_description_shortened <- factor(yeartoplot$SIC07_description_shortened, levels = sectorLQorder, ordered = T)
  
  #If I could plot both and space them out, that would be good (could get Bradford change showing too)
  p <- LQ_baseplot(df = yeartoplot, alpha = 0.1, shape = 0, sector_name = SIC07_description_shortened, 
                   LQ_column = LQ, change_over_time = slope)
  
  p <- addplacename_to_LQplot(df = yeartoplot, plot_to_addto = p, 
                              placename = place, shapenumber = 16,
                              min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                              value_column = GVA, sector_regional_proportion = sector_regional_proportion,
                              region_name = Region_name,
                              sector_name = SIC07_description_shortened, change_over_time = slope, LQ_column = LQ,
                              text = 7)
  
  p <- p + 
    # coord_cartesian(xlim = c(0.1,7)) +
    ggtitle("GVA")
  
  
  
  
  # job.sub <- job.props %>% filter(SIC07_description_shortened %in% keepthesesectors)
  
  # REPEAT FOR JOBS / BRADFORD (Using same factor order?)
  LQ_slopes <- compute_slope_or_zero(
    data = job.props, 
    Region_name, SIC07_description_shortened,#slopes will be found within whatever grouping vars are added here
    y = LQ_log, x = DATE)
  
  
  #Filter down to a single year... we may want to smooth years, let's see
  yeartoplot <- job.props %>% filter(DATE == max(DATE))#use latest year
  
  #Add slopes into data to get LQ plots
  yeartoplot <- yeartoplot %>% 
    left_join(
      LQ_slopes,
      by = c('Region_name', 'SIC07_description_shortened')
    )
  
  #Get min/max values for LQ over time as well, for each sector and place, to add as bars so range of sector is easy to see
  minmaxes <- job.props %>% 
    group_by(Region_name, SIC07_description_shortened) %>% 
    summarise(
      min_LQ_all_time = min(LQ),
      max_LQ_all_time = max(LQ)
    )
  
  #Join min and max
  yeartoplot <- yeartoplot %>% 
    left_join(
      minmaxes,
      by = c('Region_name', 'SIC07_description_shortened')
    )
  
  #Shorten here to subset of sectors
  yeartoplot <- yeartoplot %>% 
    filter(SIC07_description_shortened %in% keepthesesectors)
  
  
  
  #USE SECTOR ORDER FROM GVA LQS ABOVE
  yeartoplot$SIC07_description_shortened <- factor(yeartoplot$SIC07_description_shortened, levels = sectorLQorder, ordered = T)
  
  #If I could plot both and space them out, that would be good (could get Bradford change showing too)
  p2 <- LQ_baseplot(df = yeartoplot, alpha = 0.1, shape = 0, sector_name = SIC07_description_shortened, 
                   LQ_column = LQ, change_over_time = slope)
  
  p2 <- addplacename_to_LQplot(df = yeartoplot, plot_to_addto = p2, 
                              placename = place, shapenumber = 16,
                              min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                              value_column = JOBCOUNT, sector_regional_proportion = sector_regional_proportion,
                              region_name = Region_name,
                              sector_name = SIC07_description_shortened, change_over_time = slope, LQ_column = LQ,
                              text = 7, value_col_ismoney = F)
  
  p2 <- p2 + 
    # coord_cartesian(xlim = c(0.1,7)) +
    ggtitle("Jobs")
  
  # p2 <- p2 + scale_y_discrete(position = "right")
  p2 <- p2 +
    theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
  
  p + p2 

}

#production sector plot first
makegvajoblqplots(gva.props$SIC07_description_shortened[gva.props$productionsector == 'production'])


```


### Rest of sectors

```{r otherLQs}
#| fig-width: 9
#| fig-height: 14

makegvajoblqplots(gva.props$SIC07_description_shortened[gva.props$productionsector == 'other'])

```




## Av. percent change over time for gva vs hours worked, ITL3 zones

**Hover for place names**.

* West Yorkshire ITL3s in green - Bradford's outlined in black (95% CIs in lines).
* Core cities in blue  (95% CIs)
* London ITL3s in red/pink

And:

* Anything above/left of diagonal line is a **productivity increase**. 




### 2004 to 2013 av % change

```{r plot1}
# startyear = 2014
# endyear = 2023

#So e.g. in the plot below for 2014-2023, Rotherham's productivity increased with a **GVA increase** and a **number of hours worked *decrease***.

startyear = 2004
endyear = 2013

#Hours per week first
slopes.hours = get_slope_and_se_safely(
  hoursworked %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(hoursworkedperweek), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.hours = slopes.hours %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )


#Mark 4 SY LAs and plot
slopes.hours = slopes.hours %>% 
  ungroup() %>% 
  # mutate(
  #   SY_LAs = ifelse(qg('rotherh|doncaste|sheffie|barnsley', Region_name),'SY LA','Other'),
  #   corecities = ifelse(Region_name %in% corecities,'core city','Other'),
  #   london = ifelse(qg('london',Region_name), 'London','Other')
  # )
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'SY LA',#will overwrite sheffield as core city
      qg('bradford|kirklees|calderdale|leeds|wakefield', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
      ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','WY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.hours = slopes.hours %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )




#Then GVA CV
slopes.gva = get_slope_and_se_safely(
  itl3.all %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(gva_cv), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.gva = slopes.gva %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )



#Mark 4 WY LAs and plot
slopes.gva = slopes.gva %>% 
  ungroup() %>% 
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      qg('bradford|kirklees|calderdale|leeds|wakefield', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
    ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','WY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.gva = slopes.gva %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )


# table(slopes.hours$itl3label)
# table(slopes.gva$itl3label)


#Keep the bits we want and combine into single df
both <- slopes.gva %>% 
  select(Region_name,itl3label,
         slope_percentperyear_GVA = slope_percentperyear,
         ci95min_percentperyear_GVA = ci95min_percentperyear,
         ci95max_percentperyear_GVA = ci95max_percentperyear) %>% 
  inner_join(
    slopes.hours %>% 
      select(Region_name,
             slope_percentperyear_HOURS = slope_percentperyear,
             ci95min_percentperyear_HOURS = ci95min_percentperyear,
             ci95max_percentperyear_HOURS = ci95max_percentperyear),
    by = 'Region_name'
  )


#Poor control over visibility. Layer instead. Base plot then function.
addtogva_hours_plot = function(p,plotdata, showerrors = T, sizeval = 1, colourval = 'black', alphaval = 1){
  
  p = p + geom_point(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, group = Region_name), colour = colourval, size = sizeval, alpha = alphaval) 
  
  if(showerrors){
    
    p = p + geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, xmin = ci95min_percentperyear_HOURS, xmax = ci95max_percentperyear_HOURS),width = 0.1, size = 1, colour = colourval, alpha = alphaval) +
    geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, ymin = ci95min_percentperyear_GVA, ymax = ci95max_percentperyear_GVA),width = 0.1, size = 1, colour = colourval, alpha = alphaval) 
    
  }
  
    p  + geom_abline(slope = 1, intercept = 0, colour = 'black', alpha = 0.5, size = 2) +
      geom_hline(yintercept = 0, colour = 'black', alpha = 0.3, size = 1) +
      geom_vline(xintercept = 0, colour = 'black', alpha = 0.3, size = 1)
  
}




p = ggplot()

p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Other'), F, 1, 'black', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'London'), F, 2, 'red', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Core city'), T, 3, RColorBrewer::brewer.pal(3, "Set1")[2], 0.5)
p = addtogva_hours_plot(p, both %>% filter(Region_name == 'Bradford'), F, 10, 'black', 1)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'WY LA'), T, 7, RColorBrewer::brewer.pal(3, "Set1")[3], 1)

p = p + coord_fixed() +
  xlab("Average % change: hours worked per week") +
  ylab("Average % change: GVA") 


ggplotly(p, tooltip = 'Region_name')



```

### 2014 to 2023 av % change

```{r plot2}
startyear = 2014
endyear = 2023

# startyear = 2004
# endyear = 2013

#Hours per week first
slopes.hours = get_slope_and_se_safely(
  hoursworked %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(hoursworkedperweek), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.hours = slopes.hours %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )


#Mark 4 WY LAs and plot
slopes.hours = slopes.hours %>% 
  ungroup() %>% 
  # mutate(
  #   SY_LAs = ifelse(qg('rotherh|doncaste|sheffie|barnsley', Region_name),'WY LA','Other'),
  #   corecities = ifelse(Region_name %in% corecities,'core city','Other'),
  #   london = ifelse(qg('london',Region_name), 'London','Other')
  # )
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      qg('bradford|kirklees|calderdale|leeds|wakefield', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
      ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','WY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.hours = slopes.hours %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )




#Then GVA CV
slopes.gva = get_slope_and_se_safely(
  itl3.all %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(gva_cv), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.gva = slopes.gva %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )



#Mark 4 WY LAs and plot
slopes.gva = slopes.gva %>% 
  ungroup() %>% 
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      qg('bradford|kirklees|calderdale|leeds|wakefield', Region_name) ~ 'WY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
    ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','WY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.gva = slopes.gva %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )


# table(slopes.hours$itl3label)
# table(slopes.gva$itl3label)


#Keep the bits we want and combine into single df
both <- slopes.gva %>% 
  select(Region_name,itl3label,
         slope_percentperyear_GVA = slope_percentperyear,
         ci95min_percentperyear_GVA = ci95min_percentperyear,
         ci95max_percentperyear_GVA = ci95max_percentperyear) %>% 
  inner_join(
    slopes.hours %>% 
      select(Region_name,
             slope_percentperyear_HOURS = slope_percentperyear,
             ci95min_percentperyear_HOURS = ci95min_percentperyear,
             ci95max_percentperyear_HOURS = ci95max_percentperyear),
    by = 'Region_name'
  )


#Poor control over visibility. Layer instead. Base plot then function.
addtogva_hours_plot = function(p,plotdata, showerrors = T, sizeval = 1, colourval = 'black', alphaval = 1){
  
  p = p + geom_point(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, group = Region_name), colour = colourval, size = sizeval, alpha = alphaval) 
  
  if(showerrors){
    
    p = p + geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, xmin = ci95min_percentperyear_HOURS, xmax = ci95max_percentperyear_HOURS),width = 0.1, size = 1, colour = colourval, alpha = alphaval) +
    geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, ymin = ci95min_percentperyear_GVA, ymax = ci95max_percentperyear_GVA),width = 0.1, size = 1, colour = colourval, alpha = alphaval) 
    
  }
  
    p  + geom_abline(slope = 1, intercept = 0, colour = 'black', alpha = 0.5, size = 2) +
      geom_hline(yintercept = 0, colour = 'black', alpha = 0.3, size = 1) +
      geom_vline(xintercept = 0, colour = 'black', alpha = 0.3, size = 1)
  
}




p = ggplot()

p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Other'), F, 1, 'black', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'London'), F, 2, 'red', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Core city'), T, 3, RColorBrewer::brewer.pal(3, "Set1")[2], 0.5)
p = addtogva_hours_plot(p, both %>% filter(Region_name == 'Bradford'), F, 10, 'black', 1)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'WY LA'), T, 7, RColorBrewer::brewer.pal(3, "Set1")[3], 1)

p = p + coord_fixed() +
  xlab("Average % change: hours worked per week") +
  ylab("Average % change: GVA") 


ggplotly(p, tooltip = 'Region_name')



```









# COMPANIES HOUSE DATA


```{r, chplot}

#See here for fig-responsiveness (non)issue: https://github.com/quarto-dev/quarto-cli/issues/4836

#30 seconds-ish
# x <- Sys.time()
ch = readRDS('../../companieshouseopen/local/PROCESSED_accountextracts_n_livelist_geocoded_combined_July2025.rds')
# ch = readRDS('../companieshouseopen/local/PROCESSED_accountextracts_n_livelist_geocoded_combined_July2025.rds')
# Sys.time() - x

#get short SIC names
shorts <- readRDS('../local/data/bradford_SICs_withshortnames.rds')
#shorts <- readRDS('local/data/bradford_SICs_withshortnames.rds')

ch = ch %>%
  left_join(
    shorts %>% select(SIC_5DIGIT_CODE,SIC_SECTION_NAME_SHORT:SIC_5DIGIT_NAME_SHORT),
    by = 'SIC_5DIGIT_CODE'
  )

#Create the base summary, save, reuse
#Taken from CH code here
#https://github.com/DanOlner/companieshouseopen/blob/2ddc26d4b89716b8755c02fc6626190ad624b888/testcode/initial_datadigging.R#L417
empchange.summary <- ch %>% 
  st_set_geometry(NULL) %>% 
  filter(Employees_lastyear >= 10) %>%#Need to load cos we need to filter here so we can accurately count employees in 10+ employee firms
  group_by(localauthority_name,SIC_2DIGIT_NAME_SHORT) %>%
  summarise(
    total_employmentlastyear = sum(Employees_lastyear, na.rm = T),
    total_employmentthisyear = sum(Employees_thisyear, na.rm = T),
    employment_percentchange = percent_change(total_employmentlastyear,total_employmentthisyear),
    firmcount = n()
  ) %>% 
  ungroup() %>% 
  filter(firmcount > 5)#Reduce to sectors/places with 10+ firms
# 
# saveRDS(empchange.summary,'local/data/empchangsummary.rds')
# empchange.summary <- readRDS('../local/data/empchangsummary.rds')
#empchange.summary <- readRDS('local/data/empchangsummary.rds')

# empchange10plus <- empchange.summary %>% 
#   filter(
#     firmcount > 5,
#     Employees_lastyear >= 10#Reduce to sectors/places with 10+ firms
#   )

#Order by percent change per sector in Bradford
orderby <- empchange.summary %>% 
  filter(qg('bradford',localauthority_name)) %>%
  select(SIC_2DIGIT_NAME_SHORT,employment_percentchange) %>% 
  arrange(employment_percentchange) %>% 
  pull(SIC_2DIGIT_NAME_SHORT)

#Order by those...
empchange.summary <- empchange.summary %>% 
  mutate(
    SIC_2DIGIT_NAME_SHORT = factor(SIC_2DIGIT_NAME_SHORT, levels = orderby),
    `Place: ` = paste0(localauthority_name, ', firm count: ', firmcount,', emp last/this yr: ', total_employmentlastyear,' > ',total_employmentthisyear,' = ',round(employment_percentchange,2),'%')#for labels
    # `Place: ` = paste0(SIC_2DIGIT_NAME_SHORT,',', localauthority_name, ', firm count: ', firmcount,', emp last/this yr: ', total_employmentlastyear,' > ',total_employmentthisyear)#for labels
  ) %>% 
  filter(!is.na(SIC_2DIGIT_NAME_SHORT))#Filter these here as some will be missing spefically from Bradford, will end up NA in the factor

# table(is.na(empchange.summary$SIC_2DIGIT_NAME_SHORT))

#Get values for Bradford firms - count/%/employee counts
firmcount <- ch %>% st_set_geometry(NULL) %>% filter(localauthority_name == 'Bradford', Employees_thisyear > 0)

firmcount <- firmcount %>% 
  mutate(
    sizecategory = case_when(
      Employees_thisyear == 1 ~ "1",
      between(Employees_thisyear,2,4) ~ "2-4",
      between(Employees_thisyear,5,9) ~ "5-9",
      between(Employees_thisyear,10,20) ~ "10-20",
      between(Employees_thisyear,21,50) ~ "21-50",
      between(Employees_thisyear,51,999999) ~ "51+"
    ),
    sizecategory = factor(sizecategory, levels = c('1','2-4','5-9','10-20','21-50','51+'))
    )

#table from that to plot
firmtable <- tibble(
  `Firm size` = levels(firmcount$sizecategory),
  `Count` = table(firmcount$sizecategory),
  `Percent of firms` = paste0(round(table(firmcount$sizecategory) %>% prop.table() * 100,2),"%")
)

```


```{r eval=F}
firmcounts.by.2digit <- firmcount %>% 
  group_by(sizecategory,SIC_2DIGIT_NAME_SHORT) %>% 
  summarise(
    count = n(),
    section = max(SIC_SECTION_NAME_SHORT)
    ) %>% 
  group_by(SIC_2DIGIT_NAME_SHORT) %>% 
  mutate(
    percent = (count / sum(count)) * 100
  )

ggplot(firmcounts.by.2digit, aes(y = SIC_2DIGIT_NAME_SHORT, x = percent, fill = sizecategory)) +
  geom_bar(stat = 'identity', position = position_dodge(width = 0.8)) +
  facet_wrap(~section,ncol = 2, scales = 'free')

```



The next few plots use data extracted from individual firms' accounts **filed with Companies House between June 2024 and end of June 2025**. 

The following table shows the count (and percentage) of firms in Bradford that filed with CH, broken down by employee count.


```{r bradch_summarytable}

#escape = FALSE so that markdown in the cells (the links) is rendered correctly.
knitr::kable(firmtable, escape = FALSE, caption = "Table: Bradford count/% of firms by employee band")

```



## Comparing Companies House to BRES: location quotient plots

### Production sectors

```{r CHvBRESprod}
#| fig-width: 9

#Produced here:
#https://github.com/DanOlner/RegionalEconomicTools/blob/927f1783bfbb66293dd75870e147c858f4761ef3/bits_of_code/BradfordExplore.R#L1062
plotz <- readRDS('../local/data/ch_bres_lqplotlist.rds')
#plotz <- readRDS('local/data/ch_bres_lqplotlist.rds')

plotz[[1]] + plotz[[3]]

```

### Other sectors

```{r CHvBRESother}
#| fig-width: 9

plotz[[2]] + plotz[[4]]

```




## From Companies House firm level data: in the last year (March 2024-March 2025), percent change in employment in firms, five West Yorkshire local authorities compared to 'core cities' (hover for place names)

### Firms with 10+ employees


```{r, fig-percentempfourplaces}
#| fig-height: 20
#| label: fig-percentempfourplaces
#| fig-cap: "Companies House firm level data: percent change in employment for firms with ten+ employees between most recent accounts and previous year - core cities, with West Yorkshire's five local authorities overlaid. Hover for place name."


p <- ggplot() +
  geom_point(
    position = position_nudge(y = 0.2),
    data = empchange.summary %>% filter(
      qg('Belfast|Birmingham|Bristol|Cardiff|Glasgow|Leeds|Liverpool|Manchester|Tyne|Sheffield|Nottingham', localauthority_name)#just core cities (minus four places below)
    ),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, group = `Place: `),
    alpha = 0.5)

#Add four places
p <- p + 
  geom_point(
    position = position_nudge(y = -0.2),
    data = empchange.summary %>% filter(qg('leeds|kirklees|wakefield|calderdale',localauthority_name)),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, colour = localauthority_name, group = `Place: `),
    size = 7, shape = 17) +
  scale_colour_brewer(palette = 'Paired', direction = 1, guide = guide_legend(title = NULL)) +
  # coord_cartesian(xlim = c(-25,25)) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") 

#Add Bradford over the top to make sure visible
p <- p + 
  geom_point(
    position = position_nudge(y = -0.2),
    data = empchange.summary %>% filter(qg('bradford',localauthority_name)),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, colour = localauthority_name, group = `Place: `),
    size = 8, shape = 17) +
  scale_colour_brewer(palette = 'Paired', direction = 1, guide = guide_legend(title = NULL)) +
  # coord_cartesian(xlim = c(-25,25)) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") 



#p

# ggplotly(p, width = 800, height = 900, tooltip = 'localauthority_name')
ggplotly(p, tooltip = 'Place: ')


```

###  Firms with fewer than 10 employees



```{r, fig-percentempfourplaces_lessthan10}
#| fig-height: 20
#| label: fig-percentempfourplaces_lessthan10
#| fig-cap: "From Companies House firm level data: percent change in employment for firms with fewer than 10 employees, between most recent accounts and previous year - core cities, with West Yorkshire's five local authorities overlaid. Hover for place name."

empchange.summary <- ch %>% 
  st_set_geometry(NULL) %>% 
  filter(Employees_lastyear < 10) %>%#Need to load cos we need to filter here so we can accurately count employees in 10+ employee firms
  group_by(localauthority_name,SIC_2DIGIT_NAME_SHORT) %>%
  summarise(
    total_employmentlastyear = sum(Employees_lastyear, na.rm = T),
    total_employmentthisyear = sum(Employees_thisyear, na.rm = T),
    employment_percentchange = percent_change(total_employmentlastyear,total_employmentthisyear),
    firmcount = n()
  ) %>% 
  ungroup() %>% 
  filter(firmcount > 5)#Reduce to sectors/places with 10+ firms
# 
# saveRDS(empchange.summary,'local/data/empchangsummary.rds')
# empchange.summary <- readRDS('../local/data/empchangsummary.rds')
#empchange.summary <- readRDS('local/data/empchangsummary.rds')

# empchange10plus <- empchange.summary %>% 
#   filter(
#     firmcount > 5,
#     Employees_lastyear >= 10#Reduce to sectors/places with 10+ firms
#   )

#Order by percent change per sector in Bradford
orderby <- empchange.summary %>% 
  filter(qg('bradford',localauthority_name)) %>%
  select(SIC_2DIGIT_NAME_SHORT,employment_percentchange) %>% 
  arrange(employment_percentchange) %>% 
  pull(SIC_2DIGIT_NAME_SHORT)

#Order by those...
empchange.summary <- empchange.summary %>% 
  mutate(
    SIC_2DIGIT_NAME_SHORT = factor(SIC_2DIGIT_NAME_SHORT, levels = orderby),
    `Place: ` = paste0(localauthority_name, ', firm count: ', firmcount,', emp last/this yr: ', total_employmentlastyear,' > ',total_employmentthisyear,' = ',round(employment_percentchange,2),'%')#for labels
    # `Place: ` = paste0(SIC_2DIGIT_NAME_SHORT,',', localauthority_name, ', firm count: ', firmcount,', emp last/this yr: ', total_employmentlastyear,' > ',total_employmentthisyear)#for labels
  ) %>% 
  filter(!is.na(SIC_2DIGIT_NAME_SHORT))#Filter these here as some will be missing spefically from Bradford, will end up NA in the factor

# table(is.na(empchange.summary$SIC_2DIGIT_NAME_SHORT))




p <- ggplot() +
  geom_point(
    position = position_nudge(y = 0.2),
    data = empchange.summary %>% filter(
      qg('Belfast|Birmingham|Bristol|Cardiff|Glasgow|Leeds|Liverpool|Manchester|Tyne|Sheffield|Nottingham', localauthority_name)#just core cities (minus four places below)
    ),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, group = `Place: `),
    alpha = 0.5)

#Add four places
p <- p + 
  geom_point(
    position = position_nudge(y = -0.2),
    data = empchange.summary %>% filter(qg('leeds|kirklees|wakefield|calderdale',localauthority_name)),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, colour = localauthority_name, group = `Place: `),
    size = 7, shape = 17) +
  scale_colour_brewer(palette = 'Paired', direction = 1, guide = guide_legend(title = NULL)) +
  # coord_cartesian(xlim = c(-25,25)) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") 

#Add Bradford over the top to make sure visible
p <- p + 
  geom_point(
    position = position_nudge(y = -0.2),
    data = empchange.summary %>% filter(qg('bradford',localauthority_name)),
    aes(y = SIC_2DIGIT_NAME_SHORT, x = employment_percentchange, colour = localauthority_name, group = `Place: `),
    size = 8, shape = 17) +
  scale_colour_brewer(palette = 'Paired', direction = 1, guide = guide_legend(title = NULL)) +
  # coord_cartesian(xlim = c(-25,25)) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") 

p <- p + coord_cartesian(xlim = c(-40,80))

#p

# ggplotly(p, width = 800, height = 900, tooltip = 'localauthority_name')
ggplotly(p, tooltip = 'Place: ')


```




## Just for Bradford, individual firms with 10+ employees, percent employee change from last accounts (hover for firm names)

This section is included to be able to identify individual firms (with 10+ employees) by hovering over the points. Each is plotted by what percentage their employee number grew between their latest accounts and a year prior.

```{r, fig-indivfirmsSY}
#| fig-height: 10
#| fig-width: 9

# fig-cap: "Percent change in employment between most recent accounts and previous year, ALL Bradford firms with 10+ employees, hover for firm name."

firm.change <- ch %>% 
  filter(localauthority_name == 'Bradford', Employees_thisyear >= 10 & Employees_lastyear >= 10) %>% #10+ in both time periods
  mutate(
    employee_percentchange = percent_change(Employees_lastyear,Employees_thisyear)
  )

#As well as a minimum of ten, let's break down by firm size...
#What's the spread?
# ggplot(firm.change, aes(x = Employees_thisyear)) +
#   geom_density()

firm.change <- firm.change %>% 
  mutate(
    SIC_2DIGIT_NAME_SHORT = fct_reorder(SIC_2DIGIT_NAME_SHORT, employee_percentchange),
    `Firm: ` = paste0(CompanyName,', ',Employees_lastyear,' >> ',Employees_thisyear),
    production = ifelse(SIC_2DIGIT_CODE_NUMERIC < 44, 'production','other'),
    production = factor(production,levels = c('production','other')),
    sizecategory = case_when(
      between(Employees_thisyear,10,20) ~ "10-20",
      between(Employees_thisyear,21,50) ~ "21-50",
      between(Employees_thisyear,51,999999) ~ "51+"
    )
    ) %>% 
  filter(!is.na(SIC_2DIGIT_NAME_SHORT))

# table(firm.change$sizecategory)
 # table(firm.change$production)

#FACET doesn't work well with plotly - things overlap
#Plot them separately.
p <- ggplot(firm.change %>% filter(sizecategory == "10-20"),
            aes(y = SIC_2DIGIT_NAME_SHORT, x = employee_percentchange, group = `Firm: `)) +
  geom_jitter(height = 0.1, alpha = 0.3) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") +
  facet_wrap(~production, ncol = 2, scales = 'free') +
  theme(
    # panel.grid.major = element_line(linetype = "dotted"),
    # axis.title.x = element_text( size = 10, margin=margin(80,80,80,80)),
    axis.title.y = element_text( size = 10, margin=margin(30,30,30,30))
  ) +
  ggtitle("Firms with 10-20 employees")

ggplotly(p, tooltip = 'Firm: ')

p <- ggplot(firm.change %>% filter(sizecategory == "21-50"),
            aes(y = SIC_2DIGIT_NAME_SHORT, x = employee_percentchange, group = `Firm: `)) +
  geom_jitter(height = 0.1, alpha = 0.3) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") +
  facet_wrap(~production, ncol = 2, scales = 'free') +
  theme(
    # panel.grid.major = element_line(linetype = "dotted"),
    # axis.title.x = element_text( size = 10, margin=margin(80,80,80,80)),
    axis.title.y = element_text( size = 10, margin=margin(30,30,30,30))
  ) +
  ggtitle("Firms with 21-50 employees")


ggplotly(p, tooltip = 'Firm: ')

p <- ggplot(firm.change %>% filter(sizecategory == "51+"),
            aes(y = SIC_2DIGIT_NAME_SHORT, x = employee_percentchange, group = `Firm: `)) +
  geom_jitter(height = 0.1, alpha = 0.3) +
  geom_vline(xintercept = 0, alpha = 0.5, colour = 'green') +
  ylab("") +
  facet_wrap(~production, ncol = 2, scales = 'free') +
  theme(
    # panel.grid.major = element_line(linetype = "dotted"),
    # axis.title.x = element_text( size = 10, margin=margin(80,80,80,80)),
    axis.title.y = element_text( size = 10, margin=margin(30,30,30,30))
  ) +
  ggtitle("Firms with 50+ employees")


ggplotly(p, tooltip = 'Firm: ')




```


