---
title: "Bradford: BRES location quotients for 5 digit sectors, broken down by 2 digit"
author:
  - name: "Dan Olner"
format: 
  html:
    output-file: "Bradford_BRES_5digit_LQs_July2025"
    fig-width: 8
    fig-height: 8
execute:
  echo: false
  message: false
  warning: false
fig-responsive: false
---

```{r setup}
library(tidyverse)
library(sf)#for geo stuff
library(zoo)#for making smoothed moving averages
library(patchwork)#combine ggplots easily
# library(plotly)#For interactive plots
library(RColorBrewer)

source('../functions/misc_functions.R')
source('../functions/adhoc_functions.R')
#source('functions/misc_functions.R')
#source('functions/adhoc_functions.R')

options(scipen = 999)

#Set ggplot theme
theme_set(theme_light())




```


## BRES 5 digit sector location quotients

These plots show location quotients for full time job counts in **5 digit SIC sectors in Bradford with an average of 100+ employees 2021-23.** There are 729 5 digit codes in total, though only the 100+ employee ones are shown here.

They are grouped into their parent 2 digit sectors in each plot (quite a few with only one 5 digit category in).

```{r dataprep}
#All worked out here:
#
# bres <- read_csv("../local/data/BRES/separate_SIC_types_summedfrom5digitSIC/BRES_ALLYEARSWITHDATA_NUTS3_n_ITL321_stacked_2_Fulltimeemployees_2015_2023_SIC_5DIGIT.csv")
# 
# #This comes with ALL NUTS and ITL3 data
# #Which means 2022 is doubled up
# #E.g. see here: 
# #bres %>% filter(qg('bradford', GEOGRAPHY_NAME),qg('Museum activities',SIC_5DIGIT_NAME)) %>% View
# 
# #Remove one of them...
# #Need to filter NOT year == 2022 NOT GEOGRAPHY_CODE_NUTS2018 is NA, keep the rest
# #(2023 we need to keep!)
# bres <- bres %>% 
#   filter(!(DATE == 2022 & is.na(GEOGRAPHY_CODE_NUTS2018)))
# 
# 
# #range(bres$DATE)#tick
# 
# SIClookup <- read_csv('../data/SIClookup.csv')
# 
# #Join SIC lookup on 5 digit name
# #Keep 2 digit and section codes
# #May shorten names in a mo...
# bres <- bres %>%
#   left_join(
#     SIClookup %>% select(SIC_5DIGIT_NAME,SIC_2DIGIT_NAME,SIC_SECTION_NAME),
#     by = 'SIC_5DIGIT_NAME'
#   )
# 
# #Make shorter names, use those.
# #Make lookup so can be merged in.
# names.sections = unique(bres$SIC_SECTION_NAME)
# shortnames.sections = reduceSICnames(unique(bres$SIC_SECTION_NAME),'section')
# 
# chk.sections = data.frame(names = names.sections, shortnames = shortnames.sections)
# 
# names.2digit = unique(bres$SIC_2DIGIT_NAME)
# shortnames.2digit = reduceSICnames(unique(bres$SIC_SECTION_NAME),'2 digit')
# 
# chk.2digit = data.frame(names = names.2digit, shortnames = shortnames.2digit)
# 
# 
# #And 729 categories in the 5 digit...
# names.5digit = unique(bres$SIC_5DIGIT_NAME)
# #NOTE: THE ORDER HERE IS BASED ON THE ORDER IN bres
# #We should make an actual lookup to make sure we don't lose that order
# shortnames.5digit = reduceSICnames(unique(bres$SIC_5DIGIT_NAME),'5 digit')
# 
# chk.5digit = data.frame(names = names.5digit, shortnames = shortnames.5digit)
# 
# 
# #Merge into bres
# bres <- bres %>% 
#   left_join(chk.sections %>% rename(SIC_SECTION_NAME_SHORT = shortnames), by = c('SIC_SECTION_NAME' = 'names')) %>% 
#   left_join(chk.2digit %>% rename(SIC_2DIGIT_NAME_SHORT = shortnames), by = c('SIC_2DIGIT_NAME' = 'names')) %>% 
#   left_join(chk.5digit %>% rename(SIC_5DIGIT_NAME_SHORT = shortnames), by = c('SIC_5DIGIT_NAME' = 'names')) 
# 
# 
# #Drop some sectors
# bres <- bres %>% 
#   filter(!qg('households|membership', SIC_2DIGIT_NAME))
# 
# #Check!
# # bres %>% distinct(SIC_5DIGIT_NAME_SHORT,SIC_5DIGIT_NAME) %>% View
# 
# 
# #Let's find LQ / proportions prior to smoothing those
# #Rather than smoothing job count then doing LQ
# #That way - jobs do sum in a particular year to the correct 100%
# 
# #Before doing that...
# #We need to work out which 5 digits have all-zero job counts
# #As those can't have LQs / aren't any use to us
# #Though those will vary across years
# #Might be easier to sort after calculating?
# 
# # checkonzeros <- bres %>% 
# #   group_by(DATE,GEOGRAPHY_NAME,SIC_5DIGIT_NAME_SHORT) %>% 
# #   summarise(
# #     sectorpercent_zeroes = mean(JOBCOUNT == 0) * 100
# #   )
# # 
# # unique(checkonzeros$sectorpercent_zeroes)
# # 
# # #OK, let's check if the zeroes persist over all the years in the data
# # checkonzeros.allyears <- bres %>% 
# #   group_by(GEOGRAPHY_NAME,SIC_5DIGIT_NAME_SHORT) %>% 
# #   summarise(
# #     sectorpercent_zeroes = mean(JOBCOUNT == 0) * 100
# #   )
# 
# #And across both all years and all places?
# checkonzeros.allyearsandplaces <- bres %>% 
#   group_by(SIC_5DIGIT_NAME_SHORT) %>% 
#   summarise(
#     sectorpercent_zeroes = mean(JOBCOUNT == 0) * 100
#   )
# 
# #Final - how does that vary per year?
# # checkonzeros.peryear <- bres %>% 
# #   group_by(DATE,SIC_5DIGIT_NAME_SHORT) %>% 
# #   summarise(
# #     sectorpercent_zeroes = mean(JOBCOUNT == 0) * 100
# #   )
# 
# 
# #OK - going to remove any sectors with 100% zeroes for all years
# #Which are all agri
# #And the one NA to (also agri!)
# sectorstodrop <- checkonzeros.allyearsandplaces %>% 
#   filter(sectorpercent_zeroes == 100 | is.na(sectorpercent_zeroes)) %>% 
#   select(SIC_5DIGIT_NAME_SHORT) %>% 
#   pull
# 
# bres <- bres %>% 
#   filter(
#     !SIC_5DIGIT_NAME_SHORT %in% sectorstodrop
#   )
# 
# 
# 
# #NOW find props (will still have some issues but can deal with below)
# bres <- bres %>% 
#   arrange(DATE) %>% 
#   group_split(DATE) %>%
#   map(add_location_quotient_and_proportions,
#       regionvar = GEOGRAPHY_NAME,
#       lq_var = SIC_5DIGIT_NAME_SHORT,
#       valuevar = JOBCOUNT) %>% 
#   bind_rows()
# 
# 
# #Then smooth the props and LQs...
# bres.smooth <- bres %>% 
#   arrange(DATE) %>% 
#   group_by(GEOGRAPHY_NAME,SIC_5DIGIT_NAME_SHORT) %>%
#   mutate(
#     jobcount_movingav = rollapply(JOBCOUNT,3,mean,align='center',fill=NA),
#     sector_regional_proportion_movingav = rollapply(sector_regional_proportion,3,mean,align='center',fill=NA),
#     LQ_movingav = rollapply(LQ,3,mean,align='center',fill=NA)
#   ) %>% 
#   ungroup()
# 
# 
# #Looking at one place / sector to check it looks sane
# # bres.smooth %>% filter(qg('bradford', GEOGRAPHY_NAME),qg('prepared meals',SIC_5DIGIT_NAME_SHORT)) %>% View
# # bres.smooth %>% 
# #   # filter(qg('bradford', GEOGRAPHY_NAME)) %>% 
# #   filter(qg('Liverpool', GEOGRAPHY_NAME)) %>% 
# #   select(DATE,JOBCOUNT,SIC_5DIGIT_NAME_SHORT,sector_regional_proportion,LQ,jobcount_movingav,LQ_movingav) %>% 
# #   arrange(SIC_5DIGIT_NAME_SHORT,DATE) %>% 
# #   filter(!is.na(LQ_movingav)) %>% 
# #   View
# 
# 
# 
# 
# #Keep only smoothed years with data (the centre point)
# #And re-find the log value after smoothing...
# # x <- bres.smooth %>% 
# bres.smooth <- bres.smooth %>%
#   filter(!is.na(LQ_movingav)) %>% 
#   mutate(LQ_movingav_log = log(LQ_movingav))
# 
# #RIGHT, NOW IT'S IN A FORM TO STICK IN THE LQ PLOTS
# #Which I want to break down by section to begin with, though may make sector groupings bespoke
# 
# 
# #So - run for each group of 5 digits
# #But I want to put them in some LQ order for a place - 
# #Average order for those sectors in each grouping would make sense
# #Though could also do by section
# 
# #Get those values then we can use to order plot creation
# #Average for the most recent (smoothed) year in our chosen place
# 
place = 'Bradford'
# 
# #sections too many! Try smaller
# avLQvalues_in_groupings <- bres.smooth %>%
#   filter(DATE == max(DATE), GEOGRAPHY_NAME == place) %>% 
#   group_by(SIC_2DIGIT_NAME_SHORT) %>% 
#   summarise(mean_LQ = mean(LQ, na.rm = T)) %>% 
#   arrange(desc(mean_LQ))
# 
# 
# #Can use each of those to feed subsets into an LQ plot
# #In this case, can calculate everything beforehand then feed in
# #Order should still hold in subplots filtered by section
# LQ_slopes <- compute_slope_or_zero(
#   data = bres.smooth, 
#   GEOGRAPHY_NAME, SIC_5DIGIT_NAME_SHORT,#slopes will be found within whatever grouping vars are added here
#   y = LQ_movingav_log, x = DATE)
# 
# #Filter down to a single year... we may want to smooth years, let's see
# yeartoplot <- bres.smooth %>% filter(DATE == max(DATE))#use latest year
# 
# #Add slopes into data to get LQ plots
# yeartoplot <- yeartoplot %>% 
#   left_join(
#     LQ_slopes,
#     by = c('GEOGRAPHY_NAME', 'SIC_5DIGIT_NAME_SHORT')
#   )
# 
# #Get min/max values for LQ over time as well, for each sector and place, to add as bars so range of sector is easy to see
# minmaxes <- bres.smooth %>% 
#   group_by(GEOGRAPHY_NAME, SIC_5DIGIT_NAME_SHORT) %>% 
#   summarise(
#     min_LQ_all_time = min(LQ_movingav),
#     max_LQ_all_time = max(LQ_movingav)
#   )
# 
# #Join min and max
# yeartoplot <- yeartoplot %>% 
#   left_join(
#     minmaxes,
#     by = c('GEOGRAPHY_NAME', 'SIC_5DIGIT_NAME_SHORT')
#   )
# 
# 
# 
# 
# sectorLQorder <- yeartoplot %>% filter(
#   GEOGRAPHY_NAME == place,
#   DATE == max(DATE)#use latest data
# ) %>% 
#   arrange(-LQ) %>% 
#   select(SIC_5DIGIT_NAME_SHORT) %>% 
#   pull()
# 
# 
# #Turn the sector column into a factor and order by LCR's LQs
# yeartoplot$SIC_5DIGIT_NAME_SHORT <- factor(yeartoplot$SIC_5DIGIT_NAME_SHORT, levels = sectorLQorder, ordered = T)
# 
# #Examine to check what to filter out
# #yeartoplot %>% filter(GEOGRAPHY_NAME == 'Bradford') %>% View
# 
# 
# 
# 
# #Test some alterations
# #Use function in adhoc_functions
# #And send in the actual df already filtered to what we want to plot
# #Don't plot any 2 digit SICs with no values we want to show
# #Shorten here to subset of sectors
# 
# #Shortern the list of 2 digit SICs to ones that actually have values we want to look at
# #Keep only these 5 digit sectors for Bradford
# keep.these.5digits <- yeartoplot %>% 
#   filter(GEOGRAPHY_NAME == 'Bradford') %>% 
#   filter(LQ_movingav > 0.001, jobcount_movingav >= 100) %>% 
#   select(SIC_5DIGIT_NAME_SHORT) %>%
#   mutate(SIC_5DIGIT_NAME_SHORT = as.character(SIC_5DIGIT_NAME_SHORT)) %>% 
#   pull
# 
# #filter the yeartoplot data (which includes all other places)
# #Based on that, to get a list of what 2 digits are left
# yeartoplot.sub <- yeartoplot %>%
#   filter(
#     SIC_5DIGIT_NAME_SHORT %in% keep.these.5digits
#   ) %>%
#   mutate(jobcount_movingav = round(jobcount_movingav,0))

# length(unique(yeartoplot$SIC_2DIGIT_NAME_SHORT))
# length(unique(yeartoplot.sub$SIC_2DIGIT_NAME_SHORT))

#Plot each SIC 2 digit for the 5 digits in them

#Save and reuse
# saveRDS(yeartoplot.sub,'local/data/2023BRES_5digit_LQs.rds')
yeartoplot.sub <- readRDS('../local/data/2023BRES_5digit_LQs.rds')
#yeartoplot.sub <- readRDS('local/data/2023BRES_5digit_LQs.rds')

#Quick look at orig SIC 2 digits and shortened versions
yeartoplot.sub %>% select(SIC_2DIGIT_NAME,SIC_2DIGIT_NAME_SHORT) %>% distinct() %>% View

plotz <- map(yeartoplot.sub %>% group_split(SIC_2DIGIT_NAME_SHORT),LQplot_BRES_groupsof5digit)

#Find which ones are in which sizes and output accordingly
plotz.sectorcounts <- yeartoplot.sub %>% filter(GEOGRAPHY_NAME == 'Bradford') %>% group_by(SIC_2DIGIT_NAME_SHORT) %>% 
  summarise(sectorcount = n()) %>% select(sectorcount) %>% pull



```


```{r larger}
#| fig-height: 12

#Output larger first
subz <- plotz[which(plotz.sectorcounts >= 9)]

#Avoid list numbers
for(item in subz) print(item)

```


```{r medium}
#| fig-height: 4

#Output larger first
subz <- plotz[which(between(plotz.sectorcounts,2,8))]

#Avoid list numbers
for(item in subz) print(item)


```

```{r small}
#| fig-height: 2

#Output larger first
subz <- plotz[which(plotz.sectorcounts == 1)]

#Avoid list numbers
for(item in subz) print(item)


```







```{r}

# indices <- seq(1, 6, by = 2)
# indices <- seq(59, 62, by = 2)
# indices <- seq(1, length(plotz), by = 2)

#Avoid list numbers
# for(item in plotz) print(item)
# map(
#   indices,
#   ~ {
#     if(.x+1 <= length(plotz)){
#       patchwork::wrap_plots(plotz[.x:(.x+1)], ncol = 2)
#     } else{
#       patchwork::wrap_plots(plotz[[.x]], ncol = 2)
#     }
#   }
#   )

#x=1
# patchwork::wrap_plots(plotz[x:(x+1)], ncol = 2)

# Dynamically print each plot with custom height
# for (item in plotz) {
#   
#   n_sectors <- length(unique(ggplot_build(item)$data[[1]]$y))
#   height_in <- 1 + 0.3 * n_sectors  # adjust scale as needed
#   
#   cat(knitr::knit_print(
#     knitr::asis_output(
#       paste0(
#         "```{r, fig.height=", height_in, ", fig.width=6, echo=FALSE}\n",
#         "print(plot_list[[", i, "]])\n",
#         "```\n"
#       )
#     )
#   ))
# }

# for (i in seq_along(plotz[1:8])) {
#   
#   p <- plotz[[i]]
#   
#   n_sectors <- length(unique(ggplot_build(p)$data[[1]]$y))
#   
#   height_in <- 1 + 0.3 * n_sectors
  
  # cat(knitr::knit_print(
  #   knitr::asis_output(
  #     paste0(
  #       "```{r, fig.height=", height_in, ", fig.width=6, echo=FALSE}\n",
  #       "print(plot_list[[", i, "]])\n",
  #       "```\n"
  #     )
  #   )
  # ))
  
  # knitr::knit_child(text = 
  #                     paste0(
  #       "```{r, fig.height=", height_in, ", fig.width=6, echo=FALSE}\n",
  #       "print(plotz[[", i, "]])\n",
  #       "```\n"
  #     )
  # )
  # 
  
  # Dynamically evaluate code chunks with plot + custom height
  # for (i in seq_along(plotz)) {
  #   
  #   p <- plotz[[i]]
  #   
  #   n_sectors <- length(unique(ggplot_build(p)$data[[1]]$y))
  #   
  #   height_in <- 1 + 0.3 * n_sectors  # adjust as needed
  #   
  #   code <- paste0(
  #     "```{r, fig.height=", height_in, ", fig.width=6, echo=FALSE}\n",
  #     "print(plotz[[", i, "]])\n",
  #     "```")
  # 
  #   evaluate::evaluate(code, envir = environment())  # evaluates and renders the chunk
  # 
  # 
  # }



```




