---
title: "Greater Manchester: firm counts and percents broken down by borough, size band and broad sector"
author:
  - name: "Dan Olner"
format: 
  html:
    output-file: "GM_firmcounts_bysize_sector"
    fig-width: 9
    fig-height: 12
execute:
  echo: false
  message: false
  warning: false
fig-responsive: false
collapse: false
---

```{r setup}
library(tidyverse)
library(plotly)#For interactive plots
library(RColorBrewer)
source('../functions/misc_functions.R')

options(scipen = 999)

#orig code in businesscounts_explore.R

```


## What's this?

* Business count data (via [NOMIS](https://www.nomisweb.co.uk/sources/ukbc)) for Greater Manchester boroughs, broken down into four firm size bands and broad sectors (SIC sections). This latest data is for 2024.
* Two separate plots for local units vs enterprises. See below for more on that.
* Every data point is "percent of a borough's **total unit count**", to make them comparable across places. **Hover for exact percent as well as firm count**.
* England values included for comparison.

Prep code [here](https://github.com/DanOlner/RegionalEconomicTools/blob/gh-pages/bits_of_code/businesscounts_explore.R), code for this quarto doc [here](https://github.com/DanOlner/RegionalEconomicTools/blob/gh-pages/quarto_docs/GM_boroughs_firmcount_by_sizeband_section_2024.qmd).


## Enterprises vs local units

[This doc](https://www.ons.gov.uk/file?uri=%2Fbusinessindustryandtrade%2Fbusiness%2Factivitysizeandlocation%2Fmethodologies%2Fbusinesspopulation%2Ffurtherinformationaboutidbrsourcestcm77325481.pdf) has as good outline that includes the different data sources used for each. Both will include public and private organisations. Quick summary:

* Local units are **all business units** including head offices, branches, factories etc. These can be located in different places (e.g. supermarket branches). Source is BRES survey data.
* Enterprises are where HQs are located. From NOMIS: "An Enterprise is the smallest combination of legal units which has a certain degree of autonomy within an Enterprise Group."


# Plots

## Local units


```{r, fig1}

gm <- readRDS('../local/data/greatermanchester_businesscounts_localunits_byfirmsizeandsector2024.rds')

# saveRDS(eng,'../local/data/england_businesscounts_localunits_byfirmsizeandsector2024.rds')
eng <- readRDS('../local/data/england_businesscounts_localunits_byfirmsizeandsector2024.rds')

#Keep just two digit
gm <- gm %>% filter(INDUSTRY_TYPE == 'SIC 2007 division (2 digit)') 

#Merge in sections
SIClookup <- read_csv('../data/SIClookup.csv')

#Check match... tick
# table(unique(SIClookup$SIC_2DIGIT_CODE) %in% gm$INDUSTRY_CODE)

#Merge in sections...
gm <- gm %>% 
  rename(SIC_2DIGIT_CODE = INDUSTRY_CODE) %>% 
  left_join(
    SIClookup %>% select(SIC_2DIGIT_CODE,SIC_SECTION_CODE,SIC_SECTION_NAME) %>% distinct(),
    by = 'SIC_2DIGIT_CODE'
  )

#Shortemed!
gm <- gm %>% mutate(SIC_SECTION_NAME_SHORTERNED = reduceSICnames(SIC_SECTION_NAME,'section'))

# unique(gm$SIC_SECTION_NAME_SHORTERNED)


#Remove a couple we don't want
gm <- gm %>% filter(!qg('extrat|households',SIC_SECTION_NAME_SHORTERNED))


#So plot wise...
#Prob want 1 section, 1 firm size, all 10 boroughs on one plot? Test...

#We want "count of firms of different sizes as a percent of total firm count" in each borough
#Note also at the moment, haven't summed counts per section

#Sums per section! Let's do that first
#We don't want every single employment size band (some overlap e.g. there's 0-4 and micro 0-9) 
#Drop the ones we don't want
gm.sectionsums <- gm %>% 
  filter(
    EMPLOYMENT_SIZEBAND_NAME %in% c('Micro (0 to 9)','Small (10 to 49)','Medium-sized (50 to 249)','Large (250+)')
    # EMPLOYMENT_SIZEBAND_NAME %in% c('0 to 4','10 to 19','20 to 49','')
  ) %>% 
  group_by(EMPLOYMENT_SIZEBAND_NAME,SIC_SECTION_NAME_SHORTERNED,GEOGRAPHY_NAME) %>% 
  summarise(firmcount = sum(OBS_VALUE)) %>% 
  ungroup()
  

#Add in total firm count per borough, so can then find percent of firm count for each sizeband/section combo
#(And mull if that's the best denominator... e.g. could examine a particular sector like ICT)
gm.sectionsums <- gm.sectionsums %>% 
  group_by(GEOGRAPHY_NAME) %>% 
  mutate(
    totalfirmcount_per_LA = sum(firmcount),
    percent_firmcount = (firmcount/totalfirmcount_per_LA) * 100,
    EMPLOYMENT_SIZEBAND_NAME = factor(EMPLOYMENT_SIZEBAND_NAME, levels = c('Micro (0 to 9)','Small (10 to 49)','Medium-sized (50 to 249)','Large (250+)'))) %>% 
  ungroup() 



#REPEAT FOR ENGLAND
eng <- eng %>% filter(INDUSTRY_TYPE == 'SIC 2007 division (2 digit)') 

#Merge in sections
#Check match... tick
# table(unique(SIClookup$SIC_2DIGIT_CODE) %in% eng$INDUSTRY_CODE)

#Merge in sections...
eng <- eng %>% 
  rename(SIC_2DIGIT_CODE = INDUSTRY_CODE) %>% 
  left_join(
    SIClookup %>% select(SIC_2DIGIT_CODE,SIC_SECTION_CODE,SIC_SECTION_NAME) %>% distinct(),
    by = 'SIC_2DIGIT_CODE'
  )

#Shortemed!
eng <- eng %>% mutate(SIC_SECTION_NAME_SHORTERNED = reduceSICnames(SIC_SECTION_NAME,'section'))

# unique(eng$SIC_SECTION_NAME_SHORTERNED)


#Remove a couple we don't want
eng <- eng %>% filter(!qg('extrat|households',SIC_SECTION_NAME_SHORTERNED))

#Sums per section! Let's do that first
#We don't want every single employment size band (some overlap e.g. there's 0-4 and micro 0-9) 
#Drop the ones we don't want
eng.sectionsums <- eng %>% 
  filter(
    EMPLOYMENT_SIZEBAND_NAME %in% c('Micro (0 to 9)','Small (10 to 49)','Medium-sized (50 to 249)','Large (250+)')
  ) %>% 
  group_by(EMPLOYMENT_SIZEBAND_NAME,SIC_SECTION_NAME_SHORTERNED) %>% 
  summarise(firmcount = sum(OBS_VALUE)) %>% 
  ungroup()


#Add in total firm count per borough, so can then find percent of firm count for each sizeband/section combo
#(And mull if that's the best denominator... e.g. could examine a particular sector like ICT)
eng.sectionsums <- eng.sectionsums %>% 
  mutate(
    GEOGRAPHY_NAME = "England",#add this back in to combine below
    totalfirmcount_per_LA = sum(firmcount),
    percent_firmcount = (firmcount/totalfirmcount_per_LA) * 100,
    EMPLOYMENT_SIZEBAND_NAME = factor(EMPLOYMENT_SIZEBAND_NAME, levels = c('Micro (0 to 9)','Small (10 to 49)','Medium-sized (50 to 249)','Large (250+)'))) %>% 
  ungroup() 

#EVERYTHING PLUS ENGLAND
both <- bind_rows(gm.sectionsums,eng.sectionsums) %>% 
  mutate(
    GEOGRAPHY_NAME = factor(GEOGRAPHY_NAME, levels = c(
      "Bolton", "Bury", "Manchester", "Oldham", "Rochdale", "Salford", "Stockport", "Tameside", "Trafford", "Wigan","England"
    )),
    `%:` = paste0(round(percent_firmcount,1), ', count: ',totalfirmcount_per_LA)
  ) %>% rename(
    Borough = GEOGRAPHY_NAME
  )

p <- ggplot(
  both, 
  # gm.sectionsums, 
  aes(x = SIC_SECTION_NAME_SHORTERNED, y = percent_firmcount, fill = Borough, label = `%:`)
) +
  geom_bar(stat='identity', position = 'dodge') +
  xlab("SIC section") +
  ylab("% of borough's total local unit count (with England av)") +
  scale_fill_brewer(palette = "Paired", direction = 1) +
  theme(legend.title=element_blank()) +
  facet_wrap(~EMPLOYMENT_SIZEBAND_NAME, nrow = 1, scales = 'free_x') +
  coord_flip() +
  xlab("")
  # theme(legend.title=element_blank())#doesn't work in ggplotly
  # scale_color_discrete(name="")#Neither does this!

ggplotly(p, tooltip = '%:')



```



## Enterprises


```{r, fig2}

gm <- readRDS('../local/data/greatermanchester_businesscounts_enterprises_byfirmsizeandsector2024.rds')


#Also get for England, too add average... 2092957699
# d <- nomis_get_metadata(id = "NM_142_1", concept = "geography", type = "TYPE499")

# eng <- nomis_get_data(id = "NM_142_1", time = "latest", MEASURES = 20100, LEGAL_STATUS = 0, geography = "2092957699")

# saveRDS(eng,'../local/data/england_businesscounts_enterprises_byfirmsizeandsector2024.rds')
eng <- readRDS('../local/data/england_businesscounts_enterprises_byfirmsizeandsector2024.rds')

#Keep just two digit
gm <- gm %>% filter(INDUSTRY_TYPE == 'SIC 2007 division (2 digit)') 

#Merge in sections
#Check match... tick
# table(unique(SIClookup$SIC_2DIGIT_CODE) %in% gm$INDUSTRY_CODE)

#Merge in sections...
gm <- gm %>% 
  rename(SIC_2DIGIT_CODE = INDUSTRY_CODE) %>% 
  left_join(
    SIClookup %>% select(SIC_2DIGIT_CODE,SIC_SECTION_CODE,SIC_SECTION_NAME) %>% distinct(),
    by = 'SIC_2DIGIT_CODE'
  )

#Shortemed!
gm <- gm %>% mutate(SIC_SECTION_NAME_SHORTERNED = reduceSICnames(SIC_SECTION_NAME,'section'))

# unique(gm$SIC_SECTION_NAME_SHORTERNED)


#Remove a couple we don't want
gm <- gm %>% filter(!qg('extrat|households',SIC_SECTION_NAME_SHORTERNED))


#So plot wise...
#Prob want 1 section, 1 firm size, all 10 boroughs on one plot? Test...

#We want "count of firms of different sizes as a percent of total firm count" in each borough
#Note also at the moment, haven't summed counts per section

#Sums per section! Let's do that first
#We don't want every single employment size band (some overlap e.g. there's 0-4 and micro 0-9) 
#Drop the ones we don't want
gm.sectionsums <- gm %>% 
  filter(
    EMPLOYMENT_SIZEBAND_NAME %in% c('Micro (0 to 9)','Small (10 to 49)','Medium-sized (50 to 249)','Large (250+)')
    # EMPLOYMENT_SIZEBAND_NAME %in% c('0 to 4','10 to 19','20 to 49','')
  ) %>% 
  group_by(EMPLOYMENT_SIZEBAND_NAME,SIC_SECTION_NAME_SHORTERNED,GEOGRAPHY_NAME) %>% 
  summarise(firmcount = sum(OBS_VALUE)) %>% 
  ungroup()


#Add in total firm count per borough, so can then find percent of firm count for each sizeband/section combo
#(And mull if that's the best denominator... e.g. could examine a particular sector like ICT)
gm.sectionsums <- gm.sectionsums %>% 
  group_by(GEOGRAPHY_NAME) %>% 
  mutate(
    totalfirmcount_per_LA = sum(firmcount),
    percent_firmcount = (firmcount/totalfirmcount_per_LA) * 100,
    EMPLOYMENT_SIZEBAND_NAME = factor(EMPLOYMENT_SIZEBAND_NAME, levels = c('Micro (0 to 9)','Small (10 to 49)','Medium-sized (50 to 249)','Large (250+)'))) %>% 
  ungroup() 



#REPEAT FOR ENGLAND
eng <- eng %>% filter(INDUSTRY_TYPE == 'SIC 2007 division (2 digit)') 

#Merge in sections
#Check match... tick
# table(unique(SIClookup$SIC_2DIGIT_CODE) %in% eng$INDUSTRY_CODE)

#Merge in sections...
eng <- eng %>% 
  rename(SIC_2DIGIT_CODE = INDUSTRY_CODE) %>% 
  left_join(
    SIClookup %>% select(SIC_2DIGIT_CODE,SIC_SECTION_CODE,SIC_SECTION_NAME) %>% distinct(),
    by = 'SIC_2DIGIT_CODE'
  )

#Shortemed!
eng <- eng %>% mutate(SIC_SECTION_NAME_SHORTERNED = reduceSICnames(SIC_SECTION_NAME,'section'))

# unique(eng$SIC_SECTION_NAME_SHORTERNED)


#Remove a couple we don't want
eng <- eng %>% filter(!qg('extrat|households',SIC_SECTION_NAME_SHORTERNED))

#Sums per section! Let's do that first
#We don't want every single employment size band (some overlap e.g. there's 0-4 and micro 0-9) 
#Drop the ones we don't want
eng.sectionsums <- eng %>% 
  filter(
    EMPLOYMENT_SIZEBAND_NAME %in% c('Micro (0 to 9)','Small (10 to 49)','Medium-sized (50 to 249)','Large (250+)')
  ) %>% 
  group_by(EMPLOYMENT_SIZEBAND_NAME,SIC_SECTION_NAME_SHORTERNED) %>% 
  summarise(firmcount = sum(OBS_VALUE)) %>% 
  ungroup()


#Add in total firm count per borough, so can then find percent of firm count for each sizeband/section combo
#(And mull if that's the best denominator... e.g. could examine a particular sector like ICT)
eng.sectionsums <- eng.sectionsums %>% 
  mutate(
    GEOGRAPHY_NAME = "England",#add this back in to combine below
    totalfirmcount_per_LA = sum(firmcount),
    percent_firmcount = (firmcount/totalfirmcount_per_LA) * 100,
    EMPLOYMENT_SIZEBAND_NAME = factor(EMPLOYMENT_SIZEBAND_NAME, levels = c('Micro (0 to 9)','Small (10 to 49)','Medium-sized (50 to 249)','Large (250+)'))) %>% 
  ungroup() 


#EVERYTHING
#EVERYTHING PLUS ENGLAND
both <- bind_rows(gm.sectionsums,eng.sectionsums) %>% 
  mutate(
    GEOGRAPHY_NAME = factor(GEOGRAPHY_NAME, levels = c(
      "Bolton", "Bury", "Manchester", "Oldham", "Rochdale", "Salford", "Stockport", "Tameside", "Trafford", "Wigan","England"
    )),
    `%:` = paste0(round(percent_firmcount,1), ', count: ',totalfirmcount_per_LA)
  ) %>% rename(
    Borough = GEOGRAPHY_NAME
  )

p <- ggplot(
  both, 
  # gm.sectionsums, 
  aes(x = SIC_SECTION_NAME_SHORTERNED, y = percent_firmcount, fill = Borough, label = `%:`)
) +
  geom_bar(stat='identity', position = 'dodge') +
  xlab("SIC section") +
  ylab("% of borough's total enterprise count (with England av)") +
  scale_fill_brewer(palette = "Paired", direction = 1) +
  theme(legend.title=element_blank()) +
  facet_wrap(~EMPLOYMENT_SIZEBAND_NAME, nrow = 1, scales = 'free_x') +
  coord_flip() +
  xlab("")
# theme(legend.title=element_blank())#doesn't work in ggplotly
# scale_color_discrete(name="")#Neither does this!

ggplotly(p, tooltip = '%:')




```
