---
title: "productivity bits June 2025"
sidebar: false
author:
  - name: "Dan Olner"
format: 
  html:
    output-file: "productivity_bitsandbobs_June2025"
    fig-width: 8
    fig-height: 8
execute:
  echo: false
  message: false
  warning: false
fig-responsive: false
---

```{r setup}
library(tidyverse)
library(spdep)
library(sf)#for geo stuff
library(tmap)#for mapping
library(zoo)#for making smoothed moving averages
library(patchwork)#combine ggplots easily
library(plotly)#For interactive plots
# library(ggdist)
# library(tidyr)
# library(distributional)
library(modelsummary)
library(RColorBrewer)

source('../functions/misc_functions.R')
#source('functions/misc_functions.R')
options(scipen = 999)

#Set ggplot theme
# theme_set(theme_light())


```

```{r dataprep}
# ITL3 Hours % change v GVA % change on average----

#So need the number of hours and chained volume GVA
#Hours count is in the per hour productivity sheet
#GVA in the region by industry sheet

#hours worked first
url1 <- 'https://www.ons.gov.uk/file?uri=/employmentandlabourmarket/peopleinwork/labourproductivity/datasets/subregionalproductivitylabourproductivitygvaperhourworkedandgvaperfilledjobindicesbyuknuts2andnuts3subregions/current/labourproductivityitls1.xlsx'
p1f <- tempfile(fileext=".xlsx")
download.file(url1, p1f, mode="wb")

#"Productivity Hours Worked per Week; ITL2 and ITL3 subregions (constrained to ITL1), 2004 - 2023"
hoursworked <- readxl::read_excel(path = p1f,range = "Productivity Hours!A5:W247")

names(hoursworked) <- gsub(x = names(hoursworked), pattern = ' ', replacement = '_')

hoursworked = hoursworked %>%
  filter(ITL_level == 'ITL3') %>% 
  pivot_longer(Hours_2004:Hours_2023, names_to = 'year', values_to = 'hoursworkedperweek') %>% #get most recent year
  mutate(
    year = substr(year,7,10),
    year = as.numeric(year)
  ) %>% 
  select(-ITL_level)



#Then ITL3 CV for all industries (so has to include imputed rent as we can't sum CV)
url2 <- 'https://www.ons.gov.uk/file?uri=/economy/grossvalueaddedgva/datasets/nominalandrealregionalgrossvalueaddedbalancedbyindustry/current/regionalgrossvalueaddedbalancedbyindustryandallinternationalterritoriallevelsitlregions.xlsx'
p2f <- tempfile(fileext=".xlsx")
download.file(url2, p2f, mode="wb") 

#"Table 3b: ITL3, chained volume measures in 2019 money value, pounds million [note 2]"
itl3.all <- readxl::read_excel(path = p2f, range = "Table 3b!A2:AD11468")

names(itl3.all) <- gsub(x = names(itl3.all), pattern = ' ', replacement = '_')

itl3.all = itl3.all %>% 
  filter(SIC07_description == 'All industries') %>% 
  pivot_longer(`1998`:names(itl3.all)[length(names(itl3.all))], names_to = 'year', values_to = 'value') %>% #get most recent year
  mutate(year = as.numeric(year)) %>% 
  rename(gva_cv = value)



#Check link between two is valid... tick
# table(unique(hoursworked$ITL_code) %in% itl3.all$ITL_code)

#2004:2023
# unique(hoursworked$year)[unique(hoursworked$year) %in% unique(itl3.all$year)]

#Get list of corecities
corecities <- getdistinct('sheffield|Belfast|Birmingham|Bristol|Cardiff|Glasgow|Leeds|Liverpool|Manchester|Tyne|Nottingham', itl3.all$Region_name)

#Yes, all in there, need to remove a few...
corecities <- corecities[!grepl(x = corecities, pattern = 'Greater|shire', ignore.case = T)]
corecities <- corecities[order(corecities)]

#Check match also works in hours worked... tick
# table(corecities %in% hoursworked$Region_name)



#Get London ITL3s
itl_lookup = read_csv("../data/LAD_(December_2024)_to_LAU1_to_ITL3_to_ITL2_to_ITL1_(January_2025)_Lookup_in_the_UK.csv")

#Get all London ITL2 zones then use to get unique ITL3 zones
itl3.lon = itl_lookup %>%
  filter(qg('london',ITL225NM)) %>% 
  select(ITL325CD,ITL325NM) %>% 
  distinct()

#Check itl3 code match... tick
# table(itl3.lon$ITL325CD %in% hoursworked$ITL_code)
# table(itl3.lon$ITL325CD %in% itl3.all$ITL_code)
#tick
# table(itl3.lon$ITL325NM %in% hoursworked$Region_name)
# table(itl3.lon$ITL325NM %in% itl3.all$Region_name)


```







## AV. PERCENT CHANGE OVER TIME FOR GVA VS HOURS WORKED, ITL3 ZONES

**Hover for place names**.

* South Yorkshire ITL3s in green (95% CIs)
* Core cities in blue  (95% CIs)
* London ITL3s in red/pink

And:

* Anything above/left of diagonal line is a **productivity increase**. So e.g. in the plot below for 2014-2023, Rotherham's productivity increased with a **GVA increase** and a **number of hours worked *decrease***.

### 2004 to 2013 av % change

```{r plot1}
# startyear = 2014
# endyear = 2023

startyear = 2004
endyear = 2013

#Hours per week first
slopes.hours = get_slope_and_se_safely(
  hoursworked %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(hoursworkedperweek), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.hours = slopes.hours %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )


#Mark 4 SY LAs and plot
slopes.hours = slopes.hours %>% 
  ungroup() %>% 
  # mutate(
  #   SY_LAs = ifelse(qg('rotherh|doncaste|sheffie|barnsley', Region_name),'SY LA','Other'),
  #   corecities = ifelse(Region_name %in% corecities,'core city','Other'),
  #   london = ifelse(qg('london',Region_name), 'London','Other')
  # )
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'SY LA',#will overwrite sheffield as core city
      qg('rotherh|doncaste|sheffie|barnsley', Region_name) ~ 'SY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
      ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','SY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.hours = slopes.hours %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )




#Then GVA CV
slopes.gva = get_slope_and_se_safely(
  itl3.all %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(gva_cv), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.gva = slopes.gva %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )



#Mark 4 SY LAs and plot
slopes.gva = slopes.gva %>% 
  ungroup() %>% 
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'SY LA',#will overwrite sheffield as core city
      qg('rotherh|doncaste|sheffie|barnsley', Region_name) ~ 'SY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
    ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','SY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.gva = slopes.gva %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )


# table(slopes.hours$itl3label)
# table(slopes.gva$itl3label)


#Keep the bits we want and combine into single df
both <- slopes.gva %>% 
  select(Region_name,itl3label,
         slope_percentperyear_GVA = slope_percentperyear,
         ci95min_percentperyear_GVA = ci95min_percentperyear,
         ci95max_percentperyear_GVA = ci95max_percentperyear) %>% 
  inner_join(
    slopes.hours %>% 
      select(Region_name,
             slope_percentperyear_HOURS = slope_percentperyear,
             ci95min_percentperyear_HOURS = ci95min_percentperyear,
             ci95max_percentperyear_HOURS = ci95max_percentperyear),
    by = 'Region_name'
  )


#Poor control over visibility. Layer instead. Base plot then function.
addtogva_hours_plot = function(p,plotdata, showerrors = T, sizeval = 1, colourval = 'black', alphaval = 1){
  
  p = p + geom_point(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, group = Region_name), colour = colourval, size = sizeval, alpha = alphaval) 
  
  if(showerrors){
    
    p = p + geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, xmin = ci95min_percentperyear_HOURS, xmax = ci95max_percentperyear_HOURS),width = 0.1, size = 1, colour = colourval, alpha = alphaval) +
    geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, ymin = ci95min_percentperyear_GVA, ymax = ci95max_percentperyear_GVA),width = 0.1, size = 1, colour = colourval, alpha = alphaval) 
    
  }
  
    p  + geom_abline(slope = 1, intercept = 0, colour = 'black', alpha = 0.5, size = 2) +
      geom_hline(yintercept = 0, colour = 'black', alpha = 0.3, size = 1) +
      geom_vline(xintercept = 0, colour = 'black', alpha = 0.3, size = 1)
  
}




p = ggplot()

p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Other'), F, 1, 'black', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'London'), F, 2, 'red', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Core city'), T, 3, RColorBrewer::brewer.pal(3, "Set1")[2], 0.5)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'SY LA'), T, 7, RColorBrewer::brewer.pal(3, "Set1")[3], 1)

p = p + 
  coord_fixed() +
  xlab("Average % change: hours worked per week") +
  ylab("Average % change: GVA") 

ggplotly(p, tooltip = 'Region_name')



```

### 2014 to 2023 av % change

```{r plot2}
startyear = 2014
endyear = 2023

# startyear = 2004
# endyear = 2013

#Hours per week first
slopes.hours = get_slope_and_se_safely(
  hoursworked %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(hoursworkedperweek), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.hours = slopes.hours %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )


#Mark 4 SY LAs and plot
slopes.hours = slopes.hours %>% 
  ungroup() %>% 
  # mutate(
  #   SY_LAs = ifelse(qg('rotherh|doncaste|sheffie|barnsley', Region_name),'SY LA','Other'),
  #   corecities = ifelse(Region_name %in% corecities,'core city','Other'),
  #   london = ifelse(qg('london',Region_name), 'London','Other')
  # )
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'SY LA',#will overwrite sheffield as core city
      qg('rotherh|doncaste|sheffie|barnsley', Region_name) ~ 'SY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
      ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','SY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.hours = slopes.hours %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )




#Then GVA CV
slopes.gva = get_slope_and_se_safely(
  itl3.all %>% filter(year %in% c(startyear:endyear)), 
  Region_name,
  y = log(gva_cv), 
  x = year, 
  neweywest = F)

#Add 95% confidence intervals around the slope
slopes.gva = slopes.gva %>% 
  mutate(
    ci95min = slope - (se * 1.96),
    ci95max = slope + (se * 1.96)
  )



#Mark 4 SY LAs and plot
slopes.gva = slopes.gva %>% 
  ungroup() %>% 
  mutate(
    itl3label = case_when(
      # qg('york', Region_name) ~ 'SY LA',#will overwrite sheffield as core city
      qg('rotherh|doncaste|sheffie|barnsley', Region_name) ~ 'SY LA',#will overwrite sheffield as core city
      Region_name %in% corecities ~ 'Core city',
      Region_name %in% itl3.lon$ITL325NM ~ 'London',
      .default = 'Other'
    ),
    itl3label = factor(itl3label, ordered = T, levels = c('Other','London','Core city','SY LA'))
  )

#Adjust to get percentage change per year from log slopes
slopes.gva = slopes.gva %>% 
  mutate(
    across(c(slope,ci95min,ci95max), ~(exp(.) - 1) * 100, .names = '{.col}_percentperyear'),
    slopegroups = as.numeric(cut_number(slope_percentperyear,4))
  )


# table(slopes.hours$itl3label)
# table(slopes.gva$itl3label)


#Keep the bits we want and combine into single df
both <- slopes.gva %>% 
  select(Region_name,itl3label,
         slope_percentperyear_GVA = slope_percentperyear,
         ci95min_percentperyear_GVA = ci95min_percentperyear,
         ci95max_percentperyear_GVA = ci95max_percentperyear) %>% 
  inner_join(
    slopes.hours %>% 
      select(Region_name,
             slope_percentperyear_HOURS = slope_percentperyear,
             ci95min_percentperyear_HOURS = ci95min_percentperyear,
             ci95max_percentperyear_HOURS = ci95max_percentperyear),
    by = 'Region_name'
  )


#Poor control over visibility. Layer instead. Base plot then function.
addtogva_hours_plot = function(p,plotdata, showerrors = T, sizeval = 1, colourval = 'black', alphaval = 1){
  
  p = p + geom_point(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, group = Region_name), colour = colourval, size = sizeval, alpha = alphaval) 
  
  if(showerrors){
    
    p = p + geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, xmin = ci95min_percentperyear_HOURS, xmax = ci95max_percentperyear_HOURS),width = 0.1, size = 1, colour = colourval, alpha = alphaval) +
    geom_errorbar(data = plotdata, aes(x = slope_percentperyear_HOURS, y = slope_percentperyear_GVA, ymin = ci95min_percentperyear_GVA, ymax = ci95max_percentperyear_GVA),width = 0.1, size = 1, colour = colourval, alpha = alphaval) 
    
  }
  
    p  + geom_abline(slope = 1, intercept = 0, colour = 'black', alpha = 0.5, size = 2) +
      geom_hline(yintercept = 0, colour = 'black', alpha = 0.3, size = 1) +
      geom_vline(xintercept = 0, colour = 'black', alpha = 0.3, size = 1)
  
}




p = ggplot()

p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Other'), F, 1, 'black', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'London'), F, 2, 'red', 0.4)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'Core city'), T, 3, RColorBrewer::brewer.pal(3, "Set1")[2], 0.5)
p = addtogva_hours_plot(p, both %>% filter(itl3label == 'SY LA'), T, 7, RColorBrewer::brewer.pal(3, "Set1")[3], 1)

p = p + coord_fixed() +
  xlab("Average % change: hours worked per week") +
  ylab("Average % change: GVA") 


ggplotly(p, tooltip = 'Region_name')



```

