---
title: "Rotherham: Industrial Strategy data linkage (November 2025)"
author:
  - name: "Dan Olner"
format: 
  html:
    output-file: "Rotherham_IndStrat_Nov2025"
    fig-width: 8
    fig-height: 8
execute:
  echo: false
  message: false
  warning: false
fig-responsive: false
collapse: false
navbar: false
sidebar: false
---

```{r setup}
library(tidyverse)
# library(spdep)
library(sf)#for geo stuff
# library(tmap)#for mapping 
library(zoo)#for making smoothed moving averages
library(patchwork)#combine ggplots easily
library(plotly)#For interactive plots
# # library(ggdist)
# library(tidyr)
# library(distributional)
library(RColorBrewer)

source('../functions/misc_functions.R')
source('../functions/adhoc_functions.R')
#source('functions/misc_functions.R')
#source('functions/adhoc_functions.R')

options(scipen = 999)

#Set ggplot theme
theme_set(theme_light())

# Value for all job cutoffs
# minjobs = 99
minjobs = 1

greplsector = 'CREATIVE'

corecities = readRDS('../data/corecitiesvector.rds')

```

# Introduction

For the LQ analysis, two data sources with different strengths and weaknesses are used, to triangulate a fuller picture:

-   **Business Register and Employment Survey** (BRES) from 2015 to 2024, job counts. (2024 data was released 28th October.) The survey captures larger firms and their related sites, as well as smaller firms, but sample sizes become small at granular levels, so it can miss the full picture and is volatile between years.

-   **Companies House (CH) firm level data**, scraped for accounts submitted between September 2024 and end of September 2025. CH data's main weakness: it records only firms HQ'd in South Yorkshire. But it captures a sizeable proportion of economic activity, and can often indicate (comparing to BRES) which sectors are more 'home grown' (a high CH / BRES ratio) and sometimes pick up on jobs BRES misses, as it's not sample based. Here, it can also only compare job counts from when accounts were submitted and a year previously.

# UK INDUSTRIAL STRATEGY LINKAGE

The [UK Industrial Strategy](https://www.gov.uk/government/collections/the-uks-modern-industrial-strategy-2025) was released in 2025, with accompanying plans for 6 out of the 8 target "IS-8" sectors. It also has a [sector definition list](https://www.gov.uk/government/publications/industrial-strategy/industrial-strategy-sector-definitions-list) that provides SIC codes being targeted - top level sectors and 'frontier' subsectors (a few of which overlap; most do not). These range in specificity between 2 digit and 5 digit SIC codes.

As that webpage says, most of these SIC codes cannot capture the full range of possible firms the strategy aims to cover, and it also contains entire sections that SIC codes can't cover at all - clean energy and certain digital sectors in particular. However, it does allow an analysis of which sectors are *definitely* within those industrial strategy categories. The plots below gather job details for those specific sectors - a distinct range of types and breadth - and compares Salford to GB core cities, to see which sectors align with the industrial strategy and other analysis above. This is done for both BRES and Companies House job counts, to pick up on different job types.

The table below lists the IS-8 broad sector headings. In the IndStrat plots, the abbreviations for these IS-8 headings appear before the sector name. Clean energy - as mentioned above - has no SIC codes and so can't currently be matched. 

```{r brad_indstrattable}

indstrat <- tibble(
  `IS8 sector` = c(
    'Advanced Manufacturing',
    'Clean Energy Industries',
    'Creative Industries',
    'Defence',
    'Digital and Technologies', 
    'Financial Services',
    'Life Sciences', 
    'Professional and Business Services'),
  `Abbreviation` = c(
    'ADV MANUF',
    '--',
    'CREATIVE',
    'DEFENCE',
    'DIGITAL',
    'FIN',
    'LIFESCI',
    'PBS'
    )
)

#escape = FALSE so that markdown in the cells (the links) is rendered correctly.
knitr::kable(indstrat, escape = FALSE, caption = "Table: Industrial Strategy IS-8 names and abbreviations (dashes mean that sector doesn't appear for South Yorkshire)")

```




# LQ plots for Rotherham vs core cities

## LQ plot guide

The outputs below are **location quotient plots** for each of South Yorkshire's local authorities, for "creative" industrial strategy SIC codes.

Core cities are again included for comparison.

-   They are separated into industrial strategy 'frontier' sectors and the rest. Some are repeated, as the IndStrat definitions have two overlapping, nested categories (591 and 62011 are frontier sectors, but their parent sectors, 59 and 6201, are listed in the broader creative sectors; both are used here).

Here's a guide to reading them:

-   Location quotients for sectoral job proportions are shown for Salford with core cities overlaid
-   Sectors to the right of the centre line are more concentrated in this region when compared to the national average. (e.g. LQ = 10, it is 10x more concentrated)
-   Green circles have been increasing in concentration over time (larger circles increasing the most)
-   Red circles have been decreasing in concentration over time (larger circles decreasing the most)
-   Black horizontal lines show the min and max of concentration in the data’s history (from 2015 to 2023)
-   Circles and lines can be read together - for example, if a green circle is at the right of this line, sector job concentration has been increasing on average and is continuing.
-   Companies House plots are just comparing job counts when accounts are submitted to the 'last year's employees' figure from the same accounts - so they indicate how jobs have shifted since last accounts, but can't show longer-term change.
-   Per sector job counts and percent of the total local authority job count are shown on the right. These are useful to check sector size versus LQ - for example, a sector like jewellery can have a high LQ (as it does in Sheffield and Birmingham) with very low job counts, because there aren't many jobs in this sector overall in Great Britain.
-   Background squares are LQs for core cities. Core cities' first two letters are used. Here's a list of those, for reference:

```{r}

corecitytable <- tibble(
  `Core city` = c(
    "Belfast","Birmingham","Bristol","Cardiff","City of Edinburgh","Glasgow","Leeds","Liverpool","Manchester","Newcastle upon Tyne","Nottingham","Sheffield"),     
  `Abbreviation` = c(
    "Be","Bi","Br","Ca","Ed","Gl","Le","Li","Ma","Ne","No","Sh"
    )
)

#escape = FALSE so that markdown in the cells (the links) is rendered correctly.
knitr::kable(corecitytable, escape = FALSE, caption = "Table: core city abbreviations")

```

## LQs plots: frontier vs other sectors

# SALFORD + core cities

## "Frontier" SIC sectors

### BRES frontier sectors

```{r, fig-bresindstrat}

# islq = readRDS('../local/lqs_for_indstrat.rds') %>% filter(qg(greplsector,sic_namefrom_indstrat_combo)) %>% mutate(sic_namefrom_indstrat_combo = gsub('CREATIVE: ','',sic_namefrom_indstrat_combo))
# 
# yeartoplot = readRDS('../local/yeartoplots_for_indstrat.rds') %>% filter(qg(greplsector,sic_namefrom_indstrat_combo)) %>% mutate(sic_namefrom_indstrat_combo = gsub('CREATIVE: ','',sic_namefrom_indstrat_combo))

islq = readRDS('../local/lqs_for_indstrat.rds') 

yeartoplot = readRDS('../local/yeartoplots_for_indstrat.rds') 

place = "Rotherham"

sectorLQorder <- islq %>% filter(
  DATE == max(DATE),#use latest data
  GEOGRAPHY_NAME == place) %>% 
  arrange(-LQ) %>% 
  select(sic_namefrom_indstrat_combo) %>% 
  pull()

# Tick
# length(sectorLQorder)
# length(unique(sectorLQorder))

#Split by frontier or not

#Turn the sector column into a factor and order by LQs
yeartoplot$sic_namefrom_indstrat_combo <- factor(yeartoplot$sic_namefrom_indstrat_combo, levels = sectorLQorder, ordered = T)

#Remove some LQs
yeartoplot.lqfiltered <- yeartoplot %>% filter(LQ > 0 & LQ < 100, JOBCOUNT >= minjobs)

#Drop any sectors that Bradford now doesn't have after that filter
sectorstokeep <- yeartoplot.lqfiltered %>% filter(GEOGRAPHY_NAME == place) %>% select(sic_namefrom_indstrat_combo) %>% pull

yeartoplot.lqfiltered <- yeartoplot.lqfiltered %>% filter(sic_namefrom_indstrat_combo %in% sectorstokeep)

# Make base layer just of labelled core cities
# Remove sheffield also, don't want that labelled
corecities_baselayer = yeartoplot.lqfiltered %>% filter(GEOGRAPHY_NAME %in% corecities[!qg('sheffield',corecities)])

# Add in placename abbreviations
corecities_baselayer = corecities_baselayer %>% 
  mutate(
    placename_short = str_sub(GEOGRAPHY_NAME,1,2),
    placename_short = ifelse(placename_short == 'Ci','Ed',placename_short)#Edinburgh
  )

# debugonce(LQ_baseplot)
p <- LQ_baseplot(df = corecities_baselayer %>% filter(is_frontier == 1), alpha = 0.8, shape = 0, sector_name = sic_namefrom_indstrat_combo, LQ_column = LQ, change_over_time = slope, labelcolumn = placename_short)

p <- addplacename_to_LQplot(df = yeartoplot.lqfiltered %>% filter(is_frontier == 1), plot_to_addto = p, maxLQvalmultiplier = 2,
                            placename = place, shapenumber = 16,
                            min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                            value_column = JOBCOUNT, sector_regional_proportion = sector_regional_proportion,
                            region_name = GEOGRAPHY_NAME,
                            sector_name = sic_namefrom_indstrat_combo, change_over_time = slope, LQ_column = LQ,
                            text = 7, value_col_ismoney = F)

p <- p + 
  # coord_cartesian(xlim = c(0.1,7)) +
  ggtitle("IndStrat frontier sector")

p


```

### Companies house frontier sectors

```{r, fig-chindstrat}

ch_summary = readRDS('../local/CH_lqs_for_indstrat.rds') 

yeartoplot.ch = readRDS('../local/CH_yeartoplots_for_indstrat.rds') 

# ch_summary = readRDS('../local/CH_lqs_for_indstrat.rds') %>% 
#   filter(
#     qg(greplsector,sic_namefrom_indstrat_combo),
#     !is.na(sic_namefrom_indstrat_combo)
#     ) %>% mutate(sic_namefrom_indstrat_combo = gsub('CREATIVE: ','',sic_namefrom_indstrat_combo))
# 
# yeartoplot.ch = readRDS('../local/CH_yeartoplots_for_indstrat.rds') %>%
#   filter(
#     qg(greplsector,sic_namefrom_indstrat_combo),
#     !is.na(sic_namefrom_indstrat_combo)
#     ) %>% mutate(sic_namefrom_indstrat_combo = gsub('CREATIVE: ','',sic_namefrom_indstrat_combo))

sectorLQorder <- ch_summary %>% filter(
  GEOGRAPHY_NAME == place,
  DATE == max(DATE)#use latest data
) %>% 
  arrange(-LQ) %>% 
  select(sic_namefrom_indstrat_combo) %>% 
  pull()


#Turn the sector column into a factor and order by LCR's LQs
yeartoplot.ch$sic_namefrom_indstrat_combo <- factor(yeartoplot.ch$sic_namefrom_indstrat_combo, levels = sectorLQorder, ordered = T)

#Remove some LQs
yeartoplot.lqfiltered.ch <- yeartoplot.ch %>% filter(LQ > 0 & LQ < 100, JOBCOUNT >= minjobs)


#Drop any sectors that Bradford now doesn't have after that filter
sectorstokeep <- yeartoplot.lqfiltered.ch %>% filter(GEOGRAPHY_NAME == place) %>% select(sic_namefrom_indstrat_combo) %>% pull
yeartoplot.lqfiltered.ch <- yeartoplot.lqfiltered.ch %>% filter(sic_namefrom_indstrat_combo %in% sectorstokeep)

# Make base layer just of labelled core cities
# Remove sheffield also, don't want that labelled
corecities_baselayer.ch = yeartoplot.lqfiltered.ch %>% filter(GEOGRAPHY_NAME %in% corecities[!qg('sheffield',corecities)])

corecities_baselayer.ch = corecities_baselayer.ch %>% 
  mutate(
    placename_short = str_sub(GEOGRAPHY_NAME,1,2),
    placename_short = ifelse(placename_short == 'Ci','Ed',placename_short)#Edinburgh
  )

p1 <- LQ_baseplot(df = corecities_baselayer.ch %>% filter(is_frontier == 1), alpha = 0.8, shape = 0, sector_name = sic_namefrom_indstrat_combo, LQ_column = LQ, change_over_time = slope, labelcolumn = placename_short)
# p1 <- LQ_baseplot(df = yeartoplot.lqfiltered.ch %>% filter(is_frontier == 1), alpha = 0.03, sector_name = sic_namefrom_indstrat_combo, 
#                   LQ_column = LQ, change_over_time = slope)

# debugonce(addplacename_to_LQplot)
p1 <- addplacename_to_LQplot(df = yeartoplot.lqfiltered.ch %>% filter(is_frontier == 1), plot_to_addto = p1, maxLQvalmultiplier = 0.5,
                             placename = place, shapenumber = 16,
                             # min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                             value_column = JOBCOUNT, sector_regional_proportion = sector_regional_proportion,
                             region_name = GEOGRAPHY_NAME,
                             sector_name = sic_namefrom_indstrat_combo, change_over_time = slope, LQ_column = LQ,
                             value_col_ismoney = F, text = 7)
                             # value_col_ismoney = F, text = 7, maxLQvalmultiplier = 2,useplacenameforminmaxdisplay = T, overridetextpos = 14)

p1 <- p1 + 
  # coord_cartesian(xlim = c(0.1,7)) +
  ggtitle("IndStrat frontier sector")

p1


```

## Other industrial strategy SIC sectors

### BRES other sectors

```{r, fig-bresindstrat2}
#| fig-height: 13


p <- LQ_baseplot(df = corecities_baselayer %>% filter(is_frontier == 0), alpha = 0.8, shape = 0, sector_name = sic_namefrom_indstrat_combo, LQ_column = LQ, change_over_time = slope, labelcolumn = placename_short)

p <- addplacename_to_LQplot(df = yeartoplot.lqfiltered %>% filter(is_frontier == 0), plot_to_addto = p, maxLQvalmultiplier = 0.5,
                            placename = place, shapenumber = 16,
                            min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                            value_column = JOBCOUNT, sector_regional_proportion = sector_regional_proportion,
                            region_name = GEOGRAPHY_NAME,
                            sector_name = sic_namefrom_indstrat_combo, change_over_time = slope, LQ_column = LQ,
                            text = 7, value_col_ismoney = F)

p <- p + 
  # coord_cartesian(xlim = c(0.1,7)) +
  ggtitle("IndStrat other sector")

p


```

### Companies House other sectors

```{r, fig-chindstrat2}
#| fig-height: 13

p2 <- LQ_baseplot(df = corecities_baselayer.ch %>% filter(is_frontier == 0), alpha = 0.8, shape = 0, sector_name = sic_namefrom_indstrat_combo, LQ_column = LQ, change_over_time = slope, labelcolumn = placename_short)
# p2 <- LQ_baseplot(df = yeartoplot.lqfiltered.ch %>% filter(is_frontier == 0), alpha = 0.03, sector_name = sic_namefrom_indstrat_combo, 
#                   LQ_column = LQ, change_over_time = slope)

# debugonce(addplacename_to_LQplot)
p2 <- addplacename_to_LQplot(df = yeartoplot.lqfiltered.ch %>% filter(is_frontier == 0), plot_to_addto = p2, maxLQvalmultiplier = 0.5,
                             placename = place, shapenumber = 16,
                             # min_LQ_all_time = min_LQ_all_time,max_LQ_all_time = max_LQ_all_time,#Include minmax
                             value_column = JOBCOUNT, sector_regional_proportion = sector_regional_proportion,
                             region_name = GEOGRAPHY_NAME,
                             sector_name = sic_namefrom_indstrat_combo, change_over_time = slope, LQ_column = LQ,
                             value_col_ismoney = F, text = 7)
                             # value_col_ismoney = F, text = 7, maxLQvalmultiplier = 2,useplacenameforminmaxdisplay = T, overridetextpos = 14)

p2 <- p2 + 
  # coord_cartesian(xlim = c(0.1,7)) +
  ggtitle("IndStrat other sector")

p2



```
