---
title: "South Yorkshire BRES job statistics 2015-2024"
author:
  - name: "Dan Olner"
format: 
  html:
    output-file: "SouthYorkshire_BRESjobstats_201524"
    fig-width: 8
    fig-height: 8
execute:
  echo: false
  message: false
  warning: false
fig-responsive: false
collapse: false
navbar: false
sidebar: false
---

```{r setup}
library(tidyverse)
library(zoo)#for making smoothed moving averages
library(RColorBrewer)
library(kableExtra)


source('../functions/misc_functions.R')

options(scipen = 999)

bres = readRDS('../local/data/bres_sy2024.rds')


```

## Notes on the data

This is:

- [BRES](https://www.nomisweb.co.uk/datasets/newbres6pub) full time job count data at local authority level for 2015 to 2024
- 2 digit SIC sectors are used (shortened names)
- Keeping only sectors with more than 1% of each local authority's total job count
- The 2022-24 average jobcount is included as job data can be volatile; 'jobs2024' is the latest jobcount datapoint in BRES.
- Tables are ordered by location quotient (LQ) to show which sectors are more concentrated in each local authority. E.g. Sheffield's top LQ sector is basic metals: its LQ of 5.13 = basic metals job count there is 5.1x higher than the Great Britain average. (BRES data is GB not UK.)
- Compound Annual Growth Rate figures (just the year on year percent change taken from line-slopes through the data). Two are included: for the full 2015-2024 range, and for 2021-2024. (These use all the data, not smoothed).


## Sheffield, sectors with more than 1% of jobs

```{r}

#escape = FALSE so that markdown in the cells (the links) is rendered correctly.
# knitr::kable(
#   bres %>% filter(
#     regional_percent > 2,
#     GEOGRAPHY_NAME == 'Sheffield'
#     ) %>% 
#     select(SIC2dig_shortnames,`jobs (22-24 av)` = jobcount_movingav, regional_percent, LQ, `CAGR 15-24`, `CAGR 21-24`) %>% 
#   arrange(-LQ), 
#   escape = FALSE, caption = "")

table_data <- bres %>% 
  filter(
    regional_percent > 1,
    GEOGRAPHY_NAME == 'Sheffield'
  ) %>% 
  select(
    SIC = SIC2dig_shortnames,
    `jobs (22-24 av)` = jobcount_movingav,
    jobs2024,
    `sector %` = regional_percent, 
    LQ, 
    `CAGR 15-24`, 
    `CAGR 21-24`
  ) %>% 
  arrange(-LQ) %>%
  mutate(
    `CAGR 15-24` = cell_spec(`CAGR 15-24`, color = ifelse(`CAGR 15-24` < 0, "red", "black")),
    `CAGR 21-24` = cell_spec(`CAGR 21-24`, color = ifelse(`CAGR 21-24` < 0, "red", "black")),
    `jobs (22-24 av)` = round(`jobs (22-24 av)`,0),
    `sector %` = round(`sector %`,1),
    LQ = round(LQ,2)
  )

knitr::kable(table_data, escape = FALSE, caption = "")


```
## Rotherham, sectors with more than 1% of jobs

```{r}

#escape = FALSE so that markdown in the cells (the links) is rendered correctly.
# knitr::kable(
#   bres %>% filter(
#     regional_percent > 2,
#     GEOGRAPHY_NAME == 'Sheffield'
#     ) %>% 
#     select(SIC2dig_shortnames,`jobs (22-24 av)` = jobcount_movingav, regional_percent, LQ, `CAGR 15-24`, `CAGR 21-24`) %>% 
#   arrange(-LQ), 
#   escape = FALSE, caption = "")

table_data <- bres %>% 
  filter(
    regional_percent > 1,
    GEOGRAPHY_NAME == 'Rotherham'
  ) %>% 
  select(
    SIC = SIC2dig_shortnames,
    `jobs (22-24 av)` = jobcount_movingav,
    jobs2024,
    `sector %` = regional_percent, 
    LQ, 
    `CAGR 15-24`, 
    `CAGR 21-24`
  ) %>% 
  arrange(-LQ) %>%
  mutate(
    `CAGR 15-24` = cell_spec(`CAGR 15-24`, color = ifelse(`CAGR 15-24` < 0, "red", "black")),
    `CAGR 21-24` = cell_spec(`CAGR 21-24`, color = ifelse(`CAGR 21-24` < 0, "red", "black")),
    `jobs (22-24 av)` = round(`jobs (22-24 av)`,0),
    `sector %` = round(`sector %`,1),
    LQ = round(LQ,2)
  )

knitr::kable(table_data, escape = FALSE, caption = "")


```
## Barnsley, sectors with more than 1% of jobs

```{r}

#escape = FALSE so that markdown in the cells (the links) is rendered correctly.
# knitr::kable(
#   bres %>% filter(
#     regional_percent > 2,
#     GEOGRAPHY_NAME == 'Sheffield'
#     ) %>% 
#     select(SIC2dig_shortnames,`jobs (22-24 av)` = jobcount_movingav, regional_percent, LQ, `CAGR 15-24`, `CAGR 21-24`) %>% 
#   arrange(-LQ), 
#   escape = FALSE, caption = "")

table_data <- bres %>% 
  filter(
    regional_percent > 1,
    GEOGRAPHY_NAME == 'Barnsley'
  ) %>% 
  select(
    SIC = SIC2dig_shortnames,
    `jobs (22-24 av)` = jobcount_movingav,
    jobs2024,
    `sector %` = regional_percent, 
    LQ, 
    `CAGR 15-24`, 
    `CAGR 21-24`
  ) %>% 
  arrange(-LQ) %>%
  mutate(
    `CAGR 15-24` = cell_spec(`CAGR 15-24`, color = ifelse(`CAGR 15-24` < 0, "red", "black")),
    `CAGR 21-24` = cell_spec(`CAGR 21-24`, color = ifelse(`CAGR 21-24` < 0, "red", "black")),
    `jobs (22-24 av)` = round(`jobs (22-24 av)`,0),
    `sector %` = round(`sector %`,1),
    LQ = round(LQ,2)
  )


knitr::kable(table_data, escape = FALSE, caption = "")


```
## Doncaster, sectors with more than 1% of jobs

```{r}

#escape = FALSE so that markdown in the cells (the links) is rendered correctly.
# knitr::kable(
#   bres %>% filter(
#     regional_percent > 2,
#     GEOGRAPHY_NAME == 'Sheffield'
#     ) %>% 
#     select(SIC2dig_shortnames,`jobs (22-24 av)` = jobcount_movingav, regional_percent, LQ, `CAGR 15-24`, `CAGR 21-24`) %>% 
#   arrange(-LQ), 
#   escape = FALSE, caption = "")

table_data <- bres %>% 
  filter(
    regional_percent > 1,
    GEOGRAPHY_NAME == 'Doncaster'
  ) %>% 
  select(
    SIC = SIC2dig_shortnames,
    `jobs (22-24 av)` = jobcount_movingav,
    jobs2024,
    `sector %` = regional_percent, 
    LQ, 
    `CAGR 15-24`, 
    `CAGR 21-24`
  ) %>% 
  arrange(-LQ) %>%
  mutate(
    `CAGR 15-24` = cell_spec(`CAGR 15-24`, color = ifelse(`CAGR 15-24` < 0, "red", "black")),
    `CAGR 21-24` = cell_spec(`CAGR 21-24`, color = ifelse(`CAGR 21-24` < 0, "red", "black")),
    `jobs (22-24 av)` = round(`jobs (22-24 av)`,0),
    `sector %` = round(`sector %`,1),
    LQ = round(LQ,2)
  )


knitr::kable(table_data, escape = FALSE, caption = "")


```

