---
title: "Intro to GVA and jobs data in R"
code-overflow: wrap
execute:
  echo: true
  warning: false
  error: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 8, fig.width = 8, echo = TRUE, warning = F, error = F, message = F, comment=NA)
```

# Intro to using ONS GVA and jobs data in R

Here, we'll work with this pre-baked ONS data [from here](regionalgva_n_bres.html), which has been processed/linked to help smooth the path to getting GVA and jobs insights. As that page says, we've got the following to work with:

* GVA data at ITL2 and 3 geographies, broken down by SIC industrial codes at 2-digit, Section (20 categories) and Production/Construction/Services, 1998 to most recent available year.
* BRES job count data for the same (though with slightly more SIC codes), with a little bit better job count resolution than the originals (see above post for why/how).
* A linked version that combines those two, so we can dig into GVA per job across Great Britain[^1].

[^1]: NOMIS BRES data is not available for Northern Ireland.

These are both deep, rich datasets (though it's worth exploring [how the reginal GVA data is made](https://www.ons.gov.uk/economy/regionalaccounts/grossdisposablehouseholdincome/methodologies/regionalaccountsmethodologyguidejune2019) to get a sense of its issues). Here, we'll just run through a range of examples to show how to get started with digging into them. See the left hand bar for other ways to use them (and do please let me know if you have any thoughts, other analyses etc, via the [github issues](https://github.com/DanOlner/RegionalEconomicTools/issues) or email danolner at gmail dot com).

Here's some libraries and functions inside this repo we'll use.

```{r libs}
library(tidyverse)
library(sf)#for geo stuff
library(tmap)#for mapping 
library(zoo)#for making smoothed moving averages
library(patchwork)#combine ggplots easily

source('functions/misc_functions.R')

```



# How do production vs services vary across the UK?

Let's start with the broadest sector breakdown, and at ITL3 geography, which includes most major cities and local authorities (though some are combined into one zone, like Barnsley, Doncaster & Rotherham). The link is nabbed from the [gva n bres page](https://danolner.github.io/RegionalEconomicTools/regionalgva_n_bres.html).

**The code will also work for ITL2 zones using this data (mayoral authorities are at this level) - just [use this CSV](https://raw.githubusercontent.com/DanOlner/RegionalEconomicTools/refs/heads/gh-pages/data/regionalGVA/regionalGVA_currentprices_ITL2_SIC_3GROUPS_LONG_2022.csv) for ITL3 instead.**

We use **current prices** data - GVA values *as they were when the data was collected*, not adjusted for inflation. This is useful because (unlike inflation adjusted chained volume data), these can be legitimately summed in any way like - add up sectors, places etc. This means any change over time for current prices is only ever **relative**. (Examples below will illustrate this.)

Also, sum the two 'production' and 'construction' sectors to give us just two 'goods v services' categories.



```{r data}
#Load 3 broad sector data for ITL2 and current prices
itl3.ps <- read_csv('https://raw.githubusercontent.com/DanOlner/RegionalEconomicTools/refs/heads/gh-pages/data/regionalGVA/regionalGVA_currentprices_ITL3_SIC_3GROUPS_LONG_2022.csv')

#sum production and construction
itl3.ps <- itl3.ps %>% 
  mutate(SIC07_description = fct_collapse(SIC07_description, `Goods sector` = c("Production sector","Construction"))) %>% 
  group_by(Region_name, year, SIC07_description) %>% 
  summarise(
    value = sum(value),
    ITL_code = max(ITL_code)#This just keeps a single ITL code, they're all the same in this group. We need it to match on later as names don't perfectly match
  ) %>% 
  ungroup()


```

## Looking at location quotients

Now let's use **location quotients** (LQs) to see where goods and services are most concentrated across the UK. See [this post](sector_locationquotients_and_proportions.html) for a full explanation of LQs and a deeper dive into them - but briefly:

* If the LQ > 1, that sector is relatively **more** concentrated in the region, compared to the UK
* If the LQ < 1, it's relatively **less** concentrated in the region, compared to the UK
* ... and the value says by how much e.g. an LQ of 2 would mean it's twice as concentrated.

LQs also have a handy feature:

* As the [ONS Excel sheet on LQs](https://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/employmentandemployeetypes/datasets/locationquotientdataandindustrialspecialisationforlocalauthorities) make really clear, because (A/B)/(C/D) is equivalent to (A/C)/(B/D), the LQ actually captures two related ways of seeing the same thing: how relatively concentrated sectors are *across a whole geography* like the UK, and how concentrated *within a subgeography* like South Yorkshire they are. (See the table in the ONS document - numbers which can be read either across geographies or across sectors.)

(Note also: LQs **only make sense for current prices data**. Chained volume data can't be summed, as each row is inflation-adjusted invididually and e.g. ITL3 zones or sectors don't sum to larger geographies or SIC groupings because of that. LQs need to sum sectors and places to correctly work out proportions.)

We'll use that in a moment. First, using a function loaded above, this adds LQ details to the loaded CSV:

```{r}
lq.latestyear <- add_location_quotient_and_proportions(
  itl3.ps %>% filter(year == max(year)),#use the latest year in the data
  regionvar = Region_name,
  lq_var = SIC07_description,
  valuevar = value
)

```

Then  - ordering ITL3 zones by which has the highest goods-sector concentrations, comparing different places. Here, just outputting the ordered data and picking the top 15.


```{r}
lq.latestyear %>% filter(
  SIC07_description == 'Goods sector'
  ) %>% 
  mutate(regional_percent = sector_regional_proportion *100) %>% 
  select(Region_name,regional_percent, LQ) %>% 
  arrange(-LQ) %>% 
  slice(1:15)
```

Services is of course the mirror image of this, and we could list them - but taking a look at a **map of goods sector LQ** is also a good way of seeing what's going on. A couple of notes:

* Log LQ is used as this keeps it symmetrical each side of zero, which makes the map a lot more legible (with the downside that log values aren't as interpretable).
* Zones with LQ < 1 for goods are highlighted in thicker black lines - making clear that **more service-heavy places tend to be towns and cities (and a swathe of London and the South East)**.
* Plotting an interactive map here so it's easier to poke around / zoom in - change tmap_mode to 'plot' to just output as an image.



## Making a map

```{r}
#Load ITL3 map geodata
itl3.geo <- st_read('data/ITL_geographies/International_Territorial_Level_3_January_2021_UK_BUC_V3_2022_6920195468392554877', quiet = T) 

#Check geo and data row matches... tick
#Names don't quite match between the two, random commas etc
# table(itl3.geo$ITL321CD %in% itl3.ps$ITL_code)

#Join map data to a subset of the GVA data
LQ_map <- itl3.geo %>% 
  right_join(
    lq.latestyear %>% filter(
      SIC07_description == 'Goods sector'
      ),
    by = c('ITL321CD'='ITL_code')
  )

tmap_mode('view')#Use 'plot' to just plot an image

#Make map for interactive viewing
tm_shape(LQ_map) +
  tm_polygons('LQ_log', n = 9, id = 'Region_name', fill.scale = tm_scale(values = "RdBu")) +
tm_shape(LQ_map %>% filter(LQ < 1)) + #Highlight LQ < 1 places by overlaying thicker lines
  tm_polygons(lwd = 2, alpha = 0, border.col = 'black', , id = 'Region_name') +
  tm_layout(title = 'LQ spread of\nGoods sector\nAcross ITL3 regions', legend.outside = T) 

```






# How have the proportion of goods vs services changed over time?

What's been happening to goods vs service sectors across the UK since this dataset's earliest date in 1998? We can examine the linear slopes to get a rough idea of what direction trends have been going in.

First, let's re-run the LQ function for every year in the dataset:

```{r lqall}
itl3.ps <- itl3.ps %>% 
  split(.$year) %>% 
  map(add_location_quotient_and_proportions, 
      regionvar = Region_name,
      lq_var = SIC07_description,
      valuevar = value) %>% 
  bind_rows()

```

Then use another function that finds OLS slopes of change over time (not great time series statistics, useful for a quick look). We use the "sector regional proportion" value that came with the LQ function (this sums to one for all sectors in a place) and log values so places with different sector scales have comparable slopes (e.g. a 4-5% increase appears the same as 40-50%):


## Using linear slopes to examine biggest growers and shrinkers

```{r LQ_slopes}
#Use
#LQ_slopes %>% filter(slope==0)
#To see which didn't get slopes (all got slopes here but sometimes there'll be missing values)
LQ_slopes <- compute_slope_or_zero(
  data = itl3.ps, 
  Region_name, SIC07_description,#slopes will be found within whatever grouping vars are added here
  y = log(sector_regional_proportion), x = year
  # y = LQ_log, x = year
  ) %>% left_join(
    itl3.ps %>% select(Region_name,ITL_code) %>% distinct(), by = 'Region_name'#Add ITL code back in, so we can link to geodata below
  ) %>% 
  arrange(-slope) %>% 
  mutate(approx_percentchange_peryear = slope * 100)


```

Using that, we can ask: where does it look like the proportion of services sectors has grown the most? Just sorting by slope size:

```{r}
LQ_slopes %>% filter(
  SIC07_description == 'Services sector'
  ) %>% 
  select(Region_name,approx_percentchange_peryear) %>% 
  arrange(-approx_percentchange_peryear) %>% 
  slice(1:15)

```


We can make a map again, looking at what percentage change per year, on average, the slopes look like across the UK. There's no obvious pattern to places where services have shrunk as a proportion of their economies - but as with the list above, East Antrim is an outlier. We'll see in a second how much work "as a proportion" is doing...

(Note, not an interactive map as it appears tmap/leaflet can only display one per page at the moment - it's changed back to 'plot' mode).

```{r}
slopes_map <- itl3.geo %>% 
  right_join(
    LQ_slopes %>% filter(
      SIC07_description == 'Services sector'
      ),
    by = c('ITL321CD'='ITL_code')
  )


tmap_mode('plot')

#Make map
tm_shape(slopes_map) +
  tm_polygons('approx_percentchange_peryear', id = 'Region_name', fill.scale = tm_scale(values = "RdBu")) +
  tm_layout(title = 'Services sector approx. % GVA change per year, ITL3 regions', legend.outside = T) 



```


## Plotting time series for places where services grew/shrunk the most

Let's take a look at the actual datapoints for 'percent of regional economy that's services', using the slopes data to extract the top (highest proportional growth) and bottom (highest prop shrinkage) slopes.

* First, get the place names for top and bottom
* Then in one ggplot for each of those, just look at the 1998 to latest datapoints
* Using the excellent [patchwork library](https://patchwork.data-imaginist.com/) to easily combine and title these two plots.


```{r}
#Has already been ordered by slope, so can just filter and pull out
#We just want the place names, to then plot their slopes
topITL3names <- LQ_slopes %>%
  filter(SIC07_description == "Services sector") %>%
  slice_head(n = 12) %>% 
  select(Region_name) %>% 
  pull
  
bottomITL3names <- LQ_slopes %>%
  filter(SIC07_description == "Services sector") %>%
  slice_tail(n = 12) %>% 
  select(Region_name) %>% 
  pull
  


top.plot <- ggplot(itl3.ps %>%  
         mutate(regional_percent = sector_regional_proportion *100) %>% 
         filter(
           Region_name %in% topITL3names,
           SIC07_description == "Services sector"
           ), 
       aes(x = year, y = regional_percent, colour = fct_reorder(Region_name,-sector_regional_proportion) )) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = 'Paired', direction = 1) +
  ylab('Regional GVA percent') +
  theme(plot.title = element_text(face = 'bold')) +
  labs(colour = 'ITL region')


bottom.plot <- ggplot(itl3.ps %>%  
         mutate(regional_percent = sector_regional_proportion *100) %>% 
         filter(
           Region_name %in% bottomITL3names,
           SIC07_description == "Services sector"
           ),
       aes(x = year, y = regional_percent, colour = fct_reorder(Region_name,-sector_regional_proportion) )) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = 'Paired', direction = 1) +
  ylab('Regional GVA percent') +
  theme(plot.title = element_text(face = 'bold')) +
  labs(colour = 'ITL region')


top.plot / bottom.plot + plot_annotation(
  title = 'Percent change in proportion of service sectors,1998-2022, top/bottom ITL3 zones'
)
  

```


Blimey - a couple of places really jump out from each plot, both inflecting around 2016. **Mid and East Antrim** in the 'top proportion growers' jumps hugely. **Southampton** equally has a massive drop - never far from a mid-80% service economy, it drops to below 70% before bouncing back a little. **North / northeast Lincolnshire** also sees a large drop - and appeared previously at the top of the "% goods sectors" list, with nearly 3 times higher proportion than the UK as a whole.

Let's look at each of those, using different approaches and datasets.



## Mid and East Antrim

Mid/East Antrim's apparently huge services jump highlights why *proportional* changes need to be treated carefully. What's happening underneath that? Did services grow - or did goods and production drop, making services a bigger percent of Mid/East Antrim's economy?

Let's look at the **chained volume** data to find out; this will let us see actual sector volume changes over time, adjusted for inflation.


```{r}
itl3.cv <- read_csv('https://raw.githubusercontent.com/DanOlner/RegionalEconomicTools/refs/heads/gh-pages/data/regionalGVA/regionalGVA_chainedvolume_ITL3_SIC_SECTION_LONG_2022.csv')


ggplot(itl3.cv %>%  
         filter(Region_name == "Mid and East Antrim"), 
       aes(x = year, y = value, colour = fct_reorder(SIC07_description,-value) )) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = 'Paired', direction = 1) +
  ylab('SIC section GVA (chained volume) ') +
  theme(plot.title = element_text(face = 'bold')) +
  labs(colour = 'SIC section')



```











## Southampton's GVA vs job change over time


```{r}
gva.jobs.ITL3 <- read_csv('https://raw.githubusercontent.com/DanOlner/RegionalEconomicTools/refs/heads/gh-pages/data/regionalGVA_plus_BRESjobcounts/regionalGVA_plus_BRESjobcounts_chainedvolume_ITL3_SIC_SECTION_MINUSimputedrent_2022.csv')

smoothband = 3

gva.jobs.ITL3 <- gva.jobs.ITL3 %>% 
  mutate(gvaperjobFT = gva/JOBCOUNT_FULLTIME) %>% 
group_by(GEOGRAPHY_NAME,SIC07_description) %>%
  mutate(
  jobcountFT_movingav = rollapply(JOBCOUNT_FULLTIME,smoothband,mean,align='center',fill=NA),
  gva_movingav = rollapply(gva,smoothband,mean,align='center',fill=NA),
  `gva/FT moving av` = rollapply(gvaperjobFT * 1000,smoothband,mean,align='center',fill=NA)
) %>% 
  ungroup()



```





```{r wiggleplot}
#| fig-width: 11
#| fig-height: 11


p <- twod_generictimeplot_multipletimepoints(
  df = gva.jobs.ITL3 %>% filter(
    qg('southampton',GEOGRAPHY_NAME), 
    !qg('agri|constr',SIC07_description),
    jobcountFT_movingav > 2500
    ),
  category_var = SIC07_description,
  # x_var = gva,
  # y_var = JOBCOUNT_FULLTIME,
  # label_var = gvaperjobFT,
  x_var = gva_movingav,
  y_var = jobcountFT_movingav,
  label_var = `gva/FT moving av`,
  timevar = DATE,
  times = c(2016:2021) 
)

p + theme(aspect.ratio=1) +
  # scale_x_log10() +
  # scale_y_log10() +
  xlab("GVA (3 year moving average)") +
  ylab("Job count FT (3 year moving average)")

```



















```{r cuttinz}
#| echo: false
#| eval: false

#Chained volume for services
#Load 3 broad sector data for ITL2 and current prices
itl3.ps <- read_csv('https://raw.githubusercontent.com/DanOlner/RegionalEconomicTools/refs/heads/gh-pages/data/regionalGVA/regionalGVA_chainedvolume_ITL3_SIC_3GROUPS_LONG_2022.csv')


#Hide smaller sectors in S'hampton
#Use proportion sector size from the LQ data
itl3.sections <- read_csv('https://raw.githubusercontent.com/DanOlner/RegionalEconomicTools/refs/heads/gh-pages/data/regionalGVA/regionalGVA_currentprices_ITL3_SIC_SECTION_LONG_2022.csv') %>% 
   split(.$year) %>% 
  map(add_location_quotient_and_proportions, 
      regionvar = Region_name,
      lq_var = SIC07_description,
      valuevar = value) %>% 
  bind_rows()

sectorstokeep <- itl3.sections %>% 
  filter(
    Region_name == 'Southampton',
    sector_regional_proportion > 0.05
    ) %>% 
  select(SIC07_description) %>% 
  distinct() %>% pull

```






