---
title: "Pipeline"
format: docx
editor: visual
execute:
  echo: false
  message: false
  warning: false
fig-dpi: 300
fig-width: 8
fig-height: 5
---

```{r setup}
#Load some functions and install some packages
source('https://bit.ly/pipelinesetup')

library(tidyverse)
library(nomisr)
library(xml2)

#Set ggplot theme
theme_set(theme_light())

```

## Introduction

We'll look at 3 examples of piping data through from web to this output. Starting with...




## 1. Extracting from an Excel file on a website: productivity data example

We can copy the URL for the Excel sheet from the web page...

```{r excel}
# 1. Extracting from an Excel file on a website: productivity data example

# Copy the URL for the Excel sheet from the ONS website
# This is just setting a string named 'url'
url <- 'https://www.ons.gov.uk/file?uri=/employmentandlabourmarket/peopleinwork/labourproductivity/datasets/subregionalproductivitylabourproductivitygvaperhourworkedandgvaperfilledjobindicesbyuknuts2andnuts3subregions/current/labourproductivityitls1.xlsx'

# We then create a TEMPORARY FILE...
temp_path <- tempfile(fileext=".xlsx")
# And use that file path to DOWNLOAD A TEMP COPY OF THE EXCEL TO USE
download.file(url, temp_path, mode="wb")

# Or if you want to keep a local, month-dated version in case it changes later...
# download.file(
#   url,
#   paste0('ONS_outputperhourdata-',format(Sys.Date(),'%b-%Y'),'.xlsx'),
#   mode="wb")

# READ FROM THE EXCEL SHEET
# This goes straight into a dataframe
# Read directly from a specific sheet and range of cells from the downloaded file
# "Table A1: Current Price (smoothed) GVA (B) per hour worked indices; ITL2 and ITL3 subregions, 2004 - 2023"
cp.indices <- readxl::read_excel(path = temp_path,range = "A1!A5:W247")

# TAKE A LOOK AT THAT DATA (CLICK ON ITS NAME IN THE ENVIRONMENT PANEL)
# It's got each year IN ITS OWN COLUMN. Not what we want.

# So we RESHAPE
# Get just ITL3 level and pivot years into their own column

# An aside: 
# This symbol - %>% 
# Is a PIPING OPERATOR
# It lets us pipe one operation to the next
# So we can clearly see each step taken
# Let's number those steps here

#1. Pipe the data into a filter
#2. filter down to just ITL3 zones
#3. pivot long so years are in their own column
#4. Convert year to a sensible format
#5. Remove the ITL column, we don't need that now
# And back on the first line, overwrite the original dataframe
cp.indices = cp.indices %>% 
  filter(ITL_level == 'ITL3') %>% 
  pivot_longer(Index_2004:Index_2023, names_to = 'year', values_to = 'index_value') %>% 
  mutate(
    year = substr(year,7,10),
    year = as.numeric(year)
  ) %>% 
  select(-ITL_level)


# We're going to plot JUST THE UK CORE CITIES
# Here's a list of those I made earlier...
corecities = readRDS(gzcon(url('https://github.com/DanOlner/RegionalEconomicTools/raw/refs/heads/gh-pages/data/corecitiesvector_itl3_2025.rds')))

# Check match...
# Note, this is commented out (so it doesn't output in our final doc)
# BUT we can still run it to check by highlighting the code
# table(corecities %in% cp.indices$Region_name)

# Plot!
# Filter down to the core cities
# And reorder the places by which has the highest average index over all years
ggplot(
    cp.indices %>% filter(Region_name %in% corecities),
    aes(x = year, y = index_value, colour = fct_reorder(Region_name,index_value,.desc = T))
    ) +
  geom_line() +
  geom_hline(yintercept = 100) +
  scale_color_brewer(palette = 'Paired') +
  theme(legend.title=element_blank()) +
  ylab('Output per hour: 100 = UK av')
```




## 2. Using the NOMISR package to extract data through the NOMIS API

```{r nomis}

# Get the BRES data we extracted in the separate script
bres = readRDS('bres_nomisdatasave.rds')

# Check its size in memory
#pryr::object_size(bres)

# Quick looksee
# bres %>% filter(INDUSTRY_TYPE == 'SIC 2007 division (2 digit)') %>% View

# Check industry percentages sum to 100 per ITL3 zone
# Mostly - some roundings-down build up to not quite 100%
# bres %>% 
#   filter(INDUSTRY_TYPE == 'SIC 2007 division (2 digit)') %>% 
#   group_by(GEOGRAPHY_NAME,DATE) %>% 
#   summarise(totalpercent = sum(OBS_VALUE))

# OK, let's look at the three finance 2-digit sectors
# 2 digit SICS are 64,65 and 66: sum those
# Pick out just 2 digit first as well
finance.percent = bres %>% 
  filter(
    INDUSTRY_TYPE == 'SIC 2007 division (2 digit)',
    qg('64|65|66', INDUSTRY_NAME)
  ) %>% 
  group_by(GEOGRAPHY_NAME,DATE) %>% 
  summarise(
    percent = sum(OBS_VALUE)
  ) 

# Plot. May need to smooth
ggplot(
  finance.percent, 
  aes(x = DATE, y = percent, colour = fct_reorder(GEOGRAPHY_NAME,percent,.desc = TRUE))
  ) +
  geom_line() +
  scale_color_brewer(palette = 'Paired') +
  theme(legend.title=element_blank()) +
  scale_x_continuous(breaks = unique(finance.percent$DATE)) +
  ylab('Percent') +
  ggtitle('Percent jobs in finance sectors, GB core cities')


# While we're here, extract a single value from the data to use inline
edinburgh_latestpercent = finance.percent %>% 
  filter(
    DATE == max(DATE),
    GEOGRAPHY_NAME == 'City of Edinburgh'
    ) %>% 
  pull(percent)

```

You can run code in the text too, to include dynamic values from the data, rather than handwriting those in. For instance, we can say that **the percent of Edinburgh's finance jobs in `r max(bres$DATE)` was `r edinburgh_latestpercent`%.**

We can also add tables in from the data:

```{r tableexample}

knitr::kable(
  finance.percent %>% 
    filter(DATE == max(DATE)) %>% 
    select(City = GEOGRAPHY_NAME, percent) %>% 
    arrange(desc(percent)), 
             # escape = FALSE, 
             caption = paste0("Table: core cities' percent of finance jobs in ",max(finance.percent$DATE)," (SICs 64-66)"))

```


## 3. Scraping websites for links, downloading one and extracting from a zip

First we look at occupation percents in England/Wales core cities.

```{r scraping_census}
# From NOMIS webpage bulk Census download
website_basename = "https://www.nomisweb.co.uk"

# We use the xml2 library to read the page
doc = read_html(paste0(website_basename,"/sources/census_2021_bulk"))

# Extract all <a> tags with hrefs containing "census2021*.zip"
zip_links = xml_attr(
  xml_find_all(doc, "//a[contains(@href, 'census2021') and contains(@href, '.zip')]"),
  "href"
)

#Set global timeout to very bigly indeed for parsing and downloading
#ten hours per file covers all eventualities!
options(timeout = 36000)

# AS WE'VE DONE BEFORE, WE DOWNLOAD A ZIP TO A TEMPORARY LOCATION
url = paste0(website_basename,zip_links[qg('TS063', zip_links)][1])
temp_path = tempfile(fileext=".zip")
download.file(url, temp_path, mode="wb")

# LOOK INSIDE ZIP FOLDER - We can see separate files for different geographical scales
# With list = TRUE, this DOES NOT unzip - it just lists the zip folder's contents
filenames = unzip(temp_path, list = TRUE)

# Take a look...
# filenames
# IT CONTAINS SEPARATE CSVS FOR DIFFERNT GEOGRAPHICAL SCALES
# We want lower tier local authorities
# Here's an explanation for Census geographies...
# https://www.ons.gov.uk/census/census2021dictionary/areatypedefinitions

# Pick the geographical scale we want
occ = read_csv(unzip(temp_path, files = filenames$Name[qg('ltla',filenames$Name)]))

# TAKE A LOOK AT THE COLUMN NAMES FOR THAT DATA
# Again, we've got columns that we want to pivot
# colnames(occ)               
occ = occ %>% 
  pivot_longer(
    4:13, names_to = 'occupation', values_to = 'count'
  )


# We then keep just the subgroup occupations
# And find our own percentages for each place
occ = occ %>% 
  filter(
    !qg('all usual',occupation)
    ) %>%
  group_by(geography) %>% 
  mutate(
    percent_occ = (count / sum(count)) * 100
    ) %>% 
  ungroup()

# Confirm sums to 100% per place... tick
# occ %>% group_by(geography) %>% 
#   summarise(percentsum = sum(percent_occ))

# Nicer occupation names
occ = occ %>% 
  mutate(
    occupation = case_when(
      qg('Managers', occupation) ~ 'Managers, directors & senior officials',
      qg('Professional occ', occupation) ~ 'Professionals',
      qg('Associate', occupation) ~ 'Associate professional & technical',
      qg('secretar', occupation) ~ 'Admin & secretarial',
      qg('trades', occupation) ~ 'Skilled trades',
      qg('caring', occupation) ~ 'Caring, leisure & other service',
      qg('sales', occupation) ~ 'Sales & customer service',
      qg('process', occupation) ~ 'Process, plant & machine operatives',
      qg('elementary', occupation) ~ 'Elementary'
    )
  )

# Filter to core cities
# table(corecities %in% occ$geography)
# table(c(corecities,'Cardiff','Newcastle upon Tyne') %in% occ$geography)

# core city names from BRES will work for us here
# We lose Glasgow, Edinburgh and Belfast as it's England and Wales Census data
corecities.bres = readRDS(gzcon(url('https://github.com/DanOlner/RegionalEconomicTools/raw/refs/heads/gh-pages/data/corecities_bres.rds')))

occ.core = occ %>% 
  filter(
    geography %in% corecities.bres
  )


# Might a filled bar is the way forward? Hmm.
ggplot(
  occ.core,
  aes(y = geography, x = percent_occ, fill = occupation)
) +
  geom_bar(position="fill", stat="identity") +
  scale_fill_brewer(palette = 'Paired') +
  ylab("")


```

But how do those compare to average?


```{r census2}

# Get occupation values for England outside London
# To compare differences
occ.all = read_csv(unzip(temp_path, files = filenames$Name[6])) %>% 
  filter(geography != 'London') %>% 
  pivot_longer(
    4:13, names_to = 'occupation', values_to = 'count'
  )

# Sum for rest of England/Wales outside London
# By occupation
occ.all = occ.all %>% 
  group_by(occupation) %>% 
  summarise(count = sum(count)) %>% 
  filter(!qg('usual',occupation)) %>% 
  mutate(percent_occ = (count/sum(count)) * 100)

occ.all = occ.all %>% 
  mutate(
    occupation = case_when(
      qg('Managers', occupation) ~ 'Managers, directors & senior officials',
      qg('Professional occ', occupation) ~ 'Professionals',
      qg('Associate', occupation) ~ 'Associate professional & technical',
      qg('secretar', occupation) ~ 'Admin & secretarial',
      qg('trades', occupation) ~ 'Skilled trades',
      qg('caring', occupation) ~ 'Caring, leisure & other service',
      qg('sales', occupation) ~ 'Sales & customer service',
      qg('process', occupation) ~ 'Process, plant & machine operatives',
      qg('elementary', occupation) ~ 'Elementary'
    )
  )

# Stick them all togeher into one dataframe
# Rows can go on the end, as long as we make sure structure is the same
occ.compare = bind_rows(
  occ.core %>% select(-date,-`geography code`),
  occ.all %>% mutate(geography = 'E+W minus London')
)


# Can now see how core cities differ from average (outside London)
ggplot() +
  geom_point(
    data = occ.compare %>% filter(!qg('e\\+w', geography)), 
    aes(x = percent_occ, y = occupation, colour = geography),
    size = 3,
    shape = 17
    # height = 0.1
    ) +
  scale_color_brewer(palette = 'Paired') +
  geom_point(
    data = occ.compare %>% filter(qg('e\\+w', geography)), 
    aes(x = percent_occ, y = fct_reorder(occupation,percent_occ)),
    size = 7, shape = 3, colour = 'black'
  ) +
  xlab('Occupation % of all employed (18+)') +
  ylab("") +
  ggtitle('Census 2021: core cities % occupations\nEng/Wales minus London % overlaid (cross)')


```










